### 2005-11-14 Report/watershed_paper.tex
Added paper for InsightJournal
darcs-hash:20051114012057-fafb9-6afdc7ac3cd6339c679177e82c0571b49db7ae92.gz

\documentclass{llncs}
\usepackage{graphicx}

\begin{document}
\title{The watershed transform in ITK - discussion and new developments}
\author{Richard Beare{1} \and Gaetan Lehmann{2}}
\institute{Department of Medicine, Moansh University, Australia
\and Biologie du Developpement et de la Reproduction, INRA de Jouy-en-Josas (France)}
\maketitle

\begin{abstract}
This report discusses various definitions and implementations of the
watershed transform as well as providing an introduction to some well
known techniques for applying the watershed transform to practical
problems. The discussion will be focussed on new and existing ITK
classes.
\end{abstract}

\section{Introduction}
The watershed transform is a tool for image segmentation that is fast
and flexible and potentially fairly parameter free. It was originally
derived from a geophysical model of rain falling on a terrain and a variety
of more formal definitions have been devised to allow development of
practical algorithms. A brief discussion of some of these definitions
and algorithms is provided in Section \ref{sect:history}.

Section \ref{sect:usingWT} provides an introduction to a
methodology known as {\em marker based} watershed. Using markers in
watershed transforms is a well established technique in the
morphological image analysis community and has proved very useful for
many applications. The aim of this part of the report is to describe
the technique to a wider audience \footnote{It should also be noted
that the lack of a marker based watershed motivated the recent work on
ITK filter classes.}.

A brief description of two algorithms used to implement the watershed
transform is given in Section \ref{sect:algorithms}.

\section{Image segmentation}
Image segmentation, in the context of this report, may be broadly
defined as finding the part(s) of the image that is of interest in
order to make measurements. This is a basic definition of image
analysis, which is different from image processing and computer
vision\footnote{Image processing produces images, usually for human
consumption, while computer vision aims to produce high level semantic
descriptions of images.}.

It is very rare to achieve a useful segmentation using a single
procedure, even under the most favourable conditions. Successful
segmentation algorithms typically use a carefully constructed
combination of procedures to achieve useful results. Some modern
algorithms, especially the family of routines related to geodesic
active contours, may seem to be an exception to this rule of thumb
because they are often presented as achieving good segmentation
results entirely on their own. However close examination will show
that a variety of preprocessing steps in addition to careful parameter
tuning is necessary to achieve anything useful.

These comments are also relevant to the watershed transform -- it is
very unlikely that any segmentation problem can be solved using the
watershed alone. However it is a very useful and flexible tool that
can be applied in a number of ways. This report will emphasise the
marker based approach to using the watershed transform, but there are
a number of others including watershed trees and region merging
algorithms.

Finally, an important point that should be mentioned, is that the
output of a low level segmentation procedure such as the watershed
transform will typically be a {\em label image}. A label image,
sometimes referred to as a categorical image, has unique values for
each region. For example, if a watershed produces 2 regions, all
pixels belonging to one region would have value $A$, and all belonging
to the other might have value $B$. Unassigned pixels, such as
watershed lines, might have value of $0$.

\section{The watershed transform -- a brief history}
\label{sect:history}
The watershed transform was first proposed by Beucher and
Lantu\'{e}joul\cite{Beucher-Lantu-79} as a geophysical model of rain
falling on a terrain. The idea is that a raindrop falling on a surface
will trickle down the path of steepest descent to a minima. The set of
points on the surface that lead to the same minima is known as a
catchment basin and borders between catchment basins are watershed
lines\footnote{Major geographical features, like rivers or oceans, are
often considered as separated by watershed lines, usually along
mountain ranges.}. If an image is considered as a terrain and divided
into catchment basins then the hope is that each catchment basin would
contain an object of interest. The image representing the terrain will
be referred to as the {\em control surface} in this report.

Translating the model of rain falling on a terrain to an algorithm has
a number of problems\footnote{Although this is what the first ITK
implementation does.} which are discussed in Section
\ref{sect:rainfallmodel}. The alternative is to imagine a flooding
process, with each regional minima\footnote{A regional minima is a set
of pixels surrounded by pixels of higher value.} becoming the source
of a lake. The terrain is progressively flooded and lakes eventually
meet neighboring lakes. Virtual dams are constructed to keep the
neighboring lakes separate as the waterlevel rises. When the image
surface is completely fooded the virtual dams correspond to the
watershed lines\footnote{In 3D neighboring lakes meet at surfaces.}.

A major breakthrough in the implementation of the watershed was made
by Vincent and Soille \cite{Vincent91a} with the introduction of the
first queue based implementation of the watershed transform. This was
especially significant because it allowed similar execution times on
conventional computer systems as was being achieved with specialized
parallel processing systems. The Vincent-Soille algorithm appears to
be used in the matlab image processing toolbox watershed
implementation (based on the observation that the library function has
a {\tt \_vs} suffix).

Another major advance made soon after by Meyer and Beucher
\cite{Beucher93a,Meyer1994a}. Meyer defined the watershed transform in terms of {\em
topographic distances} through {\em lower complete} images. These
definitions rigorously defined the behavior in the presence of
plateaus and lead to some new algorithms based on graph theory. Two
examples of these algorithms are {\em hill-climbing} and {\em image
integration} and will be discussed in more detail in this report.

Readers interested in more technical detail should refer to \cite{Roerdink2000a,Meijster2004a}.

{\em what about the topological watershed?}

\section{Rainfall model}
\label{sect:rainfall}
{\em where should we put this?}  The original concept behind the
watershed transform was rain falling on a terrain and flowing down
paths of steepest descent to local minima. There has been a lot of
work on alternatives to this approach because of the problems with
it. These problems are basically to do with flat zones in an image
which have no paths of steepest descent leading from them, which means
that water would just sit in the one spot. A simple minded
implementation of this idea would therefore produce a separate
catchment basin for each pixel in the flat zone, which is not a very
useful result. The ITK watershed uses a number of steps to avoid this problem:
\begin{enumerate}
\item Label all minima and flat zones.
\item Trace all unlabelled pixels to a labelled one using gradient descent.
\item Check each flat region and find the lowest neighbor. Change the label of the flat region to that of the lowest neighbor.
\end{enumerate}

It should be evident that a number of arbitary choices have been made
here - for example it is impossible to break a flat region into
pieces, which would probably better fit the physical reality of
rainfall (consider a flat topped ridge which should perhaps be split
down the middle). What happens of there are multiple neighboring
regions with equal height - is the labelling decision now somewhat arbitary?

In general the flooding models avoid these dilemas and lead to more
elegant algorithms.

\section{Using the watershed tranform}
\label{sect:usingWT}
\subsection{Classical approach}
The simplest way of using the watershed is to preprocess the image we
want to segment so that the boundaries of our objects are bright (e.g
apply an edge detector) and compute the watershed transform of the
edge image. Watershed lines will correspond to the boundaries and our
problem will be solved. This is rarely useful in practice because
there are always more regional minima than there are objects, either
due to noise or natural variations in the object surfaces. Therefore,
while many watershed lines do lie on significant boundaries, there are
many that don't. Figure \ref{fig:overseg} illustrates the problem. The ITK
watershed implementation provides a way of addressing this problem
using a measure of significance of the boundary and allowing a
threshold to be applied first to reduce the number of local minima.

An Obvious alternative that is very similar in practice is to 
smooth the image so that there are fewer regional minima.
\begin{figure}
\begin{center}
\label{fig:overseg}
\caption{Classical over segmentation.}
\end{center}
\end{figure}

There are some situations where this direct approach can be
useful. The best known of these is probably binary object splitting,
in which the watershed transform is applied to the distance transform
of a mask. This approach is widely used to split touching,
approximately circular objects, such as cells.

\subsection{Watershed from selected minima}
An alternative approach leads us to the concept of markers mentioned
earlier. We will also see that the thresholding approach used by ITK
may be considered as a specific case of the same approach.

Suppose we have preprocessing steps that let select regional minima
that lead to the correct segmentation. How would we modify the
topography of the image so that only the desired regional minima are
present and the watershed lines that remain are the same as they were
before the modification (i.e we only remove the unwanted boundaries)?

This can be achieved by performing a {\em reconstruction by erosion}
or {\em recursive conditional erosion} from the seleted regional
minima. Figure \ref{fig:recon} illustrates the results of apply a
reconstruction by erosion to a 1D image. The original image is shown
in blue and has regional minima A, B, C, D and E. The marker image
which is recursively eroded is shown in black, with a single zero at
regional minimum B. The red plateaus illustrate the parts of the
result image that are above the orginal. The resulting image has only
on regional minima that corresponds to B.
\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.5]{Pics/recon.eps}
\label{fig:recon}
\caption{Reconstruction by erosion from a single regional minimum.}
\end{center}
\end{figure}

A more formal definition of recursive erosion is:
\begin{equation}
\label{eq:recon_erosion}
h_{n+1}=\max\{f, \varepsilon h_n\}
\end{equation}
where $h_0(x)$ is the function, with value 0 at the markers and
$\infty$ otherwise and $\varepsilon$ denotes the erosion
operation.

This procedure may not seem to offer much advantage -- after all how
are we going to decide which minima are useful to us? The key step to
using these ideas in practice is to realize that an arbitary mask
image can be used to create minima in an image in locations we want
simply by multiplying the image by the mask. Recursive erosion of the
original image using that mask then produces a topology to which we
can apply a watershed transform. The mask image is commonly called a
marker image.

This is the fundamental concept behind the use of markers in watershed
based segmentation.

Let's summarize the steps again:

\begin{enumerate}
\item Generate a marker image. The marker image will be used to create 
one, possibly extended, minima inside each object of interest. A
marker image will always need to create at least two minima, and at
least one minima will probably be used to segment the background. The
marker image may be generated via interaction with an operator or by
various preprocessing steps.
\item \label{recipe:min}Impose the minima on the control surface. This 
might be done by multiplying by the marker image (or an inverted
version of it, depending on convention).
\item \label{recipe:erode}Recursively erode the control surface to eliminate all minima besides those introduced in the previous step. This produces a new control surface.
\item Apply the watershed tranform to the new surface.
\end{enumerate}

\subsection{Efficiency issues}
The procedure outlined above introduces a lot of additional steps,
some of which could be reasonably computationally expensive
(i.e. recursive erosion). Other steps have also been mentioned in
passing ({\em lower completion} in Section \ref{sect:history} that may
be potentially time consuming. Fortunately we will discover in Section
\ref{sect:algorithms} that the appropriate choice of algorithm implementing the
watershed transform will carry out these steps implicitly.

\section{Algorithms}
\label{sect:algorithms}
The two that will be outlined here are known as {\em image
integration}\cite{Meyer1994a} and {\em hill climbing}
\cite{Beucher93a}, both of which are classed as ordered
algorithms and are examples of graph flooding approaches. These
algorithms implement the watershed tranform by topographic distance
(or approximate it) and are derived from Dijkstra-Moore minimal path
algorithms of graph theory.

\subsection{Ordered queues}
Both of these implementations depend on ordered queues to avoid the
need for explicit computation of lower complete images and recursive
erosion. An ordered queue is a heirarchical priority queue with two
levels. One level is defined by the pixel gray level, with darker
pixels having higher priority, while the second applies to pixels of
the same grey level and provides a FIFO ordering. The FIFO ordering
leads to a breadth first propogation across plateaus because the
ordering becomes related to distance from the regional minima. This is
the mechanism that avoids the to explicitly compute lower slope.

Early implementations used a special ordered queue for 8 bit data,
which may be where the misconception that a morphological watershed
can only be applied to 8 bit data types arose. The new ITK watershed
algorithms use standard c++ containers (maps and standard queues) to
implement the ordered queue\footnote{One potential problem with
applying this style of algorithm to non integer data is finding the
regional minima. Reconstruction approaches are often used to find
regional minima of integer type images, but subtracting 1, carrying
out a recursive dilation and subtracting the two. The same concept
does not extend to floating point data. However, if regional minima
are available the algorithms we are discussing can be applied to
floating point data.}.

In both algorithms each pixel is regarded as a node in a graph that is
connected to neighboring pixels by pairs of edges. Costs of the edges
are defined in terms of a cost related to the gradient.

\subsection{Image integration}
This algorithm maintains a pair of images -- a distance image and a
label image. The distance image is initialized to initialized to
$\infty$ everywhere, except at the minima where it is initialized to
the value of the minima while the label image is initialized to $0$
everywhere, except at the minima where it is initialized to the index
of the minima (or, more typically, the value of the marker image at
those points).

Wavefronts are propogated from each minima, using the ordered queue
approach. As the wavefront reaches pixel the distance travelled by the
wavefront is compared to the distance stored in the distance image. If
the distance travelled by the new wavefront is less than the stored
value then that pixel is given the label associated with the
wavefront, otherwise the label is left unchanged.

\subsection{Hill climbing}
Hill climbing is an approximation that eliminates the need to maintain
a distance image. The approximation employed is that distances between
all neighbors in a 4, 6, 8, 18 or 26\footnote{4 and 8 connected
neighborhoods are used in 2D, 6, 18 and 26 in 3D} and the center of
the neighborhood are equal -- i.e the various $\sqrt{2}$ terms are
ignored. The position of the watershed lines can have some dependence
on raster ordering because of this. In general the approach is very useful and efficient.

Hill climbing requires a label image which is initialized to the
indexes of the regional minima at the regional minima and an UNKNOWN
value elsewhere\footnote{In the ITK implementation a separate image is
used to track UNKNOWN locations because UNKNOWN values are difficult
to deal with in a type independent way.}. The priority queue is
initialised with the borders of the regional minima. Pixels pushed
onto the queue with priority defined by their grey level and labelled
as they are pushed. When pixels are popped from the queue their
neighbors are checked, labelled, and pushed onto the queue if they
aren't watershed pixels. The lower completion step is carried out
implicitly by not allowing pixels to be pushed onto the queue with a
priority lower than the priority of the neighbor which was popped from
the queue.

The assumption that the distance between a pixel and all of its
neighbors is 1 means that the steepest upper neighbors of a pixel are
simply the unlabelled neighbors.

\section{Using watersheds in ITK}

\section{Timing and performance}
streaming and threading.

\section{Discussion}
There is a level of erroneous folklore surrounding the watershed
transform and it is hoped that this report helps to clear the air.

\bibliographystyle{splncs}
\bibliography{Refnames}

\end{document}

### 2005-11-16 Report/watershed_paper.tex
rjb - minor changes
darcs-hash:20051116061051-fafb9-8f8104079a241dcc835c618f022a50e685724608.gz

@@
 \title{The watershed transform in ITK - discussion and new developments}
-\author{Richard Beare{1} \and Gaetan Lehmann{2}}
+\author{Richard Beare{$^1$} \and Gaetan Lehmann{$^2$}}
 \institute{Department of Medicine, Moansh University, Australia
@@
 
-{\em what about the topological watershed?}
+\subsection{The topological watershed}
+An alternative approach, known as the topological watershed, has
+recently been developed \cite{}. The topological watershed modifies
+the topology of the surface so that all pixels that are not watershed
+lines become regional minima and watershed lines are modified so that
+they are more suitable for use in measuring saliency of boundaries
+between regions. Algorithms implementing the topological watershed
+also exploit priority queues, but don't appear to be able to avoid the
+need for explicit lower completion or reconstruction by erosion.
+
+I don't understand the implications of the topological watershed yet.
 
@@
 which may be where the misconception that a morphological watershed
-can only be applied to 8 bit data types arose. The new ITK watershed
+can only be applied to 8 bit data types arose -- in fact some of the
+early work along these lines was covered by a patent\footnote{I'm told
+that the associated paper is \cite{Meyer91a} and that there may be an
+English version in \cite{Dougherty93a}}. The new ITK watershed
 algorithms use standard c++ containers (maps and standard queues) to
@@
 
+\section{Finding regional minima}
+The marker based watershed filter can be used as a convention
+watershed algorithm by finding the regional minima and using them as
+markers. Regional minima (and maxima) are often useful, and there are
+a number of different ways of finding them:
+\begin{itemize}
+\item Using reconstruction approaches mentioned earlier - 
+the problem with these is that a value of 1 is subtracted from the
+original to produce the image from which the reconstruction is carried
+out, which makes the approach less useful for floating point
+types. The advantage of this approach is that it can be easily
+extended to find {\em image dynamics}, which are often more stable and
+less subject to noise than regional minima.
+\item The way ITK does it, which is related to the connected 
+component labeller. Finds all flat regions and then figures out which
+are plateaus, minima and maxima by looking at the boundary values.
+\item Flooding transform (not sure where this came from) which 
+sets anything not a regional minima to a specified max value. The
+algorithm assumes that the border is set to the maximum value. Pixels
+are visited in raster order. If a pixel $A$ has a neighbour $B$ and
+$P(A) > p(B)$ (i.e grey level of A is greater than B) then all pixels
+connected to $A$ with value $P(A)$ are set to the maximum value. This
+is a recursive algorithm that finds pixels that cannot be a regional
+minimum (because they have a lower neighbour) and then eliminates any
+connected region that also cannot be a regional minima.
+
+The result would then need to be thresholded and labelled to produce
+markers. This approach is not limited by the pixel type, but does
+require a more random visit order when flooding, which could restrict
+performance on larger images.
+\end{itemize}
+
 \section{Using watersheds in ITK}

### 2005-11-29 itkRegionalExtremaImageFilter.h
fixed bug relating to input iterator
darcs-hash:20051129010957-fafb9-2a92ba1c1406e48f40f98d0993e55c467f7bf2ae.gz

@@ -124,6 +124,7 @@ private:
   bool                m_FullyConnected;
 
   typedef typename OutputImageType::IndexType OutIndexType;
+  typedef typename InputImageType::IndexType InIndexType;
   typedef ConstShapedNeighborhoodIterator<InputImageType> CNInputIterator;
   typedef ShapedNeighborhoodIterator<OutputImageType> NOutputIterator;
   typedef std::stack<OutIndexType> IndexStack;

### 2005-11-29 itkRegionalExtremaImageFilter.h
add a flat attribute
darcs-hash:20051129091014-2fc9d-53e22ca6e36b1b6137c6dffdfa93cd60268aadcb.gz

@@ -101,6 +101,8 @@ public:
   itkSetMacro(MarkerValue, typename TInputImage::PixelType);
   itkGetConstReferenceMacro(MarkerValue, typename TInputImage::PixelType);
 
+  itkGetMacro(Flat, bool);
+
 protected:
   RegionalExtremaImageFilter();
   ~RegionalExtremaImageFilter() {};
@@ -122,6 +124,7 @@ private:
   void operator=(const Self&); //purposely not implemented
   typename TInputImage::PixelType m_MarkerValue;
   bool                m_FullyConnected;
+  bool  m_Flat;
 
   typedef typename OutputImageType::IndexType OutIndexType;
   typedef typename InputImageType::IndexType InIndexType;

### 2005-11-29 itkValuedRegionalMaximaImageFilter.h
add Valued prefix to regional minima/maxima filters
darcs-hash:20051129094314-2fc9d-5363594305a509c269e3c99e3cd06c3b9fb7592e.gz

@@ -1,5 +1,5 @@
-#ifndef __itkRegionalMaximaImageFilter_h
-#define __itkRegionalMaximaImageFilter_h
+#ifndef __itkValuedRegionalMaximaImageFilter_h
+#define __itkValuedRegionalMaximaImageFilter_h
 
 #include "itkRegionalExtremaImageFilter.h"
 #include "itkNumericTraits.h"
@@ -23,14 +23,14 @@ public:
 
 
 template <class TInputImage, class TOutputImage>
-class ITK_EXPORT RegionalMaximaImageFilter :
+class ITK_EXPORT ValuedRegionalMaximaImageFilter :
     public
     RegionalExtremaImageFilter<TInputImage, TOutputImage,
 			       Function::GreaterThan<typename TInputImage::PixelType>,
 			       Function::GreaterThan<typename TOutputImage::PixelType>  >
 {
 public:
-  typedef RegionalMaximaImageFilter Self;
+  typedef ValuedRegionalMaximaImageFilter Self;
   typedef RegionalExtremaImageFilter<TInputImage, TOutputImage,
 				     Function::GreaterThan<typename TInputImage::PixelType>,
 				     Function::GreaterThan<typename TOutputImage::PixelType> > Superclass;
@@ -43,19 +43,19 @@ public:
 
 
 protected:
-  RegionalMaximaImageFilter() 
+  ValuedRegionalMaximaImageFilter() 
   {
     SetMarkerValue(NumericTraits<typename TOutputImage::PixelType>::min());
   }
-  virtual ~RegionalMaximaImageFilter() {}
+  virtual ~ValuedRegionalMaximaImageFilter() {}
 
 private:
-  RegionalMaximaImageFilter(const Self&); //purposely not implemented
+  ValuedRegionalMaximaImageFilter(const Self&); //purposely not implemented
   void operator=(const Self&); //purposely not implemented
 
 
 
-}; // end RegionalMaximaImageFilter
+}; // end ValuedRegionalMaximaImageFilter
 
 } //end namespace itk
 #endif

### 2005-11-29 itkValuedRegionalMinimaImageFilter.h

@@ -1,5 +1,5 @@
-#ifndef __itkRegionalMinimaImageFilter_h
-#define __itkRegionalMinimaImageFilter_h
+#ifndef __itkValuedRegionalMinimaImageFilter_h
+#define __itkValuedRegionalMinimaImageFilter_h
 
 #include "itkRegionalExtremaImageFilter.h"
 #include "itkNumericTraits.h"
@@ -23,7 +23,7 @@ public:
 
 
 template <class TInputImage, class TOutputImage>
-class ITK_EXPORT RegionalMinimaImageFilter :
+class ITK_EXPORT ValuedRegionalMinimaImageFilter :
     public
     RegionalExtremaImageFilter<TInputImage, TOutputImage,
 			       Function::LessThan<typename TInputImage::PixelType>,
@@ -31,7 +31,7 @@ class ITK_EXPORT RegionalMinimaImageFilter :
     >
 {
 public:
-  typedef RegionalMinimaImageFilter Self;
+  typedef ValuedRegionalMinimaImageFilter Self;
   typedef RegionalExtremaImageFilter<TInputImage, TOutputImage,
 				     Function::LessThan<typename TInputImage::PixelType>,
 				     Function::LessThan<typename TOutputImage::PixelType>  > Superclass;
@@ -44,19 +44,19 @@ public:
 
 
 protected:
-  RegionalMinimaImageFilter() 
+  ValuedRegionalMinimaImageFilter() 
   {
     SetMarkerValue(NumericTraits<typename TOutputImage::PixelType>::max());
   }
-  virtual ~RegionalMinimaImageFilter() {}
+  virtual ~ValuedRegionalMinimaImageFilter() {}
 
 private:
-  RegionalMinimaImageFilter(const Self&); //purposely not implemented
+  ValuedRegionalMinimaImageFilter(const Self&); //purposely not implemented
   void operator=(const Self&); //purposely not implemented
 
 
 
-}; // end RegionalMinimaImageFilter
+}; // end ValuedRegionalMinimaImageFilter
 
 } //end namespace itk
 #endif

### 2005-11-29 itkValuedRegionalMaximaImageFilter.h
use stl functors
darcs-hash:20051129094616-2fc9d-2b66d1cfa7b2e4be7cd77d8405d3be415604c818.gz

@@ -6,34 +6,18 @@
 
 namespace itk {
 
-namespace Function {
-template <class TInput>
-class GreaterThan
-{
-public:
-  GreaterThan(){}
-  ~GreaterThan(){}
-
-  inline bool operator()(const TInput &A, TInput &B)
-  {
-    return (A>B);
-  }
-};
-} // end namespace Function
-
-
 template <class TInputImage, class TOutputImage>
 class ITK_EXPORT ValuedRegionalMaximaImageFilter :
     public
     RegionalExtremaImageFilter<TInputImage, TOutputImage,
-			       Function::GreaterThan<typename TInputImage::PixelType>,
-			       Function::GreaterThan<typename TOutputImage::PixelType>  >
+			       std::greater<typename TInputImage::PixelType>,
+			       std::greater<typename TOutputImage::PixelType>  >
 {
 public:
   typedef ValuedRegionalMaximaImageFilter Self;
   typedef RegionalExtremaImageFilter<TInputImage, TOutputImage,
-				     Function::GreaterThan<typename TInputImage::PixelType>,
-				     Function::GreaterThan<typename TOutputImage::PixelType> > Superclass;
+				     std::greater<typename TInputImage::PixelType>,
+				     std::greater<typename TOutputImage::PixelType> > Superclass;
 
   typedef SmartPointer<Self>   Pointer;
   typedef SmartPointer<const Self>  ConstPointer;

### 2005-11-29 itkValuedRegionalMinimaImageFilter.h

@@ -6,35 +6,19 @@
 
 namespace itk {
 
-namespace Function {
-template <class TInput>
-class LessThan
-{
-public:
-  LessThan(){}
-  ~LessThan(){}
-
-  inline bool operator()(const TInput &A, TInput &B)
-  {
-    return (A<B);
-  }
-};
-} // end namespace Function
-
-
 template <class TInputImage, class TOutputImage>
 class ITK_EXPORT ValuedRegionalMinimaImageFilter :
     public
     RegionalExtremaImageFilter<TInputImage, TOutputImage,
-			       Function::LessThan<typename TInputImage::PixelType>,
-			       Function::LessThan<typename TOutputImage::PixelType>
+			       std::less<typename TInputImage::PixelType>,
+			       std::less<typename TOutputImage::PixelType>
     >
 {
 public:
   typedef ValuedRegionalMinimaImageFilter Self;
   typedef RegionalExtremaImageFilter<TInputImage, TOutputImage,
-				     Function::LessThan<typename TInputImage::PixelType>,
-				     Function::LessThan<typename TOutputImage::PixelType>  > Superclass;
+				     std::less<typename TInputImage::PixelType>,
+				     std::less<typename TOutputImage::PixelType>  > Superclass;
 
   typedef SmartPointer<Self>   Pointer;
   typedef SmartPointer<const Self>  ConstPointer;

### 2005-11-29 itkRegionalMinimaImageFilter.h
add R-Min filter and tests
darcs-hash:20051129120920-2fc9d-9b4db017a8aed50fe3d12204dabfbe5a9d51d0ac.gz

@@ -1,0 +1,151 @@
+/*=========================================================================
+
+  Program:   Insight Segmentation & Registration Toolkit
+  Module:    $RCSfile: itkRegionalMinimaImageFilter.h,v $
+  Language:  C++
+  Date:      $Date: 2005/08/23 15:09:03 $
+  Version:   $Revision: 1.4 $
+
+  Copyright (c) Insight Software Consortium. All rights reserved.
+  See ITKCopyright.txt or http://www.itk.org/HTML/Copyright.htm for details.
+
+     This software is distributed WITHOUT ANY WARRANTY; without even 
+     the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR 
+     PURPOSE.  See the above copyright notices for more information.
+
+=========================================================================*/
+#ifndef __itkRegionalMinimaImageFilter_h
+#define __itkRegionalMinimaImageFilter_h
+
+#include "itkImageToImageFilter.h"
+
+namespace itk {
+
+/** \class RegionalMinimaImageFilter
+ * \brief TODO
+ *
+ * TODO
+ *
+ * Input image MUST contain only integers. With real values, the minima will not be found
+ * as they should and the output of the filter will be wrong.
+ * Watershed pixel are labeled 0.
+ * TOutputImage should be an integer type.
+ * Labels of output image are in no particular order. You can reorder the 
+ * labels such that object labels are consecutive and sorted based on object
+ * size by passing the output of this filter to a RelabelComponentImageFilter.
+ *
+ * The morphological watershed transform algorithm is described in
+ * Chapter 9.2 of Pierre Soille's book "Morphological Image Analysis:
+ * Principles and Applications", Second Edition, Springer, 2003.
+ *
+ * \author Gaëtan Lehmann. Biologie du Développement et de la Reproduction, INRA de Jouy-en-Josas, France.
+ *
+ * \sa WatershedImageFilter, MorphologicalWatershedFromMarkersImageFilter, RelabelComponentImageFilter
+ * \ingroup ImageEnhancement  MathematicalMorphologyImageFilters
+ */
+template<class TInputImage, class TOutputImage>
+class ITK_EXPORT RegionalMinimaImageFilter : 
+    public ImageToImageFilter<TInputImage, TOutputImage>
+{
+public:
+  /** Standard class typedefs. */
+  typedef RegionalMinimaImageFilter Self;
[…]

### 2005-11-29 itkRegionalMaximaImageFilter.h
add R-Max filter and tests
darcs-hash:20051129123713-2fc9d-1eb52298c5c82a5723f1251035a3e91d15577bb7.gz

@@ -1,0 +1,151 @@
+/*=========================================================================
+
+  Program:   Insight Segmentation & Registration Toolkit
+  Module:    $RCSfile: itkRegionalMaximaImageFilter.h,v $
+  Language:  C++
+  Date:      $Date: 2005/08/23 15:09:03 $
+  Version:   $Revision: 1.4 $
+
+  Copyright (c) Insight Software Consortium. All rights reserved.
+  See ITKCopyright.txt or http://www.itk.org/HTML/Copyright.htm for details.
+
+     This software is distributed WITHOUT ANY WARRANTY; without even 
+     the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR 
+     PURPOSE.  See the above copyright notices for more information.
+
+=========================================================================*/
+#ifndef __itkRegionalMaximaImageFilter_h
+#define __itkRegionalMaximaImageFilter_h
+
+#include "itkImageToImageFilter.h"
+
+namespace itk {
+
+/** \class RegionalMaximaImageFilter
+ * \brief TODO
+ *
+ * TODO
+ *
+ * Input image MUST contain only integers. With real values, the minima will not be found
+ * as they should and the output of the filter will be wrong.
+ * Watershed pixel are labeled 0.
+ * TOutputImage should be an integer type.
+ * Labels of output image are in no particular order. You can reorder the 
+ * labels such that object labels are consecutive and sorted based on object
+ * size by passing the output of this filter to a RelabelComponentImageFilter.
+ *
+ * The morphological watershed transform algorithm is described in
+ * Chapter 9.2 of Pierre Soille's book "Morphological Image Analysis:
+ * Principles and Applications", Second Edition, Springer, 2003.
+ *
+ * \author Gaëtan Lehmann. Biologie du Développement et de la Reproduction, INRA de Jouy-en-Josas, France.
+ *
+ * \sa WatershedImageFilter, MorphologicalWatershedFromMarkersImageFilter, RelabelComponentImageFilter
+ * \ingroup ImageEnhancement  MathematicalMorphologyImageFilters
+ */
+template<class TInputImage, class TOutputImage>
+class ITK_EXPORT RegionalMaximaImageFilter : 
+    public ImageToImageFilter<TInputImage, TOutputImage>
+{
+public:
+  /** Standard class typedefs. */
+  typedef RegionalMaximaImageFilter Self;
[…]

### 2005-11-29 itkRegionalExtremaImageFilter.h
drop setConnectedPixels() method to avoid heavy list copy
darcs-hash:20051129182033-2fc9d-bc59ba02faa7272d8836361550ff3a6e88cc2554.gz

@@ -133,12 +133,6 @@ private:
   typedef std::stack<OutIndexType> IndexStack;
   //typedef NOutputIterator::IndexListType OIndexListType;
 
-  void setConnectedPixels(NOutputIterator &OIt, 
-			  OutputImagePixelType Val,
-			  IndexStack &IS,
-			  const typename NOutputIterator::IndexListType IndexList);
-			  
-
 } ; // end of class
 
 } // end namespace itk

### 2005-11-30 Report/regional_minima_paper.tex
added regional extrema paper
darcs-hash:20051130022841-fafb9-86e18681350d657e51d605d1811539776158a1fd.gz

\documentclass{llncs}
\usepackage{graphicx}

\begin{document}
\title{Finding regional extrema - methods and performance}
\author{Richard Beare{$^1$} \and Gaetan Lehmann{$^2$}}
\institute{Department of Medicine, Moansh University, Australia
\and Biologie du Developpement et de la Reproduction, INRA de Jouy-en-Josas (France)}
\maketitle

\begin{abstract}
Finding regional extrema of images is an important step in a number of
important morphological algorithms, such as some versions of the
watershed transform. Regional extrema may also be important cues of
other tasks, such as splitting objects based on distance transform
information. This report provides an overview of the methods available
in ITK and compares the performance with a new filter.
\end{abstract}

\section{Introduction}
There are two classes of regional extrema - regional maxima and
regional minima. Regional maxima are flat zones that are connected to
pixels of lower value while regional minima are flat zones that are
connected to pixels of higher value.

perhaps mention ordering and how it doesn't really apply to vector pixels.

\section{Filters in ITK}
Watershed -- labelling algorithm.

Reconstruction - relationship with image dynamics.

\section{Comparison of operations}
Labelling scheme will produce a labelled image, which might be
preferrable if the next step is some sort of shape statistics
analysis, or a marker based procedure.

Reconstruction method uses a ``substract 1'' step, which isn't well
defined for all pixel types. However it does allow the option of
placing a significance level on a regional extrema via dynamics.

Flooding method is pixel type independent. It produces an output where
all non extrema pixels are set to the opposite extreme.

\end{document}

### 2005-11-30 itkConnectedComponentAlgorithm.h
first attempt to simplify conectivity
darcs-hash:20051130114747-2fc9d-5df065a06efd3482fee0973b6b767b8e559a7bc6.gz

@@ -1,0 +1,48 @@
+#ifndef __itkConnectedComponentAlgorithm_h
+#define __itkConnectedComponentAlgorithm_h
+
+#include "itkImage.h"
+#include "itkConstShapedNeighborhoodIterator.h"
+#include "itkShapedNeighborhoodIterator.h"
+
+namespace itk
+{
+template< class TIterator >
+TIterator*
+setConnectivity( TIterator* it, bool fullyConnected=false )
+{
+  typename TIterator::OffsetType offset;
+  if( !fullyConnected) 
+    {
+    // only activate the neighbors that are face connected
+    // to the current pixel. do not include the center pixel
+    offset.Fill( 0 );
+    for( unsigned int d=0; d < TIterator::Dimension; ++d )
+      {
+      offset[d] = -1;
+      it->ActivateOffset( offset );
+      offset[d] = 1;
+      it->ActivateOffset( offset );
+      offset[d] = 0;
+      }
+    }
+  else
+    {
+    // activate all neighbors that are face+edge+vertex
+    // connected to the current pixel. do not include the center pixel
+    unsigned int centerIndex = it->GetCenterNeighborhoodIndex();
+    for( unsigned int d=0; d < centerIndex*2 + 1; d++ )
+      {
+      offset = it->GetOffset( d );
+      it->ActivateOffset( offset );
+      }
+    offset.Fill(0);
+    it->DeactivateOffset( offset );
+    }
+  return it;
+}
+
+}
+
+
+#endif

### 2005-12-06 Report/regional_minima_paper.tex
extended discussion
darcs-hash:20051206060304-fafb9-e5eec800d94dff6a9829b799771d48c3238ca838.gz

@@
 \usepackage{graphicx}
+\usepackage{listings}
 
@@
 other tasks, such as splitting objects based on distance transform
-information. This report provides an overview of the methods available
-in ITK and compares the performance with a new filter.
+information\footnote{This work started because we were trying to
+compare different watershed algorithms}. This report provides an
+overview of the methods available in ITK and compares the performance
+with a new filter.
 \end{abstract}
@@
 pixels of lower value while regional minima are flat zones that are
-connected to pixels of higher value.
+connected to pixels of higher value. The notion of higher and lower is
+obviously critcal to the definition of regional maxima and minima,
+which means that it doesn't make sense to apply the definitions
+directly to vector images.
 
-perhaps mention ordering and how it doesn't really apply to vector pixels.
+\section{Filters currently available in ITK}
+Regional extrema are already used inside the ITK watershed
+implementation and may be computed using {\em itkHConvexImageFilter}
+and {\em itkHConcaveImageFilter}.
 
-\section{Filters in ITK}
-Watershed -- labelling algorithm.
+The regional extrema computed in the watershed are not available
+externally (which could obviously be changed). The technique used in
+the watershed filters is based on a labelling approach, very similar
+to that used in the {\em ConnectedComponentImageFilter}, where each
+region of uniform intensity is labelled and checked to determine
+whether it is an extrema or plateau.
 
-Reconstruction - relationship with image dynamics.
+The convex and concave filters are based on the principles of
+morhpological reconstruction. Regional minima are located as follows:
+create a marker image by adding $h$ to the original, carry out a
+reconstruction by erosion of the marker image using the original as
+the control surface, and then, finally, take the difference between
+them. If $h=1$ then any non-zero pixels are regional minima, if $h$ is
+greater than one then the non zero pixels are minima that are at least
+$h$ below the lowest surrounding ridge point. The value $h$ is often
+referred to as the {\em dynamic} of the minima, and is often used to
+give an indication of the importance of the regional minima.
+
+The concave and convex filters are therefore capable of providing more
+information than just the location and value of the regional minima -
+they also provide some clues about the local topology. However this
+additional information comes at a computational cost as well as a
+restriction on the type of images to which the technique can be
+applied. The need to add or subtract $h$, rather than relying on
+equality, means that it is only possible to find extrema when the
+smallest step in image intensity is known. This is difficult to do
+sensibly on floating point image types.
+
+\section{{\em itkValuedRegionalMaximaImageFilter} and {\em itkValuedRegionalMinimaImageFilter}}
+The new filters use a simple flooding approach to find regional
+extrema. They produce an output image with the non minima (maxima) set
+to the maximum (minimum) of the pixel type. The minima or maxima
+retain their original value. 
+The flooding approach is very simple, and proceeds, for minima detection, as follows:
+\begin{itemize}
+\item Copy the input to the output.
+\item Visit each pixel of the input image. 
+   \begin{itemize}	
+   \item If the the corresponding output 
+	 value is not maximal (meaning this pixel hasn't already been
+      visited) then check all the neighbors. 
+    \item If any of the neighboring grey levels are less than the current pixel 
+      value, then this pixel cannot be a regional minima.
+    \item Flood fill the region, in the output image, with the same grey level 
+      as the current pixel that contains the current pixel, with the
+      maximal value for the pixel type.
+    \end{itemize}
+\item Go to next input pixel.
+\end{itemize}
+
+This algorithm requires that the image border is preset to the maximum
+value for the pixel type, which is done using a constant boundary
+condition.
+
+If an image is a constant it will be returned as either a regional
+minimum or a regional maximum, depending on the filter used, because
+the situation is not properly defined. The filter will set a flag if
+this situation arises.
+
+\section{Performance}
+A timing test comparing performance on a $371 \times 371 \times 34$
+image showed that the flooding approach was significantly faster than
+the reconstruction approach. The results achieved on an AMD Athlon 64
+3000+ (1800MHz) with 512Kb cache and 4GB ram were:
+\begin{verbatim}
+> ./perf3D ESCells.img
+#F      concave rmin
+0       20.674  3.54
+1       24.456  5.074
+\end{verbatim}
+This is an approximately 5 fold reduction in computation time. Note
+that the left-most column indicates whether the filters were run in
+fully connected(1) or face connected mode.
+
+There are a few approaches that might be taken to improve performance
+further. For example, that standard face calculator might be used to
+reduce the boundary condition check. Alternatively, if a copy of the
+input image is made, and the border of that is set to the maximum
+value, then there is no need for a boundary check at all. The cost of
+this is a missing border region and probably an extra copy of the
+image.
+
+\section{Results - just to prove that it runs}
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[scale=3]{reg_min_ov.png}
+\caption{Regional minima found using the flooding method}
+\end{center}
+\end{figure}
 
 \section{Comparison of operations}
-Labelling scheme will produce a labelled image, which might be
-preferrable if the next step is some sort of shape statistics
-analysis, or a marker based procedure.
+So which method is ``best''? Well, they are all a bit different, so
+the choice will be a bit application dependent. The labelling approach
+isn't currently available as standalone filter, so it is difficult to
+judge performance. The algorithm requires several steps, which may
+slow it down, but some of the steps are certain to visit the pixels in
+raster order, which can be a big advantage. If the regional extrema
+are going to be labelled anyway then the performance may be
+competitive. The reconstruction based approach is considerably slower,
+but is able to do a lot more. If you are interested in image dynamics
+rather than regional extrema then this is the only way to go. If you
+definitely want extrema then the new filter is currently the fastest
+way to go.
 
-Reconstruction method uses a ``substract 1'' step, which isn't well
-defined for all pixel types. However it does allow the option of
-placing a significance level on a regional extrema via dynamics.
+\section{Sample code}
+The following code is from {\tt vrmin.cxx}, and compares the operation
+of ValuedRegionalMinimaImageFilter and HConcaveImageFilter.
 
-Flooding method is pixel type independent. It produces an output where
-all non extrema pixels are set to the opposite extreme.
+\lstset{language=C,  basicstyle=\small}
+\begin{lstlisting}
+// a test routine for regional extrema using flooding
+#include "itkValuedRegionalMinimaImageFilter.h"
+#include "itkHConcaveImageFilter.h"
+#include "itkMaximumImageFilter.h"
+#include "itkInvertIntensityImageFilter.h"
+#include "itkImageFileReader.h"
+#include "itkImageFileWriter.h"
+#include "itkCommand.h"
+#include <itkRescaleIntensityImageFilter.h>
+#include <itkAndImageFilter.h>
+#include "itkCommand.h"
+
+template < class TFilter >
+class ProgressCallback : public itk::Command
+{
+public:
+  typedef ProgressCallback   Self;
+  typedef itk::Command  Superclass;
+  typedef itk::SmartPointer<Self>  Pointer;
+  typedef itk::SmartPointer<const Self>  ConstPointer;
+
+  itkTypeMacro( IterationCallback, Superclass );
+  itkNewMacro( Self );
+
+  /** Type defining the optimizer. */
+  typedef    TFilter     FilterType;
+
+  /** Method to specify the optimizer. */
+  void SetFilter( FilterType * filter )
+    {
+    m_Filter = filter;
+    m_Filter->AddObserver( itk::ProgressEvent(), this );
+    }
+
+  /** Execute method will print data at each iteration */
+  void Execute(itk::Object *caller, const itk::EventObject & event)
+    {
+    Execute( (const itk::Object *)caller, event);
+    }
+
+  void Execute(const itk::Object *, const itk::EventObject & event)
+    {
+    std::cout << m_Filter->GetNameOfClass() << ": " 
+	<< m_Filter->GetProgress() << std::endl;
+    }
+
+protected:
+  ProgressCallback() {};
+  itk::WeakPointer<FilterType>   m_Filter;
+};
+
+int main(int, char * argv[])
+{
+  const int dim = 2;
+
+  typedef unsigned char PType;
+  typedef itk::Image< PType, dim > IType;
+
+  typedef itk::ImageFileReader< IType > ReaderType;
+  ReaderType::Pointer reader = ReaderType::New();
+  reader->SetFileName( argv[2] );
+
+  typedef itk::ValuedRegionalMinimaImageFilter< IType, IType > FilterType;
+  FilterType::Pointer filter = FilterType::New();
+  filter->SetInput( reader->GetOutput() );
+  filter->SetFullyConnected( atoi(argv[1]) );
+
+  typedef ProgressCallback< FilterType > ProgressType;
+  ProgressType::Pointer progress = ProgressType::New();
+  progress->SetFilter( filter );
+
+  typedef itk::ImageFileWriter< IType > WriterType;
+  WriterType::Pointer writer = WriterType::New();
+  writer->SetInput( filter->GetOutput() );
+  writer->SetFileName( argv[3] );
+  writer->Update();
+
+  // produce the same output with other filters
+  typedef itk::HConcaveImageFilter< IType, IType > ConcaveType;
+  ConcaveType::Pointer concave = ConcaveType::New();
+  concave->SetInput( reader->GetOutput() );
+  concave->SetFullyConnected( atoi(argv[1]) );
+  concave->SetHeight( 1 );
+
+  // concave gives minima with value=1 and others with value=0
+  // rescale the image so we have minima=255 other=0
+  typedef itk::RescaleIntensityImageFilter< IType, IType > RescaleType;
+  RescaleType::Pointer rescale = RescaleType::New();
+  rescale->SetInput( concave->GetOutput() );
+  rescale->SetOutputMaximum( 255 );
+  rescale->SetOutputMinimum( 0 );
+
+  // in the input image, select the values of the pixel at the minima
+  typedef itk::AndImageFilter< IType, IType, IType > AndType;
+  AndType::Pointer a = AndType::New();
+  a->SetInput(0, rescale->GetOutput() );
+  a->SetInput(1, reader->GetOutput() );
+
+  // all pixel which are not minima must have value=255
+  // get the non minima pixel by inverting the rescaled image
+  // minima have value=0 and non minima value=255
+  typedef itk::InvertIntensityImageFilter< IType, IType > InvertType;
+  InvertType::Pointer invert = InvertType::New();
+  invert->SetInput( rescale->GetOutput() );
+
+  // get the highest value from "and" and from invert. 
+  // The minima have value>=0 in and image
+  // and the non minima have a value =0. In invert, 
+  // the non minima have a value=255 and the minima
+  // a value=0
+  typedef itk::MaximumImageFilter< IType, IType, IType > MaxType;
+  MaxType::Pointer max = MaxType::New();
+  max->SetInput(0, invert->GetOutput() );
+  max->SetInput(1, a->GetOutput() );
+
+  WriterType::Pointer writer2 = WriterType::New();
+  writer2->SetInput( max->GetOutput() );
+  writer2->SetFileName( argv[4] );
+  writer2->Update();
+
+  // to verify if the image is flat or not
+  filter->Print( std::cout );
+
+  return 0;
+}
+
+\end{lstlisting}
 

### 2005-12-06 itkRegionalMaximaImageFilter.h
remove non needed classes
darcs-hash:20051206154232-2fc9d-e8d09a045e1ea3e51f640b0052c85e47edce71f1.gz

@@ -1,151 +1,0 @@
-/*=========================================================================
-
-  Program:   Insight Segmentation & Registration Toolkit
-  Module:    $RCSfile: itkRegionalMaximaImageFilter.h,v $
-  Language:  C++
-  Date:      $Date: 2005/08/23 15:09:03 $
-  Version:   $Revision: 1.4 $
-
-  Copyright (c) Insight Software Consortium. All rights reserved.
-  See ITKCopyright.txt or http://www.itk.org/HTML/Copyright.htm for details.
-
-     This software is distributed WITHOUT ANY WARRANTY; without even 
-     the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR 
-     PURPOSE.  See the above copyright notices for more information.
-
-=========================================================================*/
-#ifndef __itkRegionalMaximaImageFilter_h
-#define __itkRegionalMaximaImageFilter_h
-
-#include "itkImageToImageFilter.h"
-
-namespace itk {
-
-/** \class RegionalMaximaImageFilter
- * \brief TODO
- *
- * TODO
- *
- * Input image MUST contain only integers. With real values, the minima will not be found
- * as they should and the output of the filter will be wrong.
- * Watershed pixel are labeled 0.
- * TOutputImage should be an integer type.
- * Labels of output image are in no particular order. You can reorder the 
- * labels such that object labels are consecutive and sorted based on object
- * size by passing the output of this filter to a RelabelComponentImageFilter.
- *
- * The morphological watershed transform algorithm is described in
- * Chapter 9.2 of Pierre Soille's book "Morphological Image Analysis:
- * Principles and Applications", Second Edition, Springer, 2003.
- *
- * \author Gaëtan Lehmann. Biologie du Développement et de la Reproduction, INRA de Jouy-en-Josas, France.
- *
- * \sa WatershedImageFilter, MorphologicalWatershedFromMarkersImageFilter, RelabelComponentImageFilter
- * \ingroup ImageEnhancement  MathematicalMorphologyImageFilters
- */
-template<class TInputImage, class TOutputImage>
-class ITK_EXPORT RegionalMaximaImageFilter : 
-    public ImageToImageFilter<TInputImage, TOutputImage>
-{
-public:
-  /** Standard class typedefs. */
-  typedef RegionalMaximaImageFilter Self;
[…]

### 2005-12-06 itkValuedRegionalMaximaImageFilter.h

@@ -1,45 +1,0 @@
-#ifndef __itkValuedRegionalMaximaImageFilter_h
-#define __itkValuedRegionalMaximaImageFilter_h
-
-#include "itkRegionalExtremaImageFilter.h"
-#include "itkNumericTraits.h"
-
-namespace itk {
-
-template <class TInputImage, class TOutputImage>
-class ITK_EXPORT ValuedRegionalMaximaImageFilter :
-    public
-    RegionalExtremaImageFilter<TInputImage, TOutputImage,
-			       std::greater<typename TInputImage::PixelType>,
-			       std::greater<typename TOutputImage::PixelType>  >
-{
-public:
-  typedef ValuedRegionalMaximaImageFilter Self;
-  typedef RegionalExtremaImageFilter<TInputImage, TOutputImage,
-				     std::greater<typename TInputImage::PixelType>,
-				     std::greater<typename TOutputImage::PixelType> > Superclass;
-
-  typedef SmartPointer<Self>   Pointer;
-  typedef SmartPointer<const Self>  ConstPointer;
-
-  /** Method for creation through the object factory. */
-  itkNewMacro(Self);
-
-
-protected:
-  ValuedRegionalMaximaImageFilter() 
-  {
-    SetMarkerValue(NumericTraits<typename TOutputImage::PixelType>::min());
-  }
-  virtual ~ValuedRegionalMaximaImageFilter() {}
-
-private:
-  ValuedRegionalMaximaImageFilter(const Self&); //purposely not implemented
-  void operator=(const Self&); //purposely not implemented
-
-
-
-}; // end ValuedRegionalMaximaImageFilter
-
-} //end namespace itk
-#endif

### 2005-12-06 itkInvertIntensityImageFilter.h
add files needed for the tests
darcs-hash:20051206163238-2fc9d-ab7ec44dde868466c7c89db53569e68af03e03c4.gz

@@ -1,0 +1,114 @@
+/*=========================================================================
+
+  Program:   Insight Segmentation & Registration Toolkit
+  Module:    $RCSfile: itkInvertIntensityImageFilter.h,v $
+  Language:  C++
+  Date:      $Date: 2004/11/29 22:18:43 $
+  Version:   $Revision: 1.11 $
+
+  Copyright (c) Insight Software Consortium. All rights reserved.
+  See ITKCopyright.txt or http://www.itk.org/HTML/Copyright.htm for details.
+
+     This software is distributed WITHOUT ANY WARRANTY; without even 
+     the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR 
+     PURPOSE.  See the above copyright notices for more information.
+
+=========================================================================*/
+#ifndef __itkInvertIntensityImageFilter_h
+#define __itkInvertIntensityImageFilter_h
+
+#include "itkUnaryFunctorImageFilter.h"
+
+namespace itk
+{
+
+namespace Functor {  
+ 
+template< typename TInput, typename  TOutput>
+class InvertIntensityTransform
+{
+public:
+  typedef typename NumericTraits< TInput >::RealType RealType;
+  InvertIntensityTransform() {}
+  ~InvertIntensityTransform() {}
+  void SetMaximum( TOutput max ) { m_Maximum = max; }
+  inline TOutput operator()( const TInput & x )
+  {
+    TOutput  result = static_cast<TOutput>( m_Maximum - x );
+    return result;
+  }
+private:
+  TInput  m_Maximum;
+}; 
+
+}  // end namespace functor
+
+
+/** \class InvertIntensityImageFilter
+ * \brief Invert intensity of an image 
+ *
+ * InvertIntensityImageFilter invert intensity of pixels by subtracting pixel value
+ * to a maximum value. The maximum value can be set with SetMaximum and defaults
+ * the maximum of input pixel type.
+ * This filter can be used to invert, for example, a binary image, a distance map, etc.
+ *
+ * \author Gaëtan Lehmann. Biologie du Développement et de la Reproduction, INRA de Jouy-en-Josas, France.
+ *
+ * \sa IntensityWindowingImageFilter ShiftScaleImageFilter
+ * \ingroup IntensityImageFilters  Multithreaded
+ *
+ */
+template <typename  TInputImage, typename  TOutputImage=TInputImage>
[…]

### 2005-12-16 article/Article.tex
rename article to Article
darcs-hash:20051216221307-2fc9d-8eb88b9fb532a428aa326c363fb6577dad2b67a0.gz

\documentclass[IJ]{cesj}
\usepackage{graphicx}
\usepackage{listings}

\author{}
\institute{}

\title{}
\abstract{}
\keyword{}
\year{}

\leftmark{}
\rightmark{}

\begin{document}
\lstset{language=c++}
\maketitle
% \tableofcontents

\end{document}

### 2006-01-11 itkMorphologicalWatershedFromMarkersImageFilter.h
add SetInput1 and SetInput2 (to use them in python)
darcs-hash:20060111170600-2fc9d-2a05a6cbe7224236867c86df4877265cf9cdd94d.gz

@@ -67,7 +67,7 @@ public:
   void SetMarkerImage(TLabelImage *input)
      {
      // Process object is not const-correct so the const casting is required.
-     this->SetNthInput(1, const_cast<TLabelImage *>(input) );
+     this->SetNthInput( 1, const_cast<TLabelImage *>(input) );
      }
 
   /** Get the marker image */
@@ -75,6 +75,18 @@ public:
     {
     return static_cast<LabelImageType*>(const_cast<DataObject *>(this->ProcessObject::GetInput(1)));
     }
+
+   /** Set the input image */
+  void SetInput1(TInputImage *input)
+     {
+     this->SetInput( input );
+     }
+
+   /** Set the marker image */
+  void SetInput2(TLabelImage *input)
+     {
+     this->SetMarkerImage( input );
+     }
 
   /**
    * Set/Get whether the connected components are defined strictly by

### 2006-01-11 itkInvertIntensityImageFilter.h
update invert filter
darcs-hash:20060111170644-2fc9d-fb7253a2703a0932656c912a2cfdeb288371ad01.gz

@@ -80,6 +80,9 @@ public:
   typedef typename TInputImage::PixelType  InputPixelType;
   typedef typename NumericTraits<InputPixelType>::RealType RealType;
 
+  /** Runtime information support. */
+  itkTypeMacro(InvertIntensityImageFilter, UnaryFunctorImageFilter);
+
   /** Method for creation through the object factory. */
   itkNewMacro(Self);
   

### 2006-01-11 itkRegionalMinimaImageFilter.h
update r-min filters
darcs-hash:20060111170924-2fc9d-e7dd84589875e17333813b17adcc98307db0616c.gz

@@ -22,25 +22,16 @@
 namespace itk {
 
 /** \class RegionalMinimaImageFilter
- * \brief TODO
+ * \brief Produce a binary image where foreground is the regional minima of the input image
  *
- * TODO
+ * Regional minima are flat zones surounded by pixels of greater value.
  *
- * Input image MUST contain only integers. With real values, the minima will not be found
- * as they should and the output of the filter will be wrong.
- * Watershed pixel are labeled 0.
- * TOutputImage should be an integer type.
- * Labels of output image are in no particular order. You can reorder the 
- * labels such that object labels are consecutive and sorted based on object
- * size by passing the output of this filter to a RelabelComponentImageFilter.
- *
- * The morphological watershed transform algorithm is described in
- * Chapter 9.2 of Pierre Soille's book "Morphological Image Analysis:
- * Principles and Applications", Second Edition, Springer, 2003.
- *
+ * If the input image is constant, the entire image can be considered as a minima or not.
+ * The SetFlatIsMinima() method let the user choose which behavior to use.
+ * 
  * \author Gaëtan Lehmann. Biologie du Développement et de la Reproduction, INRA de Jouy-en-Josas, France.
  *
- * \sa WatershedImageFilter, MorphologicalWatershedFromMarkersImageFilter, RelabelComponentImageFilter
+ * \sa ValuedRegionalMinimaImageFilter, HConcaveImageFilter, RegionalMaximaImageFilter
  * \ingroup ImageEnhancement  MathematicalMorphologyImageFilters
  */
 template<class TInputImage, class TOutputImage>
@@ -90,26 +81,27 @@ public:
   itkGetConstReferenceMacro(FullyConnected, bool);
   itkBooleanMacro(FullyConnected);
   
-  /** Set the value in the image to consider as "foreground". Defaults to
-   * maximum value of PixelType. Subclasses may alias this to
-   * DilateValue or ErodeValue.*/
+  /**
+   * Set/Get the value in the output image to consider as "foreground".
+   * Defaults to maximum value of PixelType.
+   */
   itkSetMacro(ForegroundValue, OutputImagePixelType);
-
-  /** Get the value in the image considered as "foreground". Defaults to
-   * maximum value of PixelType. */
   itkGetConstMacro(ForegroundValue, OutputImagePixelType);
 
[…]

### 2006-01-11 itkValuedRegionalExtremaImageFilter.h

@@ -1,5 +1,5 @@
-#ifndef __itkRegionalExtremaImageFilter_h
-#define __itkRegionalExtremaImageFilter_h
+#ifndef __itkValuedRegionalExtremaImageFilter_h
+#define __itkValuedRegionalExtremaImageFilter_h
 
 #include "itkImageToImageFilter.h"
 #include "itkConstShapedNeighborhoodIterator.h"
@@ -9,12 +9,12 @@
 
 namespace itk {
 
-/** \class RegionalExtremaImageFilter 
+/** \class ValuedRegionalExtremaImageFilter 
  * \brief Uses a flooding algorithm to set all voxels that are not a
  * regional extrema to the max or min of the pixel type. 
  *
- * This is the class used by RegionalMinimaImageFilter and
- * RegionalMaximaImageFilter. There is no supression of regional
+ * This is the class used by ValuedRegionalMinimaImageFilter and
+ * ValuedRegionalMaximaImageFilter. There is no supression of regional
  * minima based on dynamics, as available in HMinimaImageFilter. This
  * flooding algorithm is a very simple one, but I'm not sure where it
  * came from - I certainly didn't invent it.
@@ -43,19 +43,18 @@ namespace itk {
  * \author Richard Beare. Department of Medicine, Monash University,
  * Melbourne, Australia.
  *
- * \sa RegionalMinimaImageFilter, RegionalMaximaImageFilter,
- * HMinimaImageFilter
+ * \sa ValuedRegionalMinimaImageFilter, ValuedRegionalMaximaImageFilter, HMinimaImageFilter
  * \ingroup MathematicalMorphologyImageFilters
  */
 
 template<class TInputImage, class TOutputImage, 
 	 class TFunction1, class TFunction2>
-class ITK_EXPORT RegionalExtremaImageFilter : 
+class ITK_EXPORT ValuedRegionalExtremaImageFilter : 
     public ImageToImageFilter<TInputImage, TOutputImage>
 {
 public:
   /** Standard class typedefs. */
-  typedef RegionalExtremaImageFilter Self;
+  typedef ValuedRegionalExtremaImageFilter Self;
   typedef ImageToImageFilter<TInputImage, TOutputImage>
   Superclass;
   typedef SmartPointer<Self>        Pointer;
@@ -84,7 +83,7 @@ public:
   itkNewMacro(Self);  
 
   /** Runtime information support. */
-  itkTypeMacro(RegionalExtremaImageFilter, 
[…]

### 2006-01-11 itkValuedRegionalMinimaImageFilter.h

@@ -1,22 +1,38 @@
 #ifndef __itkValuedRegionalMinimaImageFilter_h
 #define __itkValuedRegionalMinimaImageFilter_h
 
-#include "itkRegionalExtremaImageFilter.h"
+#include "itkValuedRegionalExtremaImageFilter.h"
 #include "itkNumericTraits.h"
 
 namespace itk {
+/** \class ValuedRegionalMinimaImageFilter
+ * \brief Transforms the image so that any pixel that is not a
+ * regional minima is set to the maximum value for the pixel
+ * type. Pixels that are regional minima retain their value.
+ *
+ * Regional minima are flat zones surrounded by pixels of higher
+ * value. A completely flat image will be marked as a regional minima
+ * by this filter.
+
+ * \author Richard Beare. Department of Medicine, Monash University,
+ * Melbourne, Australia.
+ *
+ * \sa ValuedRegionalMaximaImageFilter, ValuedRegionalExtremaImageFilter, HMinimaImageFilter
+ * \ingroup MathematicalMorphologyImageFilters
+ */
+
 
 template <class TInputImage, class TOutputImage>
 class ITK_EXPORT ValuedRegionalMinimaImageFilter :
     public
-    RegionalExtremaImageFilter<TInputImage, TOutputImage,
+    ValuedRegionalExtremaImageFilter<TInputImage, TOutputImage,
 			       std::less<typename TInputImage::PixelType>,
 			       std::less<typename TOutputImage::PixelType>
     >
 {
 public:
   typedef ValuedRegionalMinimaImageFilter Self;
-  typedef RegionalExtremaImageFilter<TInputImage, TOutputImage,
+  typedef ValuedRegionalExtremaImageFilter<TInputImage, TOutputImage,
 				     std::less<typename TInputImage::PixelType>,
 				     std::less<typename TOutputImage::PixelType>  > Superclass;
 

### 2006-01-11 article/watershed_paper.tex
move parper to article dir
darcs-hash:20060111173128-2fc9d-bcb6183a2aec08a32328edb2f630117e36dcb1eb.gz

--- 2005-11-16 Report/watershed_paper.tex
@@
 
+Preventing boundary checks - not really the itk way, but if a user is
+willing to throw away the image border by treating it as a watershed
+line then there is no need to test for boundary conditions. There
+isn't even any need to use the face calculator approach because border
+pixels will be initially tagged as processed, and therefore we can be
+certain their neighbors are never checked.
+
 \section{Discussion}

### 2006-01-11 article/Article.tex
add bibliography
darcs-hash:20051216221332-2fc9d-345da32161b238dd9c79873f8c5336ec3a651250.gz
switch article template to insight template
darcs-hash:20051217121354-2fc9d-8fe58be6c8bdbe5a9783f22962e935144fd36cbc.gz
add a example of image in the template
darcs-hash:20051229181939-2fc9d-1e7f75036c9155f08e1248a3f564dc024d7d470a.gz
move article to Insight template
darcs-hash:20060111191258-2fc9d-c5d9911ad8110577a1c157a3d48a31b5c00158b6.gz

@@
-\documentclass[IJ]{cesj}
+\documentclass{InsightArticle}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%  hyperref should be the last package to be loaded.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\usepackage[dvips,
+bookmarks,
+bookmarksopen,
+backref,
+colorlinks,linkcolor={blue},citecolor={blue},urlcolor={blue},
+]{hyperref}
 \usepackage{graphicx}
 \usepackage{listings}
+\usepackage{subfigure}
+\title{Finding regional extrema - methods\\ and performance}
 
-\author{}
-\institute{}
+%\release{1.00}
 
-\title{}
-\abstract{}
-\keyword{}
-\year{}
-
-\leftmark{}
-\rightmark{}
+\author{Richard Beare{$^1$} {\small{and}} Ga\"etan Lehmann{$^2$}}
+\authoraddress{{$^1$}Department of Medicine, Moansh University, Australia.\\ 
+{$^2$}Unit\'e de Biologie du D\'eveloppement et de la Reproduction, Institut National de la Recherche Agronomique, 78350 Jouy-en-Josas, France}
 
 \begin{document}
-\lstset{language=c++}
 \maketitle
-% \tableofcontents
+
+\ifhtml
+\chapter*{Front Matter\label{front}}
+\fi
+
+\begin{abstract}
+\noindent
+This report discusses various definitions and implementations of the
+watershed transform as well as providing an introduction to some well
+known techniques for applying the watershed transform to practical
+problems. The discussion will be focussed on new and existing ITK
+classes.
+\end{abstract}
+
+\tableofcontents
+\section{Introduction}
+The watershed transform is a tool for image segmentation that is fast
+and flexible and potentially fairly parameter free. It was originally
+derived from a geophysical model of rain falling on a terrain and a variety
+of more formal definitions have been devised to allow development of
+practical algorithms. A brief discussion of some of these definitions
+and algorithms is provided in Section \ref{sect:history}.
+
+Section \ref{sect:usingWT} provides an introduction to a
+methodology known as {\em marker based} watershed. Using markers in
+watershed transforms is a well established technique in the
+morphological image analysis community and has proved very useful for
+many applications. The aim of this part of the report is to describe
+the technique to a wider audience \footnote{It should also be noted
+that the lack of a marker based watershed motivated the recent work on
+ITK filter classes.}.
+
+A brief description of two algorithms used to implement the watershed
+transform is given in Section \ref{sect:algorithms}.
+
+\section{Image segmentation}
+Image segmentation, in the context of this report, may be broadly
+defined as finding the part(s) of the image that is of interest in
+order to make measurements. This is a basic definition of image
+analysis, which is different from image processing and computer
+vision\footnote{Image processing produces images, usually for human
+consumption, while computer vision aims to produce high level semantic
+descriptions of images.}.
+
+It is very rare to achieve a useful segmentation using a single
+procedure, even under the most favourable conditions. Successful
+segmentation algorithms typically use a carefully constructed
+combination of procedures to achieve useful results. Some modern
+algorithms, especially the family of routines related to geodesic
+active contours, may seem to be an exception to this rule of thumb
+because they are often presented as achieving good segmentation
+results entirely on their own. However close examination will show
+that a variety of preprocessing steps in addition to careful parameter
+tuning is necessary to achieve anything useful.
+
+These comments are also relevant to the watershed transform -- it is
+very unlikely that any segmentation problem can be solved using the
+watershed alone. However it is a very useful and flexible tool that
+can be applied in a number of ways. This report will emphasise the
+marker based approach to using the watershed transform, but there are
+a number of others including watershed trees and region merging
+algorithms.
+
+Finally, an important point that should be mentioned, is that the
+output of a low level segmentation procedure such as the watershed
+transform will typically be a {\em label image}. A label image,
+sometimes referred to as a categorical image, has unique values for
+each region. For example, if a watershed produces 2 regions, all
+pixels belonging to one region would have value $A$, and all belonging
+to the other might have value $B$. Unassigned pixels, such as
+watershed lines, might have value of $0$.
+
+\section{The watershed transform -- a brief history}
+\label{sect:history}
+The watershed transform was first proposed by Beucher and
+Lantu\'{e}joul\cite{Beucher-Lantu-79} as a geophysical model of rain
+falling on a terrain. The idea is that a raindrop falling on a surface
+will trickle down the path of steepest descent to a minima. The set of
+points on the surface that lead to the same minima is known as a
+catchment basin and borders between catchment basins are watershed
+lines\footnote{Major geographical features, like rivers or oceans, are
+often considered as separated by watershed lines, usually along
+mountain ranges.}. If an image is considered as a terrain and divided
+into catchment basins then the hope is that each catchment basin would
+contain an object of interest. The image representing the terrain will
+be referred to as the {\em control surface} in this report.
+
+Translating the model of rain falling on a terrain to an algorithm has
+a number of problems\footnote{Although this is what the first ITK
+implementation does.} which are discussed in Section
+\ref{sect:rainfallmodel}. The alternative is to imagine a flooding
+process, with each regional minima\footnote{A regional minima is a set
+of pixels surrounded by pixels of higher value.} becoming the source
+of a lake. The terrain is progressively flooded and lakes eventually
+meet neighboring lakes. Virtual dams are constructed to keep the
+neighboring lakes separate as the waterlevel rises. When the image
+surface is completely fooded the virtual dams correspond to the
+watershed lines\footnote{In 3D neighboring lakes meet at surfaces.}.
+
+A major breakthrough in the implementation of the watershed was made
+by Vincent and Soille \cite{Vincent91a} with the introduction of the
+first queue based implementation of the watershed transform. This was
+especially significant because it allowed similar execution times on
+conventional computer systems as was being achieved with specialized
+parallel processing systems. The Vincent-Soille algorithm appears to
+be used in the matlab image processing toolbox watershed
+implementation (based on the observation that the library function has
+a {\tt \_vs} suffix).
+
+Another major advance made soon after by Meyer and Beucher
+\cite{Beucher93a,Meyer1994a}. Meyer defined the watershed transform in terms of {\em
+topographic distances} through {\em lower complete} images. These
+definitions rigorously defined the behavior in the presence of
+plateaus and lead to some new algorithms based on graph theory. Two
+examples of these algorithms are {\em hill-climbing} and {\em image
+integration} and will be discussed in more detail in this report.
+
+Readers interested in more technical detail should refer to \cite{Roerdink2000a,Meijster2004a}.
+
+\subsection{The topological watershed}
+An alternative approach, known as the topological watershed, has
+recently been developed \cite{}. The topological watershed modifies
+the topology of the surface so that all pixels that are not watershed
+lines become regional minima and watershed lines are modified so that
+they are more suitable for use in measuring saliency of boundaries
+between regions. Algorithms implementing the topological watershed
+also exploit priority queues, but don't appear to be able to avoid the
+need for explicit lower completion or reconstruction by erosion.
+
+I don't understand the implications of the topological watershed yet.
+
+\section{Rainfall model}
+\label{sect:rainfall}
+{\em where should we put this?}  The original concept behind the
+watershed transform was rain falling on a terrain and flowing down
+paths of steepest descent to local minima. There has been a lot of
+work on alternatives to this approach because of the problems with
+it. These problems are basically to do with flat zones in an image
+which have no paths of steepest descent leading from them, which means
+that water would just sit in the one spot. A simple minded
+implementation of this idea would therefore produce a separate
+catchment basin for each pixel in the flat zone, which is not a very
+useful result. The ITK watershed uses a number of steps to avoid this problem:
+\begin{enumerate}
+\item Label all minima and flat zones.
+\item Trace all unlabelled pixels to a labelled one using gradient descent.
+\item Check each flat region and find the lowest neighbor. Change the label of the flat region to that of the lowest neighbor.
+\end{enumerate}
+
+It should be evident that a number of arbitary choices have been made
+here - for example it is impossible to break a flat region into
+pieces, which would probably better fit the physical reality of
+rainfall (consider a flat topped ridge which should perhaps be split
+down the middle). What happens of there are multiple neighboring
+regions with equal height - is the labelling decision now somewhat arbitary?
+
+In general the flooding models avoid these dilemas and lead to more
+elegant algorithms.
+
+\section{Using the watershed tranform}
+\label{sect:usingWT}
+\subsection{Classical approach}
+The simplest way of using the watershed is to preprocess the image we
+want to segment so that the boundaries of our objects are bright (e.g
+apply an edge detector) and compute the watershed transform of the
+edge image. Watershed lines will correspond to the boundaries and our
+problem will be solved. This is rarely useful in practice because
+there are always more regional minima than there are objects, either
+due to noise or natural variations in the object surfaces. Therefore,
+while many watershed lines do lie on significant boundaries, there are
+many that don't. Figure \ref{fig:overseg} illustrates the problem. The ITK
+watershed implementation provides a way of addressing this problem
+using a measure of significance of the boundary and allowing a
+threshold to be applied first to reduce the number of local minima.
+
+An Obvious alternative that is very similar in practice is to 
+smooth the image so that there are fewer regional minima.
+\begin{figure}
+\begin{center}
+\label{fig:overseg}
+\caption{Classical over segmentation.}
+\end{center}
+\end{figure}
+
+There are some situations where this direct approach can be
+useful. The best known of these is probably binary object splitting,
+in which the watershed transform is applied to the distance transform
+of a mask. This approach is widely used to split touching,
+approximately circular objects, such as cells.
+
+\subsection{Watershed from selected minima}
+An alternative approach leads us to the concept of markers mentioned
+earlier. We will also see that the thresholding approach used by ITK
+may be considered as a specific case of the same approach.
+
+Suppose we have preprocessing steps that let select regional minima
+that lead to the correct segmentation. How would we modify the
+topography of the image so that only the desired regional minima are
+present and the watershed lines that remain are the same as they were
+before the modification (i.e we only remove the unwanted boundaries)?
+
+This can be achieved by performing a {\em reconstruction by erosion}
+or {\em recursive conditional erosion} from the seleted regional
+minima. Figure \ref{fig:recon} illustrates the results of apply a
+reconstruction by erosion to a 1D image. The original image is shown
+in blue and has regional minima A, B, C, D and E. The marker image
+which is recursively eroded is shown in black, with a single zero at
+regional minimum B. The red plateaus illustrate the parts of the
+result image that are above the orginal. The resulting image has only
+on regional minima that corresponds to B.
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[scale=0.5]{recon}
+\label{fig:recon}
+\caption{Reconstruction by erosion from a single regional minimum.}
+\end{center}
+\end{figure}
+
+A more formal definition of recursive erosion is:
+\begin{equation}
+\label{eq:recon_erosion}
+h_{n+1}=\max\{f, \varepsilon h_n\}
+\end{equation}
+where $h_0(x)$ is the function, with value 0 at the markers and
+$\infty$ otherwise and $\varepsilon$ denotes the erosion
+operation.
+
+This procedure may not seem to offer much advantage -- after all how
+are we going to decide which minima are useful to us? The key step to
+using these ideas in practice is to realize that an arbitary mask
+image can be used to create minima in an image in locations we want
+simply by multiplying the image by the mask. Recursive erosion of the
+original image using that mask then produces a topology to which we
+can apply a watershed transform. The mask image is commonly called a
+marker image.
+
+This is the fundamental concept behind the use of markers in watershed
+based segmentation.
+
+Let's summarize the steps again:
+
+\begin{enumerate}
+\item Generate a marker image. The marker image will be used to create 
+one, possibly extended, minima inside each object of interest. A
+marker image will always need to create at least two minima, and at
+least one minima will probably be used to segment the background. The
+marker image may be generated via interaction with an operator or by
+various preprocessing steps.
+\item \label{recipe:min}Impose the minima on the control surface. This 
+might be done by multiplying by the marker image (or an inverted
+version of it, depending on convention).
+\item \label{recipe:erode}Recursively erode the control surface to eliminate all minima besides those introduced in the previous step. This produces a new control surface.
+\item Apply the watershed tranform to the new surface.
+\end{enumerate}
+
+\subsection{Efficiency issues}
+The procedure outlined above introduces a lot of additional steps,
+some of which could be reasonably computationally expensive
+(i.e. recursive erosion). Other steps have also been mentioned in
+passing ({\em lower completion} in Section \ref{sect:history} that may
+be potentially time consuming. Fortunately we will discover in Section
+\ref{sect:algorithms} that the appropriate choice of algorithm implementing the
+watershed transform will carry out these steps implicitly.
+
+\section{Algorithms}
+\label{sect:algorithms}
+The two that will be outlined here are known as {\em image
+integration}\cite{Meyer1994a} and {\em hill climbing}
+\cite{Beucher93a}, both of which are classed as ordered
+algorithms and are examples of graph flooding approaches. These
+algorithms implement the watershed tranform by topographic distance
+(or approximate it) and are derived from Dijkstra-Moore minimal path
+algorithms of graph theory.
+
+\subsection{Ordered queues}
+Both of these implementations depend on ordered queues to avoid the
+need for explicit computation of lower complete images and recursive
+erosion. An ordered queue is a heirarchical priority queue with two
+levels. One level is defined by the pixel gray level, with darker
+pixels having higher priority, while the second applies to pixels of
+the same grey level and provides a FIFO ordering. The FIFO ordering
+leads to a breadth first propogation across plateaus because the
+ordering becomes related to distance from the regional minima. This is
+the mechanism that avoids the to explicitly compute lower slope.
+
+Early implementations used a special ordered queue for 8 bit data,
+which may be where the misconception that a morphological watershed
+can only be applied to 8 bit data types arose -- in fact some of the
+early work along these lines was covered by a patent\footnote{I'm told
+that the associated paper is \cite{Meyer91a} and that there may be an
+English version in \cite{Dougherty93a}}. The new ITK watershed
+algorithms use standard c++ containers (maps and standard queues) to
+implement the ordered queue\footnote{One potential problem with
+applying this style of algorithm to non integer data is finding the
+regional minima. Reconstruction approaches are often used to find
+regional minima of integer type images, but subtracting 1, carrying
+out a recursive dilation and subtracting the two. The same concept
+does not extend to floating point data. However, if regional minima
+are available the algorithms we are discussing can be applied to
+floating point data.}.
+
+In both algorithms each pixel is regarded as a node in a graph that is
+connected to neighboring pixels by pairs of edges. Costs of the edges
+are defined in terms of a cost related to the gradient.
+
+\subsection{Image integration}
+This algorithm maintains a pair of images -- a distance image and a
+label image. The distance image is initialized to initialized to
+$\infty$ everywhere, except at the minima where it is initialized to
+the value of the minima while the label image is initialized to $0$
+everywhere, except at the minima where it is initialized to the index
+of the minima (or, more typically, the value of the marker image at
+those points).
+
+Wavefronts are propogated from each minima, using the ordered queue
+approach. As the wavefront reaches pixel the distance travelled by the
+wavefront is compared to the distance stored in the distance image. If
+the distance travelled by the new wavefront is less than the stored
+value then that pixel is given the label associated with the
+wavefront, otherwise the label is left unchanged.
+
+\subsection{Hill climbing}
+Hill climbing is an approximation that eliminates the need to maintain
+a distance image. The approximation employed is that distances between
+all neighbors in a 4, 6, 8, 18 or 26\footnote{4 and 8 connected
+neighborhoods are used in 2D, 6, 18 and 26 in 3D} and the center of
+the neighborhood are equal -- i.e the various $\sqrt{2}$ terms are
+ignored. The position of the watershed lines can have some dependence
+on raster ordering because of this. In general the approach is very useful and efficient.
+
+Hill climbing requires a label image which is initialized to the
+indexes of the regional minima at the regional minima and an UNKNOWN
+value elsewhere\footnote{In the ITK implementation a separate image is
+used to track UNKNOWN locations because UNKNOWN values are difficult
+to deal with in a type independent way.}. The priority queue is
+initialised with the borders of the regional minima. Pixels pushed
+onto the queue with priority defined by their grey level and labelled
+as they are pushed. When pixels are popped from the queue their
+neighbors are checked, labelled, and pushed onto the queue if they
+aren't watershed pixels. The lower completion step is carried out
+implicitly by not allowing pixels to be pushed onto the queue with a
+priority lower than the priority of the neighbor which was popped from
+the queue.
+
+The assumption that the distance between a pixel and all of its
+neighbors is 1 means that the steepest upper neighbors of a pixel are
+simply the unlabelled neighbors.
+
+\section{Finding regional minima}
+The marker based watershed filter can be used as a convention
+watershed algorithm by finding the regional minima and using them as
+markers. Regional minima (and maxima) are often useful, and there are
+a number of different ways of finding them:
+\begin{itemize}
+\item Using reconstruction approaches mentioned earlier - 
+the problem with these is that a value of 1 is subtracted from the
+original to produce the image from which the reconstruction is carried
+out, which makes the approach less useful for floating point
+types. The advantage of this approach is that it can be easily
+extended to find {\em image dynamics}, which are often more stable and
+less subject to noise than regional minima.
+\item The way ITK does it, which is related to the connected 
+component labeller. Finds all flat regions and then figures out which
+are plateaus, minima and maxima by looking at the boundary values.
+\item Flooding transform (not sure where this came from) which 
+sets anything not a regional minima to a specified max value. The
+algorithm assumes that the border is set to the maximum value. Pixels
+are visited in raster order. If a pixel $A$ has a neighbour $B$ and
+$P(A) > p(B)$ (i.e grey level of A is greater than B) then all pixels
+connected to $A$ with value $P(A)$ are set to the maximum value. This
+is a recursive algorithm that finds pixels that cannot be a regional
+minimum (because they have a lower neighbour) and then eliminates any
+connected region that also cannot be a regional minima.
+
+The result would then need to be thresholded and labelled to produce
+markers. This approach is not limited by the pixel type, but does
+require a more random visit order when flooding, which could restrict
+performance on larger images.
+\end{itemize}
+
+\section{Using watersheds in ITK}
+
+\section{Timing and performance}
+streaming and threading.
+
+Preventing boundary checks - not really the itk way, but if a user is
+willing to throw away the image border by treating it as a watershed
+line then there is no need to test for boundary conditions. There
+isn't even any need to use the face calculator approach because border
+pixels will be initially tagged as processed, and therefore we can be
+certain their neighbors are never checked.
+
+\section{Discussion}
+There is a level of erroneous folklore surrounding the watershed
+transform and it is hoped that this report helps to clear the air.
+
+\appendix
+
+\bibliographystyle{plain}
+\bibliography{InsightJournal}
+\nocite{ITKSoftwareGuide}
 

### 2006-01-18 itkLabelOverlay.h
figure labels, basic color overlay generation, current itk watershed example, plateau image.
darcs-hash:20060118034932-fafb9-a957ed4bca9e80dd37374a0f69e98624d96866d9.gz

@@ -1,0 +1,135 @@
+#ifndef __itkLabelOverlay_h
+#define __itkLabelOverlay_h
+
+template< class TInputPixel, class TLabel, class TRGBPixel >
+class LabelOverlay
+{
+public:
+  LabelOverlay()
+  {
+    
+    TRGBPixel rgbPixel;
+    // the following colors are from "R", and named:
+    // "red"             "green3"          "blue"            "cyan"           
+    //"magenta"         "darkorange1"     "darkgreen"       "blueviolet"     
+    //"brown4"          "navy"            "yellow4"         "violetred1"     
+    //"salmon4"         "turquoise4"      "sienna3"         "darkorchid1"    
+    //"springgreen4"    "mediumvioletred" "orangered3"      "lightseagreen"  
+    //"slateblue"       "deeppink1"       "aquamarine4"     "royalblue1"     
+    //"tomato3"         "mediumblue"      "violetred4"      "darkmagenta"    
+    //"violet"          "red4"           
+    // They are a good selection of distinct colours for plotting and
+    // overlays.
+    
+    rgbPixel.Set(255, 0, 0);
+    m_Colors.push_back( rgbPixel );
+
+    rgbPixel.Set(0, 205, 0);
+    m_Colors.push_back( rgbPixel );
+
+    rgbPixel.Set( 0 , 0 , 255 );
+    m_Colors.push_back( rgbPixel );
+
+    rgbPixel.Set( 0 , 255 , 255 );
+    m_Colors.push_back( rgbPixel );
+
+    rgbPixel.Set( 255 , 0 , 255 );
+    m_Colors.push_back( rgbPixel );
+
+    rgbPixel.Set( 255 , 127 , 0 );
+    m_Colors.push_back( rgbPixel );
+
+    rgbPixel.Set( 0 , 100 , 0 );
+    m_Colors.push_back( rgbPixel );
+
+    rgbPixel.Set( 138 , 43 , 226 );
+    m_Colors.push_back( rgbPixel );
+
+    rgbPixel.Set( 139 , 35 , 35 );
+    m_Colors.push_back( rgbPixel );
+
+    rgbPixel.Set( 0 , 0 , 128 );
+    m_Colors.push_back( rgbPixel );
+
[…]

### 2006-02-20 article/Article.tex
add Refnames to bibliography
darcs-hash:20060111191903-2fc9d-56da7d5590030781b82d5b922c238f7d42c7abb9.gz
fix title
darcs-hash:20060111204358-2fc9d-eb8ddc6da150ce0549440d0e37939efe1d58db78.gz
figure labels, basic color overlay generation, current itk watershed example, plateau image.
darcs-hash:20060118034932-fafb9-a957ed4bca9e80dd37374a0f69e98624d96866d9.gz
More details on hill climbing.
darcs-hash:20060220010825-fafb9-7ec8dff7f53cbcd544ae90efa7b241f696a76ef0.gz

@@
 \usepackage{subfigure}
-\title{Finding regional extrema - methods\\ and performance}
+
+\title{The watershed transform in ITK - discussion and new developments}
 
@@
 
+The topological watershed appears to have been developed with the aim
+of using measures of border significance for merging criteria. This
+work lead to the discovery that it wasn't possible to compute certain
+measures with the flooding approaches used by many implementations.
+
 I don't understand the implications of the topological watershed yet.
@@
 \begin{center}
+\caption{Classical over segmentation.}
 \label{fig:overseg}
-\caption{Classical over segmentation.}
 \end{center}
@@
 \includegraphics[scale=0.5]{recon}
+\caption{Reconstruction by erosion from a single regional minimum.}
 \label{fig:recon}
-\caption{Reconstruction by erosion from a single regional minimum.}
 \end{center}
@@
 ignored. The position of the watershed lines can have some dependence
-on raster ordering because of this. In general the approach is very useful and efficient.
+on raster ordering because of this. In general the approach is very
+useful and efficient.
 
-Hill climbing requires a label image which is initialized to the
-indexes of the regional minima at the regional minima and an UNKNOWN
-value elsewhere\footnote{In the ITK implementation a separate image is
-used to track UNKNOWN locations because UNKNOWN values are difficult
-to deal with in a type independent way.}. The priority queue is
-initialised with the borders of the regional minima. Pixels pushed
-onto the queue with priority defined by their grey level and labelled
-as they are pushed. When pixels are popped from the queue their
-neighbors are checked, labelled, and pushed onto the queue if they
-aren't watershed pixels. The lower completion step is carried out
-implicitly by not allowing pixels to be pushed onto the queue with a
-priority lower than the priority of the neighbor which was popped from
-the queue.
+There seem to be at least 3 algorithms that fall into the
+hill-climbing category.
+
+%Hill climbing requires a label image which is initialized to the
+%indexes of the regional minima at the regional minima and an UNKNOWN
+%value elsewhere\footnote{In the ITK implementation a separate image is
+%used to track UNKNOWN locations because UNKNOWN values are difficult
+%to deal with in a type independent way.}. The priority queue is
+%initialised with the borders of the regional minima. Pixels pushed
+%onto the queue with priority defined by their grey level and labelled
+%as they are pushed. When pixels are popped from the queue their
+%neighbors are checked, labelled, and pushed onto the queue if they
+%aren't watershed pixels. The lower completion step is carried out
+%implicitly by not allowing pixels to be pushed onto the queue with a
+%priority lower than the priority of the neighbor which was popped from
+%the queue.
 
@@
 neighbors is 1 means that the steepest upper neighbors of a pixel are
-simply the unlabelled neighbors.
+simply the unlabelled neighbors. This applies to all 3 algorithms
+mentioned below.
+
+\subsubsection{Algorithm 1}
+\begin{itemize}
+\item lower complete the image
+\item initialize the queue with all pixels except interior of minima.
+\item while (!empty(queue))
+   \begin{itemize}
+	\item Take the pixel $p$ from the head of the queue
+        \item for all pixels $q$ that are steepest upper neighbors of $p$; do
+        \begin{itemize}
+         \item  if $q$ is unlabeled, then
+
+propagate label$(p)$ to label$(q)$;
+
+                else if (label$(q) \ne \mbox{WS})$ and  (label$(q) \ne \mbox{label}(p)$ then
+
+                  label($q$) = WS;
+	\end{itemize}
+
+   \end{itemize}
+\end{itemize}
+
+There are at least two reasons why we shouldn't use this
+approach. Firstly, the lower completion step is explicit, which we'd
+like to avoid and secondly ALL non minima pixels are placed on the
+queue, which means that the queue length is almost the same as the
+number of image pixels, increasing the overhead of queue
+management. We shall see that it is only necessary to have the pixels
+on the borders of the regions in the queue.
+
+In addition to these problems there is no provision for implicit
+imposition of minima for a marker based segmentation approach.
+
+\subsubsection{Algorithm 2}
+Using ordered queues.
+\begin{itemize}
+\item initialize the queue with edges of the regional minima. 
+The edges are pixels that are part of a regional minima with
+neighbours that are background.
+\item while (!empty(queue))
+   \begin{itemize}
+	\item Take the pixel $p$ from the head of the queue
+        \item for all pixels $q$ that are steepest upper neighbors of $p$; do
+         \item  if $q$ is unlabeled then
+
+     		propagate label$(p)$ to label$(q)$;
+
+		place $q$ in the queue
+
+                else if (label$(q) \ne \mbox{WS})$ and  (label$(q) \ne \mbox{label}(p)$ then
+
+                label($q$) = WS;
+	\end{itemize}
+
+\end{itemize}
+
+The queue only contains the borders of the growing regions.
+
+\subsubsection{Algorithm 3}
+Also using ordered queues.
+\begin{itemize}
+\item initialize the queue with background pixels that have 
+      neighbors that are members of a regional minima. In practice
+      this is done by checking all regional minima pixels to see
+      whether they have background neighbors. The background pixels
+      are placed on the queue and the label of the regional minima
+      associated with them.
+\item while (!empty(queue)
+   \begin{itemize}
+	\item Take the pixel $p$ from the head of the queue
+
+         if $p$ is unlabelled;
+
+           apply the associated label
+
+         else
+
+           check whether $p$ is a watershed pixel. $p$ is a watershed
+           pixel if there is more than one label in the neighborhood.
+
+        \item for all pixels $q$ that are steepest upper neighbors of $p$; do
+
+	if $q$ is not in the queue;
+
+           place in queue, associated with label($p$)
+
+           mark as queued
+   \end{itemize}
+\end{itemize}
+
+The difference between algorithms 2 and 3 is the order in which
+labelling and queueing happens.
+
+Algorithm 3 requires slightly more comlexity to store label
+information in the queue.
+
+{\em a summary of our findings comparing the last 2 algorithms}
 
@@
 \bibliographystyle{plain}
-\bibliography{InsightJournal}
+\bibliography{InsightJournal,Refnames.bib}
 \nocite{ITKSoftwareGuide}

### 2006-02-20 itkMorphologicalWatershedImageFilter.h

@@ -26,6 +26,7 @@ namespace itk {
  *
  * TODO
  *
+ * These restrictions on input and output type should no longer apply.
  * Input image MUST contain only integers. With real values, the minima will not be found
  * as they should and the output of the filter will be wrong.
  * Watershed pixel are labeled 0.

### 2006-04-19 article/Article.tex
Timings and connected component fix
Added timings from perf3d to the article. The connected component filter was
using long int internally, which I changed to unsigned long because the
image type used in perf3d was unsigned long.

darcs-hash:20060419045601-fafb9-9021aab8e892d96fe15c002a36efbe2d3dfaab2d.gz

@@
 \usepackage{subfigure}
+\usepackage{pseudocode}
+% \usepackage{algorithm}
+% \usepackage{algorithmic}
 
@@
 \section{Introduction}
+
 The watershed transform is a tool for image segmentation that is fast
@@
 \subsubsection{Algorithm 1}
+
+{\em Is it really needed to present this algorithm given that there is
+no lower completion ? Isn't it enough to notice somewhere that an
+algorithm exist without lower completion, and which require to put
+all the pixels in the queue ?}
+
 \begin{itemize}
@@
 approach. Firstly, the lower completion step is explicit, which we'd
-like to avoid and secondly ALL non minima pixels are placed on the
+like to avoid and secondly {\em all} non minima pixels are placed on the
 queue, which means that the queue length is almost the same as the
@@
 Using ordered queues.
-\begin{itemize}
-\item initialize the queue with edges of the regional minima. 
-The edges are pixels that are part of a regional minima with
-neighbours that are background.
-\item while (!empty(queue))
-   \begin{itemize}
-	\item Take the pixel $p$ from the head of the queue
-        \item for all pixels $q$ that are steepest upper neighbors of $p$; do
-         \item  if $q$ is unlabeled then
 
-     		propagate label$(p)$ to label$(q)$;
-
-		place $q$ in the queue
-
-                else if (label$(q) \ne \mbox{WS})$ and  (label$(q) \ne \mbox{label}(p)$ then
-
-                label($q$) = WS;
-	\end{itemize}
-
-\end{itemize}
+This algorithm does not mark the watershed lines.
 
@@
 
+This algorithm take a grayscale image $I$ and a marker image $M$
+which contains a set of markers with different labels. It can be
+splitted in to steps:
+\begin{enumerate}
+  \item copy the marker image $M$ to the output image, and put
+all the marker pixels $p$ with background neighbors in the
+ordered queue, with the priority $I[p]$.
+  \item get a pixel from the ordered queue at the highest
+priority level, and propagate the label of $p$ to all the unlabeled
+neighbors of $p$. The unlabeled neighbors $q$ are added in the
+ordered queue with priority $I[q]$, if $I[q]$ is greater than
+the priority level of $p$, or with the priority of $p$ otherwise.
+\end{enumerate}
+
+The detailed algorithm is shown in Figure \ref{BeucherAlgorithm}.
+
+\begin{figure}[htbp]
+\centering
+\small
+\begin{pseudocode}[framebox]{BeucherWatershed}{I,M}
+\COMMENT{initialization stage} \\
+queue \GETS \emptyset \\
+\FOREACH p \in M \DO
+\BEGIN
+  \COMMENT{copy the marker image to the output image}\\
+  O[p] \GETS M[p] \\
+  \COMMENT{put the edges of the marker in the queue}\\
+  \IF M[p] \neq WSLABEL
+  \THEN
+  \BEGIN
+    bgNeighbor \GETS \FALSE \\
+    \FOREACH q \in \CALL{N}{p} \DO 
+    \BEGIN
+      \IF M[q] = WSLABEL
+        \THEN bgNeighbor \GETS \TRUE \\
+    \END\\
+    \IF bgNeighbor \STMTNUM{5em}{algo:Beucher:bgNeighbor}
+      \THEN
+      queue[I[p]] \GETS queue[I[p]] \cup \{p\} \\
+  \END
+\END
+\\
+\\
+\COMMENT{flooding stage} \\
+\WHILE queue \neq \emptyset \DO
+\BEGIN
+  v \GETS \CALL{first}{queue}\\
+  \WHILE queue[v] \neq \emptyset \DO
+  \BEGIN
+    p \GETS \CALL{first}{queue[v]} \\
+    \FOREACH q \in \CALL{N}{p} \DO
+    \BEGIN
+    \IF O[q] = WSLABEL
+    \THEN
+    \BEGIN
+      O[q] \GETS O[p] \\
+      v \GETS \CALL{MAX}{I[q], v} \STMTNUM{5em}{algo:Beucher:max} \\
+      queue[v] \GETS queue[v] \cup \{q\} \\
+    \END
+    \END
+  \END
+\END
+\end{pseudocode}
+\caption{Simple watershed algorithm.\label{BeucherAlgorithm} $FIRST(queue)$ extract the highest priority element from the queue.
+$N(p)$ return the list of p's neighbors.
+$MAX(a,b)$ returns the maximum of $a$, and $b$.
+$WSLABEL$ is the label used to define background and watershed pixels (usally $0$).
+The lower completion is done by the statement (\ref{algo:Beucher:max}).
+Determine if a marker pixel before adding it in the queue (\ref{algo:Beucher:bgNeighbor}) is not required, but helps to decrease memory usage in case of large markers.}
+\end{figure}
+
+% \begin{itemize}
+% \item initialize the queue with edges of the regional minima. 
+% The edges are pixels that are part of a regional minima with
+% neighbours that are background.
+% 
+% \item while (!empty(queue))
+%    \begin{itemize}
+% 	\item Take the pixel $p$ from the head of the queue
+%         \item for all pixels $q$ that are steepest upper neighbors of $p$; do
+%          \item  if $q$ is unlabeled then 
+% 
+%      		propagate label$(p)$ to label$(q)$;
+% 	
+% 		place $q$ in the queue
+% 
+%                 else if (label$(q) \ne \mbox{WS})$ and  (label$(q) \ne \mbox{label}(p)$ then
+% 
+%                 label($q$) = WS;
+% 	\end{itemize}
+% 
+% \end{itemize}
+
 \subsubsection{Algorithm 3}
 Also using ordered queues.
-\begin{itemize}
-\item initialize the queue with background pixels that have 
-      neighbors that are members of a regional minima. In practice
-      this is done by checking all regional minima pixels to see
-      whether they have background neighbors. The background pixels
-      are placed on the queue and the label of the regional minima
-      associated with them.
-\item while (!empty(queue)
-   \begin{itemize}
-	\item Take the pixel $p$ from the head of the queue
 
-         if $p$ is unlabelled;
+This algorithm take a grayscale image $I$ and a marker image $M$
+which contains a set of markers with different labels. It can be
+splitted in to steps:
+\begin{enumerate}
+  \item copy the marker image $M$ to the output image, and put
+all the background pixels $p$ with markers in there neighbors in the
+ordered queue, with the priority $I[p]$, and mark $p$ as already in
+the queue. All the marker pixels are marked as already in the queue.
+  \item get a pixel $p$ from the ordered queue at the highest
+priority level. If the neighborhood contains only the same label
+or watershed pixels, mark $p$ with that label. The neighbors $q$ 
+of $p$ which are not already in the queue are added to the queue 
+with priority $I[q]$, if $I[q]$ is greater than the priority level
+of $p$, or with the priority of $p$ otherwise.
+\end{enumerate}
 
-           apply the associated label
+The detailed algorithm is shown in Figure \ref{MeyerAlgorithm}.
 
-         else
+Contrary to the previous algorithm, the label in the output image
+can't be used to determine if the pixel has already been added to
+the queue, and so we need to maintain this status by another way.
+It is the role of $S$ in Figure \ref{MeyerAlgorithm}. We'll discuss
+in \ref{sect:implementation} how this status can be implemented.
 
-           check whether $p$ is a watershed pixel. $p$ is a watershed
-           pixel if there is more than one label in the neighborhood.
+This algorithm mark the watershed lines.
 
-        \item for all pixels $q$ that are steepest upper neighbors of $p$; do
+The queue only contains the borders of the growing regions.
 
-	if $q$ is not in the queue;
+It must be noticed that this algorithm never propagate watershed
+lines, and so produce very thin watershed lines.
 
-           place in queue, associated with label($p$)
+\begin{figure}[htbp]
+\centering
+\small
+\begin{pseudocode}[framebox]{MeyerWatershed}{I,M}
+\COMMENT{initialization stage} \\
+\FOREACH p \in M \DO
+  S[p] \GETS \FALSE \\
+queue \GETS \emptyset \\
+\FOREACH p \in M \DO
+\BEGIN
+  \COMMENT{copy the marker image to the output image}\\
+  O[p] \GETS M[p] \\
+  \IF M[p] \neq WSLABEL
+  \THEN
+  \BEGIN
+    S[p] \GETS \TRUE \\
+    \COMMENT{add all the neighbors of $p$ with WSLABEL to the queue}\\
+    \FOREACH q \in \CALL{N}{p} \DO
+    \BEGIN
+      \IF \neg S[q] \AND M[q] = WSLABEL 
+      \THEN
+      \BEGIN
+        S[q] \GETS \TRUE \\
+        queue[I[q]] \GETS queue[I[q]] \cup \{q\} \\
+      \END
+    \END
+  \END
+\END
+\\
+\\
+\COMMENT{flooding stage} \\
+\WHILE queue \neq \emptyset \DO
+\BEGIN
+  v \GETS \CALL{first}{queue}\\
+  \WHILE queue[v] \neq \emptyset \DO
+  \BEGIN
+    p \GETS \CALL{first}{queue[v]} \\
+    \COMMENT{found the label of $p$ and determine if it is a watershed pixel} \\
+    label \GETS WSLABEL \\
+    watershed \GETS \FALSE \\
+    \FOREACH q \in \CALL{N}{p} \DO
+    \BEGIN
+      \IF O[q] \neq WSLABEL \AND \neg watershed
+      \THEN
+        \BEGIN
+        \IF label \neq WSLABEL \AND O[q] \neq label \\
+        \THEN
+          watershed \GETS \TRUE \\
+        \ELSE
+          label \GETS O[q] \\
+        \END
+    \END \\
+    \COMMENT{if $p$ is a watershed pixel, don't propagate this pixel} \\
+    \IF \neg watershed
+    \THEN
+      \BEGIN
+        \COMMENT{assign the label value to the current pixel} \\
+        O[p] \GETS label \\
+        \COMMENT{put all the neighbors of p in the queue} \\
+        \FOREACH q \in \CALL{N}{p} \DO
+        \BEGIN
+          \IF \neg S[q]
+          \THEN
+            \BEGIN
+            S[q] \GETS \TRUE \\
+            h \GETS \CALL{MAX}{I[q], v} \STMTNUM{5em}{algo:Meyer:max} \\
+            queue[h] \GETS queue[h] \cup \{q\} \\
+            \END
+        \END
+      \END
+  \END
+\END
+\end{pseudocode}
+\caption{Watershed algorithm with watershed lines.\label{MeyerAlgorithm} $FIRST(queue)$ extract the highest priority element from the queue.
+$N(p)$ return the list of p's neighbors.
+$MAX(a,b)$ returns the maximum of $a$, and $b$.
+$WSLABEL$ is the label used to define background and watershed pixels (usally $0$).
+The lower completion is done by the statement (\ref{algo:Meyer:max})}
+\end{figure}
 
-           mark as queued
-   \end{itemize}
-\end{itemize}
+% \begin{itemize}
+% \item initialize the queue with background pixels that have 
+%       neighbors that are members of a regional minima. In practice
+%       this is done by checking all regional minima pixels to see
+%       whether they have background neighbors. The background pixels
+%       are placed on the queue and the label of the regional minima
+%       associated with them.
+% \item while (!empty(queue)
+%    \begin{itemize}
+% 	\item Take the pixel $p$ from the head of the queue
+% 	
+%          if $p$ is unlabelled;
+% 
+%            apply the associated label
+% 
+%          else
+%  
+%            check whether $p$ is a watershed pixel. $p$ is a watershed
+%            pixel if there is more than one label in the neighborhood.
+% 
+%         \item for all pixels $q$ that are steepest upper neighbors of $p$; do
+% 	
+% 	if $q$ is not in the queue;
+% 
+%            place in queue, associated with label($p$)
+%         
+%            mark as queued
+%    \end{itemize}
+% \end{itemize}
 
@@
 
-Algorithm 3 requires slightly more comlexity to store label
+Algorithm 3 requires slightly more complexity to store label
 information in the queue.
@@
 {\em a summary of our findings comparing the last 2 algorithms}
+
+The new work implements algorithms 2 and 3 -- algorithm 2 is used when
+the watershed lines are not marked while algorithm 3 is used when they
+are.
 
@@
 
-\section{Using watersheds in ITK}
+\section{Implementation}
+\label{sect:implementation}
+
+We provide with this article two new classes:
+
+  \subsection{MorphologicalWatershedFromMarkersImageFilter}
+
+\verb$MorphologicalWatershedFromMarkersImageFilter$ implement the
+algorithms \ref{sect:BeucherAlgorithm} and \ref{sect:MeyerAlgorithm}.
+They are accessible respectively when the attribute \verb$MarkWatershed$
+is \verb$false$ or is \verb$true$. The filter is templated over the
+input image type, which can be any scalar type, and the label image type
+which should be an integer. The filter take 2 inputs:
+\begin{itemize}
+  \item a grayscale image which can be set with \verb$SetInput(image)$,
+\verb$SetInput(0, image)$ or \verb$SetInput1(image)$.
+  \item a marker image which can be set with \verb$SetMarkerImage(image)$
+and \verb$SetInput2(image)$. \verb$SetInput(0, image)$ should be avoided
+because it will fail if the input image and the marker image are not of the
+same type. The markers are all the non-zero values in the image.
+\end{itemize}
+
+The algorithm \ref{sect:MeyerAlgorithm} require to keep the status of all
+the pixels: {\em already added to the queue} or {\em not yet added to the queue}.
+This task is often done by using negative values in the output image
+for the pixels not yet added to the queue. In our implementation, we have
+chosen to use a binary image to store this, while labels always take values
+above 0 and watershed lines are marked with value 0. The negative values
+option have several drawbacks:
+\begin{itemize}
+  \item it requires to use a different image type than the label type defined
+by the user in case the user have chosen an unsigned pixel type.
+  \item it requires to use one of the value which can be possibly aready used
+by a label in the marker image, in case of a signed type.
+\end{itemize}
+Those problems can be avoided\footnote{It's not true for the unsigned long
+pixel type} by using a type which contains a larger range of value than the
+type choosed by the user, but then, we must copy the data to an image of
+the type chosen by the user at the end of the algorithm.
+
+We prefered using a separated binary image (of type \verb$itk::Image<bool, dim>$)
+to store this information. This solution avoid the need to copy the image at
+the end, and so should be more efficient in computation time, and require only
+one bit per pixel and so should be more memory efficient as well.
+
+The connectivity is defined using the usual ITK way: \verb$SetFullConnected(false)$
+(4-connected in 2D, 6-connected in 3D) or \verb$SetFullConnected(true)$ (8-connected
+in 2D, 26-connected in 3D)\footnote{One should regret to not have the 18-connected
+connectivity in 3D}.
+
+  \subsection{MorphologicalWatershedImageFilter}
+
+\verb$MorphologicalWatershedImageFilter$ perform a watershed transformation
+without marker. It encapsulate several filters:
+\begin{itemize}
+  \item \verb$HMinimaImageFilter$
+  \item \verb$RegionalMinimaImageFilter$
+  \item \verb$ConnectedComponentImageFilter$
+  \item \verb$MorphologicalWatershedFromMarkersImageFilter$
+\end{itemize}
+While \verb$RegionalMinimaImageFilter$, \verb$ConnectedComponentImageFilter$,
+and \verb$MorphologicalWatershedFromMarkersImageFilter$ are the usal steps
+of the waterhed usage, the \verb$HMinimaImageFilter$ is a convenient way
+to avoid the classical watershed's over segmentation by allowing the user
+to select the dynamic of the minima. The minimum dynamic of minima can be
+selected with \verb$SetThreshold()$.
 
@@
 
-Preventing boundary checks - not really the itk way, but if a user is
+\subsection{Other performance issues/comments}
+\begin{itemize}
+\item Preventing boundary checks - not really the itk way, but if a user is
 willing to throw away the image border by treating it as a watershed
@@
 certain their neighbors are never checked.
+\item Streaming and threading  - queue based algorithms are generally difficult 
+to thread efficiently and generally need to have all of the input
+available so prevent streaming from working well. The existing ITK
+watershed does provide a multithreaded implementation and similar
+techniques may be possible with the new implementation. Other
+possibilities include using threads to improve performance of queues.
+\end{itemize}
+
+\section{Using watersheds in ITK}
+
+\begin{figure}[htbp]
+\begin{center}
+\subfigure[Input image]{\includegraphics[scale=4]{threshold}}
+\subfigure[No minima threshold]{\includegraphics[scale=4]{threshold00-rgb}}
+\subfigure[Minima threshold = 10]{\includegraphics[scale=4]{threshold10-rgb}}
+\subfigure[Minima threshold = 20]{\includegraphics[scale=4]{threshold20-rgb}}
+\subfigure[Minima threshold = 30 and Minima threshold = 40]{\includegraphics[scale=4]{threshold30-rgb}}
+\subfigure[Minima threshold = 50]{\includegraphics[scale=4]{threshold50-rgb}}
+\subfigure[markers]{\includegraphics[scale=4]{threshold-markers}}
+\subfigure[Watershed from markers]{\includegraphics[scale=4]{threshold-markersMf-rgb}}
+\caption{Different results of watershed on an articial image.
+The labeled image are overlayed to the input image with {\em LabelOverlayImageFilter}.
+From $(b)$ to $(f)$, an example of reduction of the classical watershed over segmentation by minima dynamic filtering.
+$(g)$ and $(h)$ show a marker defined by a user, and the result over the input image.
+All the watersheds were performed with $MarkWatershed = true$ and $FullyConnected = false$.}
+\end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+\begin{center}
+\subfigure[Input image]{\includegraphics[scale=4]{threshold}}
+\subfigure[No minima threshold]{\includegraphics[scale=4]{threshold00-itk-rgb}}
+\subfigure[Minima threshold = 10]{\includegraphics[scale=4]{threshold10-itk-rgb}}
+% \subfigure[Minima threshold = 20]{\includegraphics[scale=4]{threshold20-itk-rgb}}
+\subfigure[Minima threshold = 40]{\includegraphics[scale=4]{threshold40-itk-rgb}}
+\subfigure[Minima threshold = 63]{\includegraphics[scale=4]{threshold63-itk-rgb}}
+\subfigure[Minima threshold = 65]{\includegraphics[scale=4]{threshold65-itk-rgb}}
+\subfigure[markers]{\includegraphics[scale=4]{threshold-markers}}
+\subfigure[Watershed from markers]{\includegraphics[scale=4]{threshold-markers-itk-rgb}}
+\caption{Different results of watershed on an articial image.
+The labeled image are overlayed to the input image with {\em LabelOverlayImageFilter}.
+From $(b)$ to $(f)$, an example of reduction of the classical watershed over segmentation by minima dynamic filtering.
+$(g)$ and $(h)$ show a marker defined by a user, and the result over the input image..}
+\end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+\begin{center}
+\subfigure[]{\includegraphics[scale=10]{pass-values}}
+\subfigure[]{\includegraphics[scale=10]{pass-values-mf-rgb}}
+\subfigure[]{\includegraphics[scale=10]{pass-values-mF-rgb}}
+\subfigure[]{\includegraphics[scale=10]{pass-values-Mf-rgb}}
+\subfigure[]{\includegraphics[scale=10]{pass-values-MF-rgb}}
+\subfigure[]{\includegraphics[scale=10]{pass-values-itk-rgb}}
+
+\subfigure[]{\includegraphics[scale=4]{plateau}}
+\subfigure[]{\includegraphics[scale=4]{plateau-mf-rgb}}
+\subfigure[]{\includegraphics[scale=4]{plateau-mF-rgb}}
+\subfigure[]{\includegraphics[scale=4]{plateau-Mf-rgb}}
+\subfigure[]{\includegraphics[scale=4]{plateau-MF-rgb}}
+\subfigure[]{\includegraphics[scale=4]{plateau-itk-rgb}}
+
+\subfigure[]{\includegraphics[scale=10]{thick-lines}}
+\subfigure[]{\includegraphics[scale=10]{thick-lines-mf-rgb}}
+\subfigure[]{\includegraphics[scale=10]{thick-lines-mF-rgb}}
+\subfigure[]{\includegraphics[scale=10]{thick-lines-Mf-rgb}}
+\subfigure[]{\includegraphics[scale=10]{thick-lines-MF-rgb}}
+\subfigure[]{\includegraphics[scale=10]{thick-lines-itk-rgb}}
+
+\subfigure[]{\includegraphics[scale=6]{button-hole}}
+\subfigure[]{\includegraphics[scale=6]{button-hole-mf-rgb}}
+\subfigure[]{\includegraphics[scale=6]{button-hole-mF-rgb}}
+\subfigure[]{\includegraphics[scale=6]{button-hole-Mf-rgb}}
+\subfigure[]{\includegraphics[scale=6]{button-hole-MF-rgb}}
+\subfigure[]{\includegraphics[scale=6]{button-hole-itk-rgb}}
+
+\caption{}
+\end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+\begin{center}
+\subfigure[]{\includegraphics[scale=0.5]{cthead1}}
+\subfigure[]{\includegraphics[scale=0.5]{cthead1-markers}}
+\subfigure[]{\includegraphics[scale=0.5]{cthead1mf-rgb}}
+\subfigure[]{\includegraphics[scale=0.5]{cthead1mF-rgb}}
+\subfigure[]{\includegraphics[scale=0.5]{cthead1Mf-rgb}}
+\subfigure[]{\includegraphics[scale=0.5]{cthead1MF-rgb}}
+\subfigure[]{\includegraphics[scale=0.5]{cthead1itk-rgb}}
+\caption{}
+\end{center}
+\end{figure}
 

### 2006-04-19 itkConnectedComponentImageFilter.h

@@ -96,7 +96,7 @@ public:
   itkBooleanMacro(FullyConnected);
   
   // only set after completion
-  itkGetConstReferenceMacro(ObjectCount, long);
+  itkGetConstReferenceMacro(ObjectCount, unsigned long);
 
   // Concept checking -- input and output dimensions must be the same
   itkConceptMacro(SameDimension,
@@ -132,7 +132,7 @@ protected:
 private:
 
   bool m_FullyConnected;
-  long m_ObjectCount;
+  unsigned long m_ObjectCount;
   // some additional types
   typedef typename TOutputImage::RegionType::SizeType OutSizeType;
 
@@ -142,7 +142,7 @@ private:
     public:
       long int length;  // run length information - may be a more type safe way of doing this
       typename InputImageType::IndexType where;  // Index of the start of the run
-      long int label; // the initial label of the run
+      unsigned long int label; // the initial label of the run
     };
 
   typedef std::vector<runLength> lineEncoding;
@@ -153,18 +153,18 @@ private:
   typedef std::vector<long> OffsetVec;
 
   // the types to support union-find operations
-  typedef std::vector<long int> UnionFindType;
+  typedef std::vector<unsigned long int> UnionFindType;
   UnionFindType m_UnionFind;
   UnionFindType m_Consecutive;
   // functions to support union-find operations
-  void InitUnion(const long int size) 
+  void InitUnion(const unsigned long int size) 
   {
     m_UnionFind = UnionFindType(size + 1);
   }
-  void InsertSet(const long int label);
-  long int LookupSet(const long int label);
-  void LinkLabels(const long int lab1, const long int lab2);
-  long int CreateConsecutive();
+  void InsertSet(const unsigned long int label);
+  unsigned long int LookupSet(const unsigned long int label);
+  void LinkLabels(const unsigned long int lab1, const unsigned long int lab2);
+  unsigned long int CreateConsecutive();
   //////////////////
   void CompareLines(lineEncoding &current, const lineEncoding &Neighbour);
 

### 2006-04-23 article/Article.tex
add algorithms 2 and 3 pseudocode
darcs-hash:20060423124020-2fc9d-c3d5616c7aa5a6eabd1c501041bfcee086aaa0a4.gz

--- 2006-02-20 article/Article.tex
@@
 \subsubsection{Algorithm 2}
+
 Using ordered queues.
-\begin{itemize}
-\item initialize the queue with edges of the regional minima. 
-The edges are pixels that are part of a regional minima with
-neighbours that are background.
-\item while (!empty(queue))
-   \begin{itemize}
-	\item Take the pixel $p$ from the head of the queue
-        \item for all pixels $q$ that are steepest upper neighbors of $p$; do
-         \item  if $q$ is unlabeled then
 
-     		propagate label$(p)$ to label$(q)$;
-
-		place $q$ in the queue
-
-                else if (label$(q) \ne \mbox{WS})$ and  (label$(q) \ne \mbox{label}(p)$ then
-
-                label($q$) = WS;
-	\end{itemize}
-
-\end{itemize}
+This algorithm does not mark the watershed lines.
 
@@
 
+This algorithm take a grayscale image $I$ and a marker image $M$
+which contains a set of markers with different labels. It can be
+splitted in to steps:
+\begin{enumerate}
+  \item copy the marker image $M$ to the output image, and put
+all the marker pixels $p$ with background neighbors in the
+ordered queue, with the priority $I[p]$.
+  \item get a pixel from the ordered queue at the highest
+priority level, and propagate the label of $p$ to all the unlabeled
+neighbors of $p$. The unlabeled neighbors $q$ are added in the
+ordered queue with priority $I[q]$, if $I[q]$ is greater than
+the priority level of $p$, or with the priority of $p$ otherwise.
+\end{enumerate}
+
+The detailed algorithm is shown in Figure \ref{BeucherAlgorithm}.
+
+\begin{figure}[htbp]
+\centering
+\small
+\begin{pseudocode}[framebox]{BeucherWatershed}{I,M}
+\COMMENT{initialization stage} \\
+queue \GETS \emptyset \\
+\FOREACH p \in M \DO
+\BEGIN
+  \COMMENT{copy the marker image to the output image}\\
+  O[p] \GETS M[p] \\
+  \COMMENT{put the edges of the marker in the queue}\\
+  \IF M[p] \neq WSLABEL
+  \THEN
+  \BEGIN
+    bgNeighbor \GETS \FALSE \\
+    \FOREACH q \in \CALL{N}{p} \DO 
+    \BEGIN
+      \IF M[q] = WSLABEL
+        \THEN bgNeighbor \GETS \TRUE \\
+    \END\\
+    \IF bgNeighbor \STMTNUM{5em}{algo:Beucher:bgNeighbor}
+      \THEN
+      queue[I[p]] \GETS queue[I[p]] \cup \{p\} \\
+  \END
+\END
+\\
+\\
+\COMMENT{flooding stage} \\
+\WHILE queue \neq \emptyset \DO
+\BEGIN
+  v \GETS \CALL{first}{queue}\\
+  \WHILE queue[v] \neq \emptyset \DO
+  \BEGIN
+    p \GETS \CALL{first}{queue[v]} \\
+    \FOREACH q \in \CALL{N}{p} \DO
+    \BEGIN
+    \IF O[q] = WSLABEL
+    \THEN
+    \BEGIN
+      O[q] \GETS O[p] \\
+      v \GETS \CALL{MAX}{I[q], v} \STMTNUM{5em}{algo:Beucher:max} \\
+      queue[v] \GETS queue[v] \cup \{q\} \\
+    \END
+    \END
+  \END
+\END
+\end{pseudocode}
+\caption{Simple watershed algorithm.\label{BeucherAlgorithm} $FIRST(queue)$ extract the highest priority element from the queue.
+$N(p)$ return the list of p's neighbors.
+$MAX(a,b)$ returns the maximum of $a$, and $b$.
+$WSLABEL$ is the label used to define background and watershed pixels (usally $0$).
+The lower completion is done by the statement (\ref{algo:Beucher:max}).
+Determine if a marker pixel before adding it in the queue (\ref{algo:Beucher:bgNeighbor}) is not required, but helps to decrease memory usage in case of large markers.}
+\end{figure}
+
+% \begin{itemize}
+% \item initialize the queue with edges of the regional minima. 
+% The edges are pixels that are part of a regional minima with
+% neighbours that are background.
+% 
+% \item while (!empty(queue))
+%    \begin{itemize}
+% 	\item Take the pixel $p$ from the head of the queue
+%         \item for all pixels $q$ that are steepest upper neighbors of $p$; do
+%          \item  if $q$ is unlabeled then 
+% 
+%      		propagate label$(p)$ to label$(q)$;
+% 	
+% 		place $q$ in the queue
+% 
+%                 else if (label$(q) \ne \mbox{WS})$ and  (label$(q) \ne \mbox{label}(p)$ then
+% 
+%                 label($q$) = WS;
+% 	\end{itemize}
+% 
+% \end{itemize}
+
 \subsubsection{Algorithm 3}
+
 Also using ordered queues.
-\begin{itemize}
-\item initialize the queue with background pixels that have 
-      neighbors that are members of a regional minima. In practice
-      this is done by checking all regional minima pixels to see
-      whether they have background neighbors. The background pixels
-      are placed on the queue and the label of the regional minima
-      associated with them.
-\item while (!empty(queue)
-   \begin{itemize}
-	\item Take the pixel $p$ from the head of the queue
 
-         if $p$ is unlabelled;
+This algorithm take a grayscale image $I$ and a marker image $M$
+which contains a set of markers with different labels. It can be
+splitted in to steps:
+\begin{enumerate}
+  \item copy the marker image $M$ to the output image, and put
+all the background pixels $p$ with markers in there neighbors in the
+ordered queue, with the priority $I[p]$, and mark $p$ as already in
+the queue. All the marker pixels are marked as already in the queue.
+  \item get a pixel $p$ from the ordered queue at the highest
+priority level. If the neighborhood contains only the same label
+or watershed pixels, mark $p$ with that label. The neighbors $q$ 
+of $p$ which are not already in the queue are added to the queue 
+with priority $I[q]$, if $I[q]$ is greater than the priority level
+of $p$, or with the priority of $p$ otherwise.
+\end{enumerate}
 
-           apply the associated label
+The detailed algorithm is shown in Figure \ref{MeyerAlgorithm}.
 
-         else
+Contrary to the previous algorithm, the label in the output image
+can't be used to determine if the pixel has already been added to
+the queue, and so we need to maintain this status by another way.
+It is the role of $S$ in Figure \ref{MeyerAlgorithm}. We'll discuss
+in ??? how this status can be implemented.
 
-           check whether $p$ is a watershed pixel. $p$ is a watershed
-           pixel if there is more than one label in the neighborhood.
+This algorithm mark the watershed lines.
 
-        \item for all pixels $q$ that are steepest upper neighbors of $p$; do
+The queue only contains the borders of the growing regions.
 
-	if $q$ is not in the queue;
+It must be noticed that this algorithm never propagate watershed
+lines, and so produce very thin watershed lines.
 
-           place in queue, associated with label($p$)
+\begin{figure}[htbp]
+\centering
+\small
+\begin{pseudocode}[framebox]{MeyerWatershed}{I,M}
+\COMMENT{initialization stage} \\
+\FOREACH p \in M \DO
+  S[p] \GETS \FALSE \\
+queue \GETS \emptyset \\
+\FOREACH p \in M \DO
+\BEGIN
+  \COMMENT{copy the marker image to the output image}\\
+  O[p] \GETS M[p] \\
+  \IF M[p] \neq WSLABEL
+  \THEN
+  \BEGIN
+    S[p] \GETS \TRUE \\
+    \COMMENT{add all the neighbors of $p$ with WSLABEL to the queue}\\
+    \FOREACH q \in \CALL{N}{p} \DO
+    \BEGIN
+      \IF \neg S[q] \AND M[q] = WSLABEL 
+      \THEN
+      \BEGIN
+        S[q] \GETS \TRUE \\
+        queue[I[q]] \GETS queue[I[q]] \cup \{q\} \\
+      \END
+    \END
+  \END
+\END
+\\
+\\
+\COMMENT{flooding stage} \\
+\WHILE queue \neq \emptyset \DO
+\BEGIN
+  v \GETS \CALL{first}{queue}\\
+  \WHILE queue[v] \neq \emptyset \DO
+  \BEGIN
+    p \GETS \CALL{first}{queue[v]} \\
+    \COMMENT{found the label of $p$ and determine if it is a watershed pixel} \\
+    label \GETS WSLABEL \\
+    watershed \GETS \FALSE \\
+    \FOREACH q \in \CALL{N}{p} \DO
+    \BEGIN
+      \IF O[q] \neq WSLABEL \AND \neg watershed
+      \THEN
+        \BEGIN
+        \IF label \neq WSLABEL \AND O[q] \neq label \\
+        \THEN
+          watershed \GETS \TRUE \\
+        \ELSE
+          label \GETS O[q] \\
+        \END
+    \END \\
+    \COMMENT{if $p$ is a watershed pixel, don't propagate this pixel} \\
+    \IF \neg watershed
+    \THEN
+      \BEGIN
+        \COMMENT{assign the label value to the current pixel} \\
+        O[p] \GETS label \\
+        \COMMENT{put all the neighbors of p in the queue} \\
+        \FOREACH q \in \CALL{N}{p} \DO
+        \BEGIN
+          \IF \neg S[q]
+          \THEN
+            \BEGIN
+            S[q] \GETS \TRUE \\
+            h \GETS \CALL{MAX}{I[q], v} \STMTNUM{5em}{algo:Meyer:max} \\
+            queue[h] \GETS queue[h] \cup \{q\} \\
+            \END
+        \END
+      \END
+  \END
+\END
+\end{pseudocode}
+\caption{Watershed algorithm with watershed lines.\label{MeyerAlgorithm} $FIRST(queue)$ extract the highest priority element from the queue.
+$N(p)$ return the list of p's neighbors.
+$MAX(a,b)$ returns the maximum of $a$, and $b$.
+$WSLABEL$ is the label used to define background and watershed pixels (usally $0$).
+The lower completion is done by the statement (\ref{algo:Meyer:max})}
+\end{figure}
 
-           mark as queued
-   \end{itemize}
-\end{itemize}
+% \begin{itemize}
+% \item initialize the queue with background pixels that have 
+%       neighbors that are members of a regional minima. In practice
+%       this is done by checking all regional minima pixels to see
+%       whether they have background neighbors. The background pixels
+%       are placed on the queue and the label of the regional minima
+%       associated with them.
+% \item while (!empty(queue)
+%    \begin{itemize}
+% 	\item Take the pixel $p$ from the head of the queue
+% 	
+%          if $p$ is unlabelled;
+% 
+%            apply the associated label
+% 
+%          else
+%  
+%            check whether $p$ is a watershed pixel. $p$ is a watershed
+%            pixel if there is more than one label in the neighborhood.
+% 
+%         \item for all pixels $q$ that are steepest upper neighbors of $p$; do
+% 	
+% 	if $q$ is not in the queue;
+% 
+%            place in queue, associated with label($p$)
+%         
+%            mark as queued
+%    \end{itemize}
+% \end{itemize}
 

### 2006-04-23 article/Article.tex
add implementation section
darcs-hash:20060423153043-2fc9d-8b6205fc26a93aa83db676431c1013a53b69e72e.gz

@@
 \subsubsection{Algorithm 2}
+\label{sect:BeucherAlgorithm}
 
@@
 \subsubsection{Algorithm 3}
+\label{sect:MeyerAlgorithm}
 
@@
 It is the role of $S$ in Figure \ref{MeyerAlgorithm}. We'll discuss
-in ??? how this status can be implemented.
+in \ref{sect:implementation} how this status can be implemented.
 
@@
 
-\section{Using watersheds in ITK}
+\section{Implementation}
+\label{sect:implementation}
+
+We provide with this article two new classes:
+
+  \subsection{MorphologicalWatershedFromMarkersImageFilter}
+
+\verb$MorphologicalWatershedFromMarkersImageFilter$ implement the
+algorithms \ref{sect:BeucherAlgorithm} and \ref{sect:MeyerAlgorithm}.
+They are accessible respectively when the attribute \verb$MarkWatershed$
+is \verb$false$ or is \verb$true$. The filter is templated over the
+input image type, which can be any scalar type, and the label image type
+which should be an integer. The filter take 2 inputs:
+\begin{itemize}
+  \item a grayscale image which can be set with \verb$SetInput(image)$,
+\verb$SetInput(0, image)$ or \verb$SetInput1(image)$.
+  \item a marker image which can be set with \verb$SetMarkerImage(image)$
+and \verb$SetInput2(image)$. \verb$SetInput(0, image)$ should be avoided
+because it will fail if the input image and the marker image are not of the
+same type. The markers are all the non-zero values in the image.
+\end{itemize}
+
+The algorithm \ref{sect:MeyerAlgorithm} require to keep the status of all
+the pixels: $already added to the queue$ or $not yet added to the queue$.
+This task is often done by using negative values in the output image
+for the pixels not yet added to the queue. In our implementation, we have
+chosen to use a binary image to store this, while labels always take values
+above 0 and watershed lines are marked with value 0. The negative values
+option have several drawbacks:
+\begin{itemize}
+  \item it requires to use a different image type than the label type defined
+by the user in case the user have chosen an unsigned pixel type.
+  \item it requires to use one of the value which can be possibly aready used
+by a label in the marker image, in case of a signed type.
+\end{itemize}
+Those problems can be avoided\footnote{It's not true for the unsigned long
+pixel type} by using a type which contains a larger range of value than the
+type choosed by the user, but then, we must copy the data to an image of
+the type chosen by the user at the end of the algorithm.
+
+We prefered using a separated binary image (of type \verb$itk::Image<bool, dim>$)
+to store this information. This solution avoid the need to copy the image at
+the end, and so should be more efficient in computation time, and require only
+one bit per pixel and so should be more memory efficient as well.
+
+  \subsection{MorphologicalWatershedImageFilter}
+
+\verb$MorphologicalWatershedImageFilter$ perform a watershed transformation
+without marker. It encapsulate several filters:
+\begin{itemize}
+  \item \verb$HMinimaImageFilter$
+  \item \verb$RegionalMinimaImageFilter$
+  \item \verb$ConnectedComponentImageFilter$
+  \item \verb$MorphologicalWatershedFromMarkersImageFilter$
+\end{itemize}
+While \verb$RegionalMinimaImageFilter$, \verb$ConnectedComponentImageFilter$,
+and \verb$MorphologicalWatershedFromMarkersImageFilter$ are the usal steps
+of the waterhed usage, the \verb$HMinimaImageFilter$ is a convenient way
+to avoid the classical watershed's over segmentation by allowing the user
+to select the dynamic of the minima.
 
 \section{Timing and performance}
+
 streaming and threading.
@@
 certain their neighbors are never checked.
+
+\section{Using watersheds in ITK}
 

### 2006-04-23 article/Article.tex
article continuated
darcs-hash:20060423203001-2fc9d-8c629973fe2020663d13cf8f6a114ef38d1395f6.gz

@@
 The algorithm \ref{sect:MeyerAlgorithm} require to keep the status of all
-the pixels: $already added to the queue$ or $not yet added to the queue$.
+the pixels: {\em already added to the queue} or {\em not yet added to the queue}.
 This task is often done by using negative values in the output image
@@
 
+The connectivity is defined using the usual ITK way: \verb$SetFullConnected(false)$
+(4-connected in 2D, 6-connected in 3D) or \verb$SetFullConnected(true)$ (8-connected
+in 2D, 26-connected in 3D)\footnote{One should regret to not have the 18-connected
+connectivity in 3D}.
+
   \subsection{MorphologicalWatershedImageFilter}
@@
 to avoid the classical watershed's over segmentation by allowing the user
-to select the dynamic of the minima.
+to select the dynamic of the minima. The minimum dynamic of minima can be
+selected with \verb$SetThreshold()$.
 
@@
 
+\begin{figure}[htbp]
+\begin{center}
+\subfigure[Input image]{\includegraphics[scale=4]{threshold}}
+\subfigure[No minima threshold]{\includegraphics[scale=4]{threshold00-rgb}}
+\subfigure[Minima threshold = 10]{\includegraphics[scale=4]{threshold10-rgb}}
+\subfigure[Minima threshold = 20]{\includegraphics[scale=4]{threshold20-rgb}}
+\subfigure[Minima threshold = 30 and Minima threshold = 40]{\includegraphics[scale=4]{threshold30-rgb}}
+\subfigure[Minima threshold = 50]{\includegraphics[scale=4]{threshold50-rgb}}
+\subfigure[markers]{\includegraphics[scale=4]{threshold-markers}}
+\subfigure[Watershed from markers]{\includegraphics[scale=4]{threshold-markersMf-rgb}}
+\caption{Different results of watershed on an articial image.
+The labeled image are overlayed to the input image with {\em LabelOverlayImageFilter}.
+From $(b)$ to $(f)$, an example of reduction of the classical waterhsed oversegmentation by minima dynamic filtering.
+$(g)$ and $(h)$ show a marker defined by a user, and the result over the input image.
+All the watersheds were performed with $MarkWatershed = true$ and $FullyConnected = false$.}
+\end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+\begin{center}
+\subfigure[]{\includegraphics[scale=10]{pass-values}}
+\subfigure[]{\includegraphics[scale=10]{pass-values-mf-rgb}}
+\subfigure[]{\includegraphics[scale=10]{pass-values-mF-rgb}}
+\subfigure[]{\includegraphics[scale=10]{pass-values-Mf-rgb}}
+\subfigure[]{\includegraphics[scale=10]{pass-values-MF-rgb}}
+\subfigure[]{\includegraphics[scale=10]{pass-values}}
+
+\subfigure[]{\includegraphics[scale=4]{plateau}}
+\subfigure[]{\includegraphics[scale=4]{plateau-mf-rgb}}
+\subfigure[]{\includegraphics[scale=4]{plateau-mF-rgb}}
+\subfigure[]{\includegraphics[scale=4]{plateau-Mf-rgb}}
+\subfigure[]{\includegraphics[scale=4]{plateau-MF-rgb}}
+\subfigure[]{\includegraphics[scale=4]{plateau}}
+
+\subfigure[]{\includegraphics[scale=10]{thick-lines}}
+\subfigure[]{\includegraphics[scale=10]{thick-lines-mf-rgb}}
+\subfigure[]{\includegraphics[scale=10]{thick-lines-mF-rgb}}
+\subfigure[]{\includegraphics[scale=10]{thick-lines-Mf-rgb}}
+\subfigure[]{\includegraphics[scale=10]{thick-lines-MF-rgb}}
+\subfigure[]{\includegraphics[scale=10]{thick-lines}}
+
+\subfigure[]{\includegraphics[scale=6]{button-hole}}
+\subfigure[]{\includegraphics[scale=6]{button-hole-mf-rgb}}
+\subfigure[]{\includegraphics[scale=6]{button-hole-mF-rgb}}
+\subfigure[]{\includegraphics[scale=6]{button-hole-Mf-rgb}}
+\subfigure[]{\includegraphics[scale=6]{button-hole-MF-rgb}}
+\subfigure[]{\includegraphics[scale=6]{button-hole}}
+
+\caption{}
+\end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+\begin{center}
+\subfigure[]{\includegraphics[scale=0.7]{cthead1}}
+\subfigure[]{\includegraphics[scale=0.7]{cthead1-markers}}
+\subfigure[]{\includegraphics[scale=0.7]{cthead1mf-rgb}}
+\subfigure[]{\includegraphics[scale=0.7]{cthead1mF-rgb}}
+\subfigure[]{\includegraphics[scale=0.7]{cthead1Mf-rgb}}
+\subfigure[]{\includegraphics[scale=0.7]{cthead1MF-rgb}}
+\caption{}
+\end{center}
+\end{figure}
+
 \section{Discussion}

### 2006-04-23 itkLabelOverlayImageFilter.h

@@ -1,6 +1,8 @@
 #ifndef __itkLabelOverlayImageFilter_h
 #define __itkLabelOverlayImageFilter_h
 
+#include "itkLabelToRGBImageFilter.h"
+
 #include "itkBinaryFunctorImageFilter.h"
 
 namespace itk
@@ -12,58 +14,13 @@ template< class TInputPixel, class TLabel, class TRGBPixel >
 class LabelOverlay
 {
 public:
-  LabelOverlay()
-  {
-    
-    TRGBPixel rgbPixel;
-    // the following colors are from "R", and named:
-    // "red"             "green3"          "blue"            "cyan"           
-    //"magenta"         "darkorange1"     "darkgreen"       "blueviolet"     
-    //"brown4"          "navy"            "yellow4"         "violetred1"     
-    //"salmon4"         "turquoise4"      "sienna3"         "darkorchid1"    
-    //"springgreen4"    "mediumvioletred" "orangered3"      "lightseagreen"  
-    //"slateblue"       "deeppink1"       "aquamarine4"     "royalblue1"     
-    //"tomato3"         "mediumblue"      "violetred4"      "darkmagenta"    
-    //"violet"          "red4"           
-    // They are a good selection of distinct colours for plotting and
-    // overlays.
-    
-    addColor( 255, 0, 0 );
-    addColor( 0, 205, 0 );
-    addColor( 0, 0, 255 );
-    addColor( 0, 255, 255 );
-    addColor( 255, 0, 255 );
-    addColor( 255, 127, 0 );
-    addColor( 0, 100, 0 );
-    addColor( 138, 43, 226 );
-    addColor( 139, 35, 35 );
-    addColor( 0, 0, 128 );
-    addColor( 139, 139, 0 );
-    addColor( 255, 62, 150 );
-    addColor( 139, 76, 57 );
-    addColor( 0, 134, 139 );
-    addColor( 205, 104, 57 );
-    addColor( 191, 62, 255 );
-    addColor( 0, 139, 69 );
-    addColor( 199, 21, 133 );
-    addColor( 205, 55, 0 );
-    addColor( 32, 178, 170 );
-    addColor( 106, 90, 205 );
-    addColor( 255, 20, 147 );
-    addColor( 69, 139, 116 );
-    addColor( 72, 118, 255 );
-    addColor( 205, 79, 57 );
-    addColor( 0, 0, 205 );
[…]

### 2006-04-23 itkLabelToRGBImageFilter.h

@@ -1,0 +1,145 @@
+#ifndef __itkLabelToRGBImageFilter_h
+#define __itkLabelToRGBImageFilter_h
+
+#include "itkUnaryFunctorImageFilter.h"
+
+namespace itk
+{
+
+namespace Functor {  
+ 
+template< class TLabel, class TRGBPixel >
+class LabelToRGBFunctor
+{
+public:
+  LabelToRGBFunctor()
+    {
+    
+    TRGBPixel rgbPixel;
+    // the following colors are from "R", and named:
+    // "red"             "green3"          "blue"            "cyan"           
+    //"magenta"         "darkorange1"     "darkgreen"       "blueviolet"     
+    //"brown4"          "navy"            "yellow4"         "violetred1"     
+    //"salmon4"         "turquoise4"      "sienna3"         "darkorchid1"    
+    //"springgreen4"    "mediumvioletred" "orangered3"      "lightseagreen"  
+    //"slateblue"       "deeppink1"       "aquamarine4"     "royalblue1"     
+    //"tomato3"         "mediumblue"      "violetred4"      "darkmagenta"    
+    //"violet"          "red4"           
+    // They are a good selection of distinct colours for plotting and
+    // overlays.
+    
+    addColor( 255, 0, 0 );
+    addColor( 0, 205, 0 );
+    addColor( 0, 0, 255 );
+    addColor( 0, 255, 255 );
+    addColor( 255, 0, 255 );
+    addColor( 255, 127, 0 );
+    addColor( 0, 100, 0 );
+    addColor( 138, 43, 226 );
+    addColor( 139, 35, 35 );
+    addColor( 0, 0, 128 );
+    addColor( 139, 139, 0 );
+    addColor( 255, 62, 150 );
+    addColor( 139, 76, 57 );
+    addColor( 0, 134, 139 );
+    addColor( 205, 104, 57 );
+    addColor( 191, 62, 255 );
+    addColor( 0, 139, 69 );
+    addColor( 199, 21, 133 );
+    addColor( 205, 55, 0 );
+    addColor( 32, 178, 170 );
+    addColor( 106, 90, 205 );
+    addColor( 255, 20, 147 );
+    addColor( 69, 139, 116 );
+    addColor( 72, 118, 255 );
+    addColor( 205, 79, 57 );
+    addColor( 0, 0, 205 );
+    addColor( 139, 34, 82 );
+    addColor( 139, 0, 139 );
+    addColor( 238, 130, 238 );
+    addColor( 139, 0, 0 );
+    }
+
[…]

### 2006-04-23 article/Article.tex
add image from current itk watershed
darcs-hash:20060423230435-2fc9d-1951809248b9364c0ddbf1f99feff5071c99261c.gz

@@
 The labeled image are overlayed to the input image with {\em LabelOverlayImageFilter}.
-From $(b)$ to $(f)$, an example of reduction of the classical waterhsed oversegmentation by minima dynamic filtering.
+From $(b)$ to $(f)$, an example of reduction of the classical watershed over segmentation by minima dynamic filtering.
 $(g)$ and $(h)$ show a marker defined by a user, and the result over the input image.
 All the watersheds were performed with $MarkWatershed = true$ and $FullyConnected = false$.}
+\end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+\begin{center}
+\subfigure[Input image]{\includegraphics[scale=4]{threshold}}
+\subfigure[No minima threshold]{\includegraphics[scale=4]{threshold00-itk-rgb}}
+\subfigure[Minima threshold = 10]{\includegraphics[scale=4]{threshold10-itk-rgb}}
+% \subfigure[Minima threshold = 20]{\includegraphics[scale=4]{threshold20-itk-rgb}}
+\subfigure[Minima threshold = 40]{\includegraphics[scale=4]{threshold40-itk-rgb}}
+\subfigure[Minima threshold = 63]{\includegraphics[scale=4]{threshold63-itk-rgb}}
+\subfigure[Minima threshold = 65]{\includegraphics[scale=4]{threshold65-itk-rgb}}
+\subfigure[markers]{\includegraphics[scale=4]{threshold-markers}}
+\subfigure[Watershed from markers]{\includegraphics[scale=4]{threshold-markers-itk-rgb}}
+\caption{Different results of watershed on an articial image.
+The labeled image are overlayed to the input image with {\em LabelOverlayImageFilter}.
+From $(b)$ to $(f)$, an example of reduction of the classical watershed over segmentation by minima dynamic filtering.
+$(g)$ and $(h)$ show a marker defined by a user, and the result over the input image..}
 \end{center}
@@
 \subfigure[]{\includegraphics[scale=10]{pass-values-MF-rgb}}
-\subfigure[]{\includegraphics[scale=10]{pass-values}}
+\subfigure[]{\includegraphics[scale=10]{pass-values-itk-rgb}}
 
@@
 \subfigure[]{\includegraphics[scale=4]{plateau-MF-rgb}}
-\subfigure[]{\includegraphics[scale=4]{plateau}}
+\subfigure[]{\includegraphics[scale=4]{plateau-itk-rgb}}
 
@@
 \subfigure[]{\includegraphics[scale=10]{thick-lines-MF-rgb}}
-\subfigure[]{\includegraphics[scale=10]{thick-lines}}
+\subfigure[]{\includegraphics[scale=10]{thick-lines-itk-rgb}}
 
@@
 \subfigure[]{\includegraphics[scale=6]{button-hole-MF-rgb}}
-\subfigure[]{\includegraphics[scale=6]{button-hole}}
+\subfigure[]{\includegraphics[scale=6]{button-hole-itk-rgb}}
 
@@
 \begin{center}
-\subfigure[]{\includegraphics[scale=0.7]{cthead1}}
-\subfigure[]{\includegraphics[scale=0.7]{cthead1-markers}}
-\subfigure[]{\includegraphics[scale=0.7]{cthead1mf-rgb}}
-\subfigure[]{\includegraphics[scale=0.7]{cthead1mF-rgb}}
-\subfigure[]{\includegraphics[scale=0.7]{cthead1Mf-rgb}}
-\subfigure[]{\includegraphics[scale=0.7]{cthead1MF-rgb}}
+\subfigure[]{\includegraphics[scale=0.5]{cthead1}}
+\subfigure[]{\includegraphics[scale=0.5]{cthead1-markers}}
+\subfigure[]{\includegraphics[scale=0.5]{cthead1mf-rgb}}
+\subfigure[]{\includegraphics[scale=0.5]{cthead1mF-rgb}}
+\subfigure[]{\includegraphics[scale=0.5]{cthead1Mf-rgb}}
+\subfigure[]{\includegraphics[scale=0.5]{cthead1MF-rgb}}
+\subfigure[]{\includegraphics[scale=0.5]{cthead1itk-rgb}}
 \caption{}

### 2006-04-23 itkMinimaImpositionImageFilter.h
add minima imposition filters
darcs-hash:20060423230510-2fc9d-8377849c37332a9dcd7a1f35026aaae12d4141ce.gz

@@ -1,0 +1,144 @@
+/*=========================================================================
+
+  Program:   Insight Segmentation & Registration Toolkit
+  Module:    $RCSfile: itkMinimaImpositionImageFilter.h,v $
+  Language:  C++
+  Date:      $Date: 2005/08/23 15:09:03 $
+  Version:   $Revision: 1.4 $
+
+  Copyright (c) Insight Software Consortium. All rights reserved.
+  See ITKCopyright.txt or http://www.itk.org/HTML/Copyright.htm for details.
+
+     This software is distributed WITHOUT ANY WARRANTY; without even 
+     the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR 
+     PURPOSE.  See the above copyright notices for more information.
+
+=========================================================================*/
+#ifndef __itkMinimaImpositionImageFilter_h
+#define __itkMinimaImpositionImageFilter_h
+
+#include "itkImageToImageFilter.h"
+
+namespace itk {
+
+/** \class MinimaImpositionImageFilter
+ * \brief Impose the minima defined in a marker image to the input image
+ *
+ * TLabelImage should be an integer type. All non zero pixels of the image are
+ * considered as markers.
+ *
+ * The minima imposition algorithm is described in
+ * Chapter 6.3.6 of Pierre Soille's book "Morphological Image Analysis:
+ * Principles and Applications", Second Edition, Springer, 2003.
+ *
+ * \author Gaëtan Lehmann. Biologie du Développement et de la Reproduction, INRA de Jouy-en-Josas, France.
+ *
+ * \sa WatershedImageFilter
+ * \ingroup ImageEnhancement  MathematicalMorphologyImageFilters
+ */
+template<class TInputImage, class TLabelImage>
+class ITK_EXPORT MinimaImpositionImageFilter : 
+    public ImageToImageFilter<TInputImage, TLabelImage>
+{
+public:
+  /** Standard class typedefs. */
+  typedef MinimaImpositionImageFilter Self;
+  typedef ImageToImageFilter<TInputImage, TLabelImage>
+  Superclass;
+  typedef SmartPointer<Self>        Pointer;
+  typedef SmartPointer<const Self>  ConstPointer;
+
+  /** Some convenient typedefs. */
+  typedef TInputImage InputImageType;
+  typedef TLabelImage LabelImageType;
+  typedef typename InputImageType::Pointer         InputImagePointer;
[…]

### 2006-04-23 article/Article.tex
Fixed conflicts in article
darcs-hash:20060423232523-fafb9-9348987c8c297faf893ae4c66478b9ffc4695056.gz

@@
 
-\subsubsection{Algorithm 2}
+\subsubsection{Algorithm 2 - commonly attributed to Beucher.}
 \label{sect:BeucherAlgorithm}
@@
 
-\subsubsection{Algorithm 3}
+\subsubsection{Algorithm 3 - commonly attributed to Meyer.}
 \label{sect:MeyerAlgorithm}
@@
 \section{Timing and performance}
+The relative performance of the new filters are shown in
+Table~\ref{tbl:perf}. The timings were obtained using the {\tt
+./perf3D images/ESCells.img}. The times for steps involved in a
+morphological watershed are shown in the first few columns, with the
+time from a combined filter and the itk watershed shown last. It
+should be noted that the itk watershed does not allow connectivity to
+be modified by the user - it always uses face connected mode and
+doesn't have an inbuilt mechanism for boundary selection. The existing
+ITK watershed is therefore somewhat slower than the new implementation
+when used to compute a traditional watershed. The new implementations
+are not easily threadable and therefore may prove to be slower on some
+systems. However, as discussed in earlier sections, the new
+implementation produces more stable results than the rainfall model
+and can be used in a marker based approach.
 
-streaming and threading.
+\begin{table}[htbp]
+\centering
+\begin{tabular}{cccccccc}
+\hline
+FC & MB & Find Min &  Label mins & Watershed from markers & Total & Morph watershed &  itk watershed \\
+\hline
+0  &    0  &     1.948  & 0.522  & 5.752  & 8.222   & 8.218  & 12.596 \\
+0  &    1  &     1.922  & 0.516  & 6.724  & 9.162   & 9.28   & 12.61 \\
+1  &    0  &     4.26   & 0.154  & 13.274 & 17.688  & 18.158 & 12.61 \\
+1  &    1  &     4.284  & 0.15   & 17.23  & 21.664  & 21.76  & 12.606 \\
+\hline
+\end{tabular}
+\caption{Execution times. Note the FC and MB refer to FullyConnected and MarkBoundary.\label{tbl:perf}}
+\end{table}
 
-Preventing boundary checks - not really the itk way, but if a user is
+\subsection{Other performance issues/comments}
+\begin{itemize}
+\item Preventing boundary checks - not really the itk way, but if a user is
 willing to throw away the image border by treating it as a watershed
@@
 certain their neighbors are never checked.
+\item Streaming and threading  - queue based algorithms are generally difficult 
+to thread efficiently and generally need to have all of the input
+available so prevent streaming from working well. The existing ITK
+watershed does provide a multithreaded implementation and similar
+techniques may be possible with the new implementation. Other
+possibilities include using threads to improve performance of queues.
+\end{itemize}
 

### 2006-04-24 article/Article.tex
change the size of some image
darcs-hash:20060424071352-2fc9d-a080b596340c797115623aa946abc365d8b48028.gz

@@
 \begin{center}
-\subfigure[Input image]{\includegraphics[scale=4]{threshold}}
-\subfigure[No minima threshold]{\includegraphics[scale=4]{threshold00-rgb}}
-\subfigure[Minima threshold = 10]{\includegraphics[scale=4]{threshold10-rgb}}
-\subfigure[Minima threshold = 20]{\includegraphics[scale=4]{threshold20-rgb}}
-\subfigure[Minima threshold = 30 and Minima threshold = 40]{\includegraphics[scale=4]{threshold30-rgb}}
-\subfigure[Minima threshold = 50]{\includegraphics[scale=4]{threshold50-rgb}}
-\subfigure[markers]{\includegraphics[scale=4]{threshold-markers}}
-\subfigure[Watershed from markers]{\includegraphics[scale=4]{threshold-markersMf-rgb}}
+\subfigure[Input image]{\includegraphics[scale=3.2]{threshold}}
+\subfigure[No minima threshold]{\includegraphics[scale=3.2]{threshold00-rgb}}
+\subfigure[Minima threshold = 10]{\includegraphics[scale=3.2]{threshold10-rgb}}
+\subfigure[Minima threshold = 20]{\includegraphics[scale=3.2]{threshold20-rgb}}
+\subfigure[Minima threshold = 30 and Minima threshold = 40]{\includegraphics[scale=3.2]{threshold30-rgb}}
+\subfigure[Minima threshold = 50]{\includegraphics[scale=3.2]{threshold50-rgb}}
+\subfigure[markers]{\includegraphics[scale=3.2]{threshold-markers}}
+\subfigure[Watershed from markers]{\includegraphics[scale=3.2]{threshold-markersMf-rgb}}
 \caption{Different results of watershed on an articial image.
@@
 \begin{center}
-\subfigure[Input image]{\includegraphics[scale=4]{threshold}}
-\subfigure[No minima threshold]{\includegraphics[scale=4]{threshold00-itk-rgb}}
-\subfigure[Minima threshold = 10]{\includegraphics[scale=4]{threshold10-itk-rgb}}
-% \subfigure[Minima threshold = 20]{\includegraphics[scale=4]{threshold20-itk-rgb}}
-\subfigure[Minima threshold = 40]{\includegraphics[scale=4]{threshold40-itk-rgb}}
-\subfigure[Minima threshold = 63]{\includegraphics[scale=4]{threshold63-itk-rgb}}
-\subfigure[Minima threshold = 65]{\includegraphics[scale=4]{threshold65-itk-rgb}}
-\subfigure[markers]{\includegraphics[scale=4]{threshold-markers}}
-\subfigure[Watershed from markers]{\includegraphics[scale=4]{threshold-markers-itk-rgb}}
+\subfigure[Input image]{\includegraphics[scale=3.2]{threshold}}
+\subfigure[No minima threshold]{\includegraphics[scale=3.2]{threshold00-itk-rgb}}
+\subfigure[Level = 0.1]{\includegraphics[scale=3.2]{threshold10-itk-rgb}}
+% \subfigure[Level = 0.2]{\includegraphics[scale=3.2]{threshold20-itk-rgb}}
+\subfigure[Level = 0.4]{\includegraphics[scale=3.2]{threshold40-itk-rgb}}
+\subfigure[Level = 0.63]{\includegraphics[scale=3.2]{threshold63-itk-rgb}}
+\subfigure[Level = 0.65]{\includegraphics[scale=3.2]{threshold65-itk-rgb}}
+\subfigure[markers]{\includegraphics[scale=3.2]{threshold-markers}}
+\subfigure[Watershed from markers]{\includegraphics[scale=3.2]{threshold-markers-itk-rgb}}
 \caption{Different results of watershed on an articial image.

### 2006-04-24 article/Article.tex
article continued, with better images
darcs-hash:20060424144440-2fc9d-5d280e7e505eba5058ee9638e57ee2f1bde7558a.gz

@@
 integration} and will be discussed in more detail in this report.
+Figure \ref{fig:cthead1-markers} and \ref{fig:difficult-cases} illustrate
+the possible problem with plateaus.
+
+\begin{figure}[htbp]
+\begin{center}
+\subfigure[Input image]{\includegraphics[scale=0.7]{cthead1}}
+\subfigure[User defined labels colored with LabelOverlayImageFilter]{\includegraphics[scale=0.7]{cthead1-markers-rgb}}
+% \subfigure[]{\includegraphics[scale=0.7]{cthead1mf-rgb}}
+% \subfigure[]{\includegraphics[scale=0.7]{cthead1mF-rgb}}
+% \subfigure[]{\includegraphics[scale=0.7]{cthead1Mf-rgb}}
+\subfigure[Morphological watershed]{\includegraphics[scale=0.7]{cthead1MF-rgb}}
+\subfigure[ITK watershed]{\includegraphics[scale=0.7]{cthead1itk-rgb}}
+\caption{Example of plateau problem. The blue marker is clearly extended where it should not with the ITK watershed.\label{fig:cthead1-markers}}
+\end{center}
+\end{figure}
 
@@
 while many watershed lines do lie on significant boundaries, there are
-many that don't. Figure \ref{fig:overseg} illustrates the problem. The ITK
+many that don't. Figure \ref{fig:overseg} illustrates the problem.
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.75]{cthead1-over-segmentation-rgb}
+\caption{Classical over segmentation.}
+\label{fig:overseg}
+\end{center}
+\end{figure}
+
+The ITK
 watershed implementation provides a way of addressing this problem
@@
 
+\begin{figure}[htbp]
+\begin{center}
+\subfigure[Input image]{\includegraphics[scale=3.2]{threshold}}
+\subfigure[No minima threshold]{\includegraphics[scale=3.2]{threshold00-rgb}}
+\subfigure[Minima threshold = 10]{\includegraphics[scale=3.2]{threshold10-rgb}}
+\subfigure[Minima threshold = 20]{\includegraphics[scale=3.2]{threshold20-rgb}}
+\subfigure[Minima threshold = 30 and Minima threshold = 40]{\includegraphics[scale=3.2]{threshold30-rgb}}
+\subfigure[Minima threshold = 50]{\includegraphics[scale=3.2]{threshold50-rgb}}
+\subfigure[markers]{\includegraphics[scale=3.2]{threshold-markers-rgb}}
+\subfigure[Watershed from markers]{\includegraphics[scale=3.2]{threshold-markersMf-rgb}}
+\caption{Different results of morphological watershed on an articial image.
+The labeled image are overlayed to the input image with {\em LabelOverlayImageFilter}.
+From $(b)$ to $(f)$, an example of reduction of the classical watershed over segmentation by minima dynamic filtering.
+$(g)$ and $(h)$ show a marker defined by a user, and the result over the input image.
+All the watersheds were performed with $MarkWatershed = true$ and $FullyConnected = false$.}
+\end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+\begin{center}
+\subfigure[Input image]{\includegraphics[scale=3.2]{threshold}}
+\subfigure[No minima threshold]{\includegraphics[scale=3.2]{threshold00-itk-rgb}}
+\subfigure[Level = 0.1]{\includegraphics[scale=3.2]{threshold10-itk-rgb}}
+% \subfigure[Level = 0.2]{\includegraphics[scale=3.2]{threshold20-itk-rgb}}
+\subfigure[Level = 0.4]{\includegraphics[scale=3.2]{threshold40-itk-rgb}}
+\subfigure[Level = 0.63]{\includegraphics[scale=3.2]{threshold63-itk-rgb}}
+\subfigure[Level = 0.65]{\includegraphics[scale=3.2]{threshold65-itk-rgb}}
+\subfigure[markers]{\includegraphics[scale=3.2]{threshold-markers-rgb}}
+\subfigure[Watershed from markers]{\includegraphics[scale=3.2]{threshold-markers-itk-rgb}}
+\caption{Different results of ITK's watershed on an articial image.
+The labeled image are overlayed to the input image with {\em LabelOverlayImageFilter}.
+From $(b)$ to $(f)$, an example of reduction of the classical watershed over segmentation by minima dynamic filtering.
+$(g)$ and $(h)$ show a marker defined by a user, and the result over the input image.}
+\end{center}
+\end{figure}
+
 An Obvious alternative that is very similar in practice is to 
 smooth the image so that there are fewer regional minima.
-\begin{figure}
-\begin{center}
-\caption{Classical over segmentation.}
-\label{fig:overseg}
-\end{center}
-\end{figure}
 
@@
 
+\begin{figure}[htbp]
+\begin{center}
+% \subfigure[]{\includegraphics[scale=10]{pass-values}}
+% \subfigure[]{\includegraphics[scale=10]{pass-values-mf-rgb}}
+% \subfigure[]{\includegraphics[scale=10]{pass-values-mF-rgb}}
+% \subfigure[]{\includegraphics[scale=10]{pass-values-Mf-rgb}}
+% \subfigure[]{\includegraphics[scale=10]{pass-values-MF-rgb}}
+% \subfigure[]{\includegraphics[scale=10]{pass-values-itk-rgb}}
+
+\subfigure[Plateau]{\includegraphics[scale=4]{plateau}}
+\subfigure[$F=false$ and $M=false$]{\includegraphics[scale=4]{plateau-mf-rgb}}
+\subfigure[$F=true$ and $M=false$]{\includegraphics[scale=4]{plateau-mF-rgb}}
+\subfigure[$F=false$ and $M=true$]{\includegraphics[scale=4]{plateau-Mf-rgb}}
+\subfigure[$F=true$ and $M=true$]{\includegraphics[scale=4]{plateau-MF-rgb}}
+\subfigure[ITK]{\includegraphics[scale=4]{plateau-itk-rgb}}
+
+\subfigure[Thick line]{\includegraphics[scale=10]{thick-lines}}
+\subfigure[$F=false$ and $M=false$]{\includegraphics[scale=10]{thick-lines-mf-rgb}}
+\subfigure[$F=true$ and $M=false$]{\includegraphics[scale=10]{thick-lines-mF-rgb}}
+\subfigure[$F=false$ and $M=true$]{\includegraphics[scale=10]{thick-lines-Mf-rgb}}
+\subfigure[$F=true$ and $M=true$]{\includegraphics[scale=10]{thick-lines-MF-rgb}}
+\subfigure[ITK]{\includegraphics[scale=10]{thick-lines-itk-rgb}}
+
+\subfigure[Button hole]{\includegraphics[scale=8]{button-hole}}
+\subfigure[$F=false$ and $M=false$]{\includegraphics[scale=8]{button-hole-mf-rgb}}
+\subfigure[$F=true$ and $M=false$]{\includegraphics[scale=8]{button-hole-mF-rgb}}
+\subfigure[$F=false$ and $M=true$]{\includegraphics[scale=8]{button-hole-Mf-rgb}}
+\subfigure[$F=true$ and $M=true$]{\includegraphics[scale=8]{button-hole-MF-rgb}}
+\subfigure[ITK]{\includegraphics[scale=8]{button-hole-itk-rgb}}
+
+\caption{Three difficult test images.
+The first image, plateau, show that ITK's watershed produce a bad result on this image: only one minima take all the image.
+This problem is also shown in Figure \ref{fig:cthead1-markers}.
+The morphological watershed give a better result, but not yet perfect: the separation line should not be curved as is it.
+The next image, thick line, again show that ITK watershed can produce strange results with a green label which is where we should have the light blue one.
+Morphological watershed gives good results.
+The last image, button hole, is a very difficult case for watershed algorithms which mark watershed line: the watershed pixel at the bottom of the image stop the label propagation.
+There is several possible behavior: produce a large watershed line which cover all the zone (Vincent-Soille algorithm, not shown here), let the label on the top of the image propagate in the zone (Meyer algorithm), split the zone in 2 zones with the 2 labels of the bottom of the image (topological watershed, not shown here).\label{fig:difficult-cases}}
+\end{center}
+\end{figure}
+
 \section{Finding regional minima}
@@
 
-\begin{figure}[htbp]
-\begin{center}
-\subfigure[Input image]{\includegraphics[scale=3.2]{threshold}}
-\subfigure[No minima threshold]{\includegraphics[scale=3.2]{threshold00-rgb}}
-\subfigure[Minima threshold = 10]{\includegraphics[scale=3.2]{threshold10-rgb}}
-\subfigure[Minima threshold = 20]{\includegraphics[scale=3.2]{threshold20-rgb}}
-\subfigure[Minima threshold = 30 and Minima threshold = 40]{\includegraphics[scale=3.2]{threshold30-rgb}}
-\subfigure[Minima threshold = 50]{\includegraphics[scale=3.2]{threshold50-rgb}}
-\subfigure[markers]{\includegraphics[scale=3.2]{threshold-markers}}
-\subfigure[Watershed from markers]{\includegraphics[scale=3.2]{threshold-markersMf-rgb}}
-\caption{Different results of watershed on an articial image.
-The labeled image are overlayed to the input image with {\em LabelOverlayImageFilter}.
-From $(b)$ to $(f)$, an example of reduction of the classical watershed over segmentation by minima dynamic filtering.
-$(g)$ and $(h)$ show a marker defined by a user, and the result over the input image.
-All the watersheds were performed with $MarkWatershed = true$ and $FullyConnected = false$.}
-\end{center}
-\end{figure}
+\subsection{Pronucleus segmentation in bovine embryo}
 
-\begin{figure}[htbp]
-\begin{center}
-\subfigure[Input image]{\includegraphics[scale=3.2]{threshold}}
-\subfigure[No minima threshold]{\includegraphics[scale=3.2]{threshold00-itk-rgb}}
-\subfigure[Level = 0.1]{\includegraphics[scale=3.2]{threshold10-itk-rgb}}
-% \subfigure[Level = 0.2]{\includegraphics[scale=3.2]{threshold20-itk-rgb}}
-\subfigure[Level = 0.4]{\includegraphics[scale=3.2]{threshold40-itk-rgb}}
-\subfigure[Level = 0.63]{\includegraphics[scale=3.2]{threshold63-itk-rgb}}
-\subfigure[Level = 0.65]{\includegraphics[scale=3.2]{threshold65-itk-rgb}}
-\subfigure[markers]{\includegraphics[scale=3.2]{threshold-markers}}
-\subfigure[Watershed from markers]{\includegraphics[scale=3.2]{threshold-markers-itk-rgb}}
-\caption{Different results of watershed on an articial image.
-The labeled image are overlayed to the input image with {\em LabelOverlayImageFilter}.
-From $(b)$ to $(f)$, an example of reduction of the classical watershed over segmentation by minima dynamic filtering.
-$(g)$ and $(h)$ show a marker defined by a user, and the result over the input image..}
-\end{center}
-\end{figure}
-
-\begin{figure}[htbp]
-\begin{center}
-\subfigure[]{\includegraphics[scale=10]{pass-values}}
-\subfigure[]{\includegraphics[scale=10]{pass-values-mf-rgb}}
-\subfigure[]{\includegraphics[scale=10]{pass-values-mF-rgb}}
-\subfigure[]{\includegraphics[scale=10]{pass-values-Mf-rgb}}
-\subfigure[]{\includegraphics[scale=10]{pass-values-MF-rgb}}
-\subfigure[]{\includegraphics[scale=10]{pass-values-itk-rgb}}
-
-\subfigure[]{\includegraphics[scale=4]{plateau}}
-\subfigure[]{\includegraphics[scale=4]{plateau-mf-rgb}}
-\subfigure[]{\includegraphics[scale=4]{plateau-mF-rgb}}
-\subfigure[]{\includegraphics[scale=4]{plateau-Mf-rgb}}
-\subfigure[]{\includegraphics[scale=4]{plateau-MF-rgb}}
-\subfigure[]{\includegraphics[scale=4]{plateau-itk-rgb}}
-
-\subfigure[]{\includegraphics[scale=10]{thick-lines}}
-\subfigure[]{\includegraphics[scale=10]{thick-lines-mf-rgb}}
-\subfigure[]{\includegraphics[scale=10]{thick-lines-mF-rgb}}
-\subfigure[]{\includegraphics[scale=10]{thick-lines-Mf-rgb}}
-\subfigure[]{\includegraphics[scale=10]{thick-lines-MF-rgb}}
-\subfigure[]{\includegraphics[scale=10]{thick-lines-itk-rgb}}
-
-\subfigure[]{\includegraphics[scale=6]{button-hole}}
-\subfigure[]{\includegraphics[scale=6]{button-hole-mf-rgb}}
-\subfigure[]{\includegraphics[scale=6]{button-hole-mF-rgb}}
-\subfigure[]{\includegraphics[scale=6]{button-hole-Mf-rgb}}
-\subfigure[]{\includegraphics[scale=6]{button-hole-MF-rgb}}
-\subfigure[]{\includegraphics[scale=6]{button-hole-itk-rgb}}
-
-\caption{}
-\end{center}
-\end{figure}
-
-\begin{figure}[htbp]
-\begin{center}
-\subfigure[]{\includegraphics[scale=0.5]{cthead1}}
-\subfigure[]{\includegraphics[scale=0.5]{cthead1-markers}}
-\subfigure[]{\includegraphics[scale=0.5]{cthead1mf-rgb}}
-\subfigure[]{\includegraphics[scale=0.5]{cthead1mF-rgb}}
-\subfigure[]{\includegraphics[scale=0.5]{cthead1Mf-rgb}}
-\subfigure[]{\includegraphics[scale=0.5]{cthead1MF-rgb}}
-\subfigure[]{\includegraphics[scale=0.5]{cthead1itk-rgb}}
-\caption{}
-\end{center}
-\end{figure}
+\subsection{???}
 

### 2006-04-25 article/Article.tex
pronucleus segmentation in bovine embryo
darcs-hash:20060425150425-2fc9d-474338ada516754e8a5ad079e60e8367a9501286.gz

@@
 This problem is also shown in Figure \ref{fig:cthead1-markers}.
-The morphological watershed give a better result, but not yet perfect: the separation line should not be curved as is it.
+The morphological watershed give a better result, but not yet perfect: the separation line should not be curved as it is.
+This problem is also visible in Figure \ref{fig:embryo-mws}.
 The next image, thick line, again show that ITK watershed can produce strange results with a green label which is where we should have the light blue one.
@@
 
+In this example, we use the watershed filter to split two nearly round
+objects: the male and female pronucleus in a bovine embryo, around 18
+hours after fecondation.
+
+The program is in python for conciseness\footnote{And also to promote
+python wrappers.}.
+
+\begin{figure}[htbp]
+\begin{center}
+\subfigure[Input]{\includegraphics[scale=0.22]{embryo}}
+\subfigure[median]{\includegraphics[scale=0.22]{embryo-median}}
+\subfigure[threshold]{\includegraphics[scale=0.22]{embryo-th}}
+\subfigure[opening by reconstruction]{\includegraphics[scale=0.22]{embryo-open}}
+\subfigure[fill hole]{\includegraphics[scale=0.22]{embryo-fill}}
+\subfigure[invert]{\includegraphics[scale=0.22]{embryo-invert}}
+\subfigure[distance]{\includegraphics[scale=0.22]{embryo-distance}}
+\subfigure[inverted distance]{\includegraphics[scale=0.22]{embryo-idistance}}
+\subfigure[morphological watershed\label{fig:embryo-mws}]{\includegraphics[scale=0.45]{embryo-overlay}}
+\subfigure[ITK watershed]{\includegraphics[scale=0.45]{embryo-overlay2}}
+
+\caption{The results of the different steps of the segmentation of bovine pronucleus.}
+\end{center}
+\end{figure}
+
+\small \begin{verbatim}
+import itk
+reader = itk.ImageFileReader.IUC2.New(FileName='embryo.png')
+\end{verbatim} \normalsize
+
+We use a median image filter to remove some noise, and threshold the image
+to get a binary image\footnote{The methods to remove the noise and binarize the
+image are volontary kept simple for this example}.
+
+\small \begin{verbatim}
+median = itk.MedianImageFilter.IUC2IUC2.New(reader)
+th = itk.BinaryThresholdImageFilter.IUC2IUC2.New(median, LowerThreshold=40)
+\end{verbatim} \normalsize
+
+The binary image produced by the \verb$ThresholdImageFilter$ contains lot
+of small objects. To remove them, we use an {\em opening by reconstruction}
+which have the advantage to preserve the egdes of the object.
+
+\small \begin{verbatim}
+kernel = itk.strel(2, 30)
+open = itk.OpeningByReconstructionImageFilter.IUC2IUC2.New(th, Kernel=kernel)
+\end{verbatim} \normalsize
+
+There is now only one binary object in the image, but it contains lots of holes
+which will produce an unsable distance map for the watershed. We fill them
+with a {\em hole filling} filter. This algorithm, based on a morphological
+reconstruction, fill the hole in a binary object but preserve the edges of the
+object. In this case, it is important to keep the two cavities between the two
+pronucleus unchanged.
+
+\small \begin{verbatim}
+fill = itk.GrayscaleFillholeImageFilter.IUC2IUC2.New(open)
+\end{verbatim} \normalsize
+
+Then we compute the distance map which will be used by the watershed. The
+binary object must be inverted to compute the distance {\em inside} the
+object, and the distance map must be inverted to used by the watershed.
+
+\small \begin{verbatim}
+invert = itk.InvertIntensityImageFilter.IUC2IUC2.New(fill)
+distance = itk.DanielssonDistanceMapImageFilter.IUC2IUC2.New(invert)
+idistance = itk.InvertIntensityImageFilter.IUC2IUC2.New(distance)
+\end{verbatim} \normalsize
+
+The watershed is then performed to split the two pronucleus. A threshold
+is used to avoid the too small minima in the border of the object and thus
+avoid over segmentation. The label produced by the watershed are then
+displayed on top of the original image.
+
+The watershed extends the labels on the entire image. We use a
+\verb$MaskImageFilter$ to limit the labels to the object. Another commonly
+used method is to use the background as a marker and to perform a watershed
+from markers.
+
+\small \begin{verbatim}
+ws = itk.MorphologicalWatershedImageFilter.IUC2IUC2.New(idistance,
+                                                        MarkWatershed=False,
+                                                        Threshold=10)
+mask = itk.MaskImageFilter.IUC2IUC2IUC2.New(ws, fill)
+overlay = itk.LabelOverlayImageFilter.IUC2IUC2IRGBUC2.New(reader, mask, Opacity=0.5)
+\end{verbatim} \normalsize
+
+The same watershed is done with the ITK watershed filter\footnote{A
+$RelabelComponentImageFilter$ is used here because unsigned long is
+not available in python for the $MaskImageFilter$}.
+
+\small \begin{verbatim}
+ws2 = itk.WatershedImageFilter.IUC2.New(idistance, Level=0.5)
+relabel = itk.RelabelComponentImageFilter.IUL2IUC2.New(ws2)
+mask2 = itk.MaskImageFilter.IUC2IUC2IUC2.New(relabel, fill)
+overlay2 = itk.LabelOverlayImageFilter.IUC2IUC2IRGBUC2.New(reader, mask2, Opacity=0.5)
+\end{verbatim} \normalsize
+
+ITK watershed clearly produce a wrong segmentation in that case.
+The result of the morphological watershed is quite good, even if not perfect:
+there is some curves in the separation line of the two labels. This is the
+result of the distance approximation in the watershed algorithm.
+
 \subsection{???}

### 2006-04-27 article/Article.tex
acknowledgments
darcs-hash:20060425150623-2fc9d-527fc267aabb5a8efedd6ab3fea99d8237f05f89.gz
rename Threshold attribute to Level to be consistent with ITK notation
darcs-hash:20060427092832-2fc9d-0eb7d917074f6ef3a26863a9dd791e8538044a84.gz

@@
 \begin{center}
-\subfigure[Input image]{\includegraphics[scale=3.2]{threshold}}
-\subfigure[No minima threshold]{\includegraphics[scale=3.2]{threshold00-rgb}}
-\subfigure[Minima threshold = 10]{\includegraphics[scale=3.2]{threshold10-rgb}}
-\subfigure[Minima threshold = 20]{\includegraphics[scale=3.2]{threshold20-rgb}}
-\subfigure[Minima threshold = 30 and Minima threshold = 40]{\includegraphics[scale=3.2]{threshold30-rgb}}
-\subfigure[Minima threshold = 50]{\includegraphics[scale=3.2]{threshold50-rgb}}
-\subfigure[markers]{\includegraphics[scale=3.2]{threshold-markers-rgb}}
-\subfigure[Watershed from markers]{\includegraphics[scale=3.2]{threshold-markersMf-rgb}}
+\subfigure[Input image]{\includegraphics[scale=3.2]{level}}
+\subfigure[No minima level]{\includegraphics[scale=3.2]{level00-rgb}}
+\subfigure[Minima level = 10]{\includegraphics[scale=3.2]{level10-rgb}}
+\subfigure[Minima level = 20]{\includegraphics[scale=3.2]{level20-rgb}}
+\subfigure[Minima level = 30 and Minima level = 40]{\includegraphics[scale=3.2]{level30-rgb}}
+\subfigure[Minima level = 50]{\includegraphics[scale=3.2]{level50-rgb}}
+\subfigure[markers]{\includegraphics[scale=3.2]{level-markers-rgb}}
+\subfigure[Watershed from markers]{\includegraphics[scale=3.2]{level-markersMf-rgb}}
 \caption{Different results of morphological watershed on an articial image.
@@
 \begin{center}
-\subfigure[Input image]{\includegraphics[scale=3.2]{threshold}}
-\subfigure[No minima threshold]{\includegraphics[scale=3.2]{threshold00-itk-rgb}}
-\subfigure[Level = 0.1]{\includegraphics[scale=3.2]{threshold10-itk-rgb}}
-% \subfigure[Level = 0.2]{\includegraphics[scale=3.2]{threshold20-itk-rgb}}
-\subfigure[Level = 0.4]{\includegraphics[scale=3.2]{threshold40-itk-rgb}}
-\subfigure[Level = 0.63]{\includegraphics[scale=3.2]{threshold63-itk-rgb}}
-\subfigure[Level = 0.65]{\includegraphics[scale=3.2]{threshold65-itk-rgb}}
-\subfigure[markers]{\includegraphics[scale=3.2]{threshold-markers-rgb}}
-\subfigure[Watershed from markers]{\includegraphics[scale=3.2]{threshold-markers-itk-rgb}}
+\subfigure[Input image]{\includegraphics[scale=3.2]{level}}
+\subfigure[No minima level]{\includegraphics[scale=3.2]{level00-itk-rgb}}
+\subfigure[Level = 0.1]{\includegraphics[scale=3.2]{level10-itk-rgb}}
+% \subfigure[Level = 0.2]{\includegraphics[scale=3.2]{level20-itk-rgb}}
+\subfigure[Level = 0.4]{\includegraphics[scale=3.2]{level40-itk-rgb}}
+\subfigure[Level = 0.63]{\includegraphics[scale=3.2]{level63-itk-rgb}}
+\subfigure[Level = 0.65]{\includegraphics[scale=3.2]{level65-itk-rgb}}
+\subfigure[markers]{\includegraphics[scale=3.2]{level-markers-rgb}}
+\subfigure[Watershed from markers]{\includegraphics[scale=3.2]{level-markers-itk-rgb}}
 \caption{Different results of ITK's watershed on an articial image.
@@
 to select the dynamic of the minima. The minimum dynamic of minima can be
-selected with \verb$SetThreshold()$.
+selected with \verb$SetLevel()$.
 
@@
 In this example, we use the watershed filter to split two nearly round
-objects: the male and female pronucleus in a bovine embryo, around 18
+objects: the male and female pronucleus in a bovine embryo, around 22
 hours after fecondation.
@@
 
-The watershed is then performed to split the two pronucleus. A threshold
+The watershed is then performed to split the two pronucleus. A level
 is used to avoid the too small minima in the border of the object and thus
@@
                                                         MarkWatershed=False,
-                                                        Threshold=10)
+                                                        Level=10)
 mask = itk.MaskImageFilter.IUC2IUC2IUC2.New(ws, fill)
@@
 
+\section{Acknowledgments}
+We thank Dr Pierre Adenot and MIMA2 confocal facilities
+(\url{http://mima2.jouy.inra.fr}) for providing ESCells image used for
+timing and for the bovine embryo image.
+
 \appendix

### 2006-04-27 itkMorphologicalWatershedImageFilter.h

@@ -102,8 +102,8 @@ public:
 
   /**
    */
-  itkSetMacro(Threshold, InputImagePixelType);
-  itkGetMacro(Threshold, InputImagePixelType);
+  itkSetMacro(Level, InputImagePixelType);
+  itkGetMacro(Level, InputImagePixelType);
 
 protected:
   MorphologicalWatershedImageFilter();
@@ -131,7 +131,7 @@ private:
 
   bool m_MarkWatershed;
 
-  InputImagePixelType m_Threshold;
+  InputImagePixelType m_Level;
 
 } ; // end of class
 

### 2006-04-27 itkLabelOverlayImageFilter.h
update label overlay
darcs-hash:20060427101804-2fc9d-f99a2f1e9cfeb6254802060a301d7e44614a3381.gz

@@ -15,11 +15,15 @@ class LabelOverlay
 {
 public:
 
-  LabelOverlay() {}
+  LabelOverlay() 
+    {
+    m_UseBackground = false;
+    m_BackgroundValue = NumericTraits<TLabel>::Zero;
+    }
 
   inline TRGBPixel operator()(  const TInputPixel & p1,
 				const TLabel & p2)
-  {
+    {
     if( m_UseBackground && p2 == m_BackgroundValue )
       {
       typename TRGBPixel::ValueType p = static_cast< typename TRGBPixel::ValueType >( p1 );
@@ -33,11 +37,12 @@ public:
        {
        rgbPixel[i] = static_cast< typename TRGBPixel::ValueType >( opaque[i] * m_Opacity + p1 * ( 1.0 - m_Opacity ) );
        }
+std::cout << opaque << std::endl;
      return rgbPixel;
-  }
+    }
 
   bool operator != (const LabelOverlay &l) const
-  { return l.m_Opacity == m_Opacity; }
+  { return l.m_Opacity == m_Opacity || m_UseBackground != l.m_UseBackground || m_BackgroundValue != l.m_BackgroundValue; }
 
   ~LabelOverlay() {}
 
@@ -66,9 +71,8 @@ public:
  * can be defined by the user.
  *
  * \author Gaëtan Lehmann. Biologie du Développement et de la Reproduction, INRA de Jouy-en-Josas, France.
- * \author Richard Beare. Department of Medicine, Monash University, Melbourne, Australia.
  *
- * \sa ScalarToRGBPixelFunctor
+ * \sa ScalarToRGBPixelFunctor LabelToRGBImageFilter
  * \ingroup Multithreaded
  *
  */

### 2006-04-27 itkLabelToRGBImageFilter.h

@@ -58,10 +58,19 @@ public:
     addColor( 139, 0, 139 );
     addColor( 238, 130, 238 );
     addColor( 139, 0, 0 );
+
+    m_UseBackground = false;
+    m_BackgroundValue = NumericTraits<TLabel>::Zero;
     }
 
   inline TRGBPixel operator()( const TLabel & p)
     {
+    if( m_UseBackground && p == m_BackgroundValue )
+      {
+      TRGBPixel rgbPixel;
+      rgbPixel.Set( p, p, p );
+      return rgbPixel;
+      }
     return m_Colors[ p % m_Colors.size()];
     }
 
@@ -76,30 +85,38 @@ public:
     }
 
   bool operator != (const LabelToRGBFunctor &l) const
-  { return true; }
+    { return m_UseBackground != l.m_UseBackground || m_BackgroundValue != l.m_BackgroundValue; }
+
+  void SetBackgroundValue( TLabel v ) { m_BackgroundValue = v; }
+
+  void SetUseBackground( bool b ) { m_UseBackground = b; }
 
   ~LabelToRGBFunctor() {}
 
   std::vector< TRGBPixel > m_Colors;
+
+  bool m_UseBackground;
+
+  TLabel m_BackgroundValue;
+
 };
 }  // end namespace functor
 
 
 /** \class LabelToRGBImageFilter
- * \brief Apply a colormap to a label image and put it on top of the input image
+ * \brief Apply a colormap to a label image
  *
- * Apply a colormap to a label image and put it on top of the input image. The set of colors
- * is a good selection of distinct colors. The opacity of the label image
- * can be defined by the user.
+ * Apply a colormap to a label image. The set of colors
+ * is a good selection of distinct colors.
  *
  * \author Gaëtan Lehmann. Biologie du Développement et de la Reproduction, INRA de Jouy-en-Josas, France.
  * \author Richard Beare. Department of Medicine, Monash University, Melbourne, Australia.
  *
- * \sa ScalarToRGBPixelFunctor
+ * \sa ScalarToRGBPixelFunctor LabelOverlayImageFilter
  * \ingroup Multithreaded
  *
  */
-template <typename  TInputImage, class TLabelImage, typename  TOutputImage>
+template <class TLabelImage, typename  TOutputImage>
 class ITK_EXPORT LabelToRGBImageFilter :
     public
 UnaryFunctorImageFilter<TLabelImage, TOutputImage, 
[…]

### 2006-05-15 article/Article.tex
Minor additions to article
I've kept the first hill climbing algorithm because I recently had a reviewer
of an article claim that version was the "one true" hill climbing algorithm.
It is good to be complete, anyway.

darcs-hash:20060515054736-fafb9-e667ea21d523943faeca93b1bf9ccfd3fe50f87f.gz

@@
 (i.e. recursive erosion). Other steps have also been mentioned in
-passing ({\em lower completion} in Section \ref{sect:history} that may
+passing ({\em lower completion} in Section \ref{sect:history}) that may
 be potentially time consuming. Fortunately we will discover in Section
@@
 
-{\em Is it really needed to present this algorithm given that there is
-no lower completion ? Isn't it enough to notice somewhere that an
-algorithm exist without lower completion, and which require to put
-all the pixels in the queue ?}
-
 \begin{itemize}
@@
 
-There are at least two reasons why we shouldn't use this
-approach. Firstly, the lower completion step is explicit, which we'd
-like to avoid and secondly {\em all} non minima pixels are placed on the
-queue, which means that the queue length is almost the same as the
-number of image pixels, increasing the overhead of queue
+This is the most basic version of hill climbing, and is presented here
+for completeness. There are at least two reasons why we shouldn't use
+this approach. Firstly, the lower completion step is explicit, which
+we'd like to avoid and secondly {\em all} non minima pixels are placed
+on the queue, which means that the queue length is almost the same as
+the number of image pixels, increasing the overhead of queue
 management. We shall see that it is only necessary to have the pixels
@@
 
-It must be noticed that this algorithm never propagate watershed
-lines, and so produce very thin watershed lines.
+It should be noted that this algorithm never propagates watershed
+lines, and that the watershed lines it produces have a width of 1 pixel.
 
@@
 The result of the morphological watershed is quite good, even if not perfect:
-there is some curves in the separation line of the two labels. This is the
+there are some curves in the separation line of the two labels. This is the
 result of the distance approximation in the watershed algorithm.

### 2006-05-15 article/Article.tex
added the artificial plateau image with colour.
darcs-hash:20060515070152-fafb9-3be25a780d6b4ede47f6d06a0a0fe32db592334c.gz

@@
 integration} and will be discussed in more detail in this report.
-Figure \ref{fig:cthead1-markers} and \ref{fig:difficult-cases} illustrate
-the possible problem with plateaus.
+Figures \ref{fig:cthead1-markers}, \ref{fig:circle}and \ref{fig:difficult-cases} illustrate
+the possible problems with plateaus.
 
@@
 
-Readers interested in more technical detail should refer to \cite{Roerdink2000a,Meijster2004a}.
+\begin{figure}[htbp]
+\begin{center}
+\subfigure[Input image - bright ring on constant background]{\includegraphics[scale=0.7]{circle_plateau}}
+% \subfigure[]{\includegraphics[scale=0.7]{cthead1mf-rgb}}
+% \subfigure[]{\includegraphics[scale=0.7]{cthead1mF-rgb}}
+% \subfigure[]{\includegraphics[scale=0.7]{cthead1Mf-rgb}}
+\subfigure[Morphological watershed]{\includegraphics[scale=0.7]{circle_rgb2}}
+\subfigure[ITK watershed]{\includegraphics[scale=0.7]{circle_rgb1}}
+\caption{Another example of plateau problem.
+The inside and outside of the ring have the same brightness, but the ring is arbitarily assigned to the outer region by the ITK watershed.
+It is split by the morphological watershed.\label{fig:circle}}
+\end{center}
+\end{figure}
+
+Readers interested in more technical detail should refer to
+\cite{Roerdink2000a,Meijster2004a}.
 

### 2006-05-15 itkLabelShapeImageFilter.h
add label shape filter - required by splitCells.cxx
darcs-hash:20060515122824-2fc9d-67a8c6a53ff1a227aa8b636bedc1880f52847f92.gz

@@ -1,0 +1,233 @@
+/*=========================================================================
+
+  Program:   Insight Segmentation & Registration Toolkit
+  Module:    $RCSfile: itkLabelShapeImageFilter.h,v $
+  Language:  C++
+  Date:      $Date: 2005/05/02 19:27:25 $
+  Version:   $Revision: 1.4 $
+
+  Copyright (c) Insight Software Consortium. All rights reserved.
+  See ITKCopyright.txt or http://www.itk.org/HTML/Copyright.htm for details.
+
+     This software is distributed WITHOUT ANY WARRANTY; without even 
+     the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR 
+     PURPOSE.  See the above copyright notices for more information.
+
+=========================================================================*/
+#ifndef __itkLabelShapeImageFilter_h
+#define __itkLabelShapeImageFilter_h
+
+#include "itkImageToImageFilter.h"
+#include "itkNumericTraits.h"
+#include "itkArray.h"
+#include "itkSimpleDataObjectDecorator.h"
+#include "itk_hash_map.h"
+#include "itkFixedArray.h"
+#include "itkImageRegion.h"
+#include <vector>
+
+namespace itk {
+
+/** \class LabelShapeImageFilter 
+ * \brief Given a label map, compute center of gravity and bounding box of each component
+ *
+ * LabelShapeImageFilter computes the center of gravity and bounding box
+ * of regions defined via a label map.  The image should be integral type.
+ * The filter needs all of its input image.  It behaves as a filter with
+ * an input and output. Thus it can be inserted in a pipline with other
+ * filters and the statistics will only be recomputed if a downstream
+ * filter changes.
+ *
+ * The filter passes its intensity input through unmodified.  The filter is
+ * threaded. It computes statistics in each thread then combines them in
+ * its AfterThreadedGenerate method.
+ *
+ * \ingroup MathematicalStatisticsImageFilters
+ */
+template<class ImageType>
+class ITK_EXPORT LabelShapeImageFilter : 
+    public ImageToImageFilter<ImageType, ImageType>
+{
+public:
+  /** Standard Self typedef */
+  typedef LabelShapeImageFilter Self;
+  typedef ImageToImageFilter<ImageType, ImageType>  Superclass;
[…]

### 2006-05-15 article/Article.tex
more of the article
darcs-hash:20060515142026-2fc9d-fa340f3526b977f262c2cfdfd5185152b04cd742.gz

@@
 integration} and will be discussed in more detail in this report.
-Figures \ref{fig:cthead1-markers}, \ref{fig:circle}and \ref{fig:difficult-cases} illustrate
+Figures \ref{fig:cthead1-markers}, \ref{fig:circle} and \ref{fig:difficult-cases} illustrate
 the possible problems with plateaus.
@@
 An alternative approach, known as the topological watershed, has
-recently been developed \cite{}. The topological watershed modifies
+recently been developed \cite{topologicalWatershed}. The topological watershed modifies
 the topology of the surface so that all pixels that are not watershed
@@
 
-The ITK
-watershed implementation provides a way of addressing this problem
-using a measure of significance of the boundary and allowing a
-threshold to be applied first to reduce the number of local minima.
+A common way to address this problem is to use a ignore the minima
+with a too small dynamic. Figures \ref{fig:artificial-itk} and
+\ref{fig:artificial-morpho} illustrate the effect of this technique.
+An usage example is shown in Section \ref{sect:pronucleus}.
+The ITK watershed implementation also allow a threshold to be applied
+first to reduce the number of local minima.
 
@@
 $(g)$ and $(h)$ show a marker defined by a user, and the result over the input image.
-All the watersheds were performed with $MarkWatershed = true$ and $FullyConnected = false$.}
+All the watersheds were performed with $MarkWatershed = true$ and $FullyConnected = false$.\label{fig:artificial-itk}}
 \end{center}
@@
 From $(b)$ to $(f)$, an example of reduction of the classical watershed over segmentation by minima dynamic filtering.
-$(g)$ and $(h)$ show a marker defined by a user, and the result over the input image.}
+$(g)$ and $(h)$ show a marker defined by a user, and the result over the input image.\label{fig:artificial-morpho}}
 \end{center}
@@
 \subsection{Watershed from selected minima}
+\label{sect:ws-selected-minima}
 An alternative approach leads us to the concept of markers mentioned
@@
 It is the role of $S$ in Figure \ref{MeyerAlgorithm}. We'll discuss
-in \ref{sect:implementation} how this status can be implemented.
+in Section \ref{sect:implementation} how this status can be implemented.
 
@@
 
-We provide with this article two new classes:
+  \subsection{With ITK watershed}
+
+The watershed from markers can be implemented with the ITK watershed,
+by using the technique of {\em minima imposition} exposed in Section
+\ref{sect:ws-selected-minima}. Several steps are required:
+\begin{itemize}
+  \item Impose the minima to the input image, with the
+\verb$MinimaImpositionImageFilter$
+  \item Perform the watershed with the \verb$WatershedImageFilter$
+  \item Restore the labels produced by the \verb$WatershedImageFilter$.
+\verb$WatershedImageFilter$ produce arbitrary labels which doesn't match
+the labels in the marker image. There is currently no filter able to
+find the label correspondance in ITK, so this step must be performed by
+hand with a code like the one before. We can use a
+\verb$ChangeLabelImageFilter$ to change the labels.
+\small \begin{verbatim}
+filter->Update();
+
+typedef itk::ChangeLabelImageFilter< LIType, IType > ChangeLabelType;
+ChangeLabelType::Pointer changeLabel = ChangeLabelType::New();
+changeLabel->SetInput( filter->GetOutput() );
+
+itk::ImageRegionConstIterator< LIType > wit( filter->GetOutput(),
+                filter->GetOutput()->GetLargestPossibleRegion() );
+itk::ImageRegionConstIterator< IType > mit( reader2->GetOutput(),
+                filter->GetOutput()->GetLargestPossibleRegion() );
+for ( mit.GoToBegin(), wit.GoToBegin();
+  !mit.IsAtEnd();
+  ++wit, ++mit )
+  {
+  if( mit.Get() != 0 )
+    { changeLabel->SetChange( wit.Get(), mit.Get() ); }
+  }
+\end{verbatim} \normalsize
+\end{itemize}
+
+This approach has several drawbacks:
+\begin{itemize}
+  \item it doesn't work if there some markers not separated by background
+pixels - they are considered as a single marker
+  \item restoring the labels is not included in a filter and so break the
+ITK pipeline model
+  \item it's less efficient, as shown in Section \ref{sect:perf}
+  \item it gives poor results with plateaus
+\end{itemize}
+
+A full example of implementation can be found in \verb$iwsm.cxx$.
 
@@
 \section{Timing and performance}
+\label{sect:perf}
+
+  \subsection{Classical watershed}
+\label{sect:perf-classical}
+
 The relative performance of the new filters are shown in
@@
 are not easily threadable and therefore may prove to be slower on some
-systems. However, as discussed in earlier sections, the new
+systems. Moreover, this results doesn't show the performance in the
+case where the user modify the \verb$Level$ parameters of the ITK
+filter, case where the time needed to update the filter should be a lot
+shorter. However, as discussed in earlier sections, the new
 implementation produces more stable results than the rainfall model
-and can be used in a marker based approach.
+and is more dedicated to a marker based approach.
 
@@
 \caption{Execution times. Note the FC and MB refer to FullyConnected and MarkBoundary.\label{tbl:perf}}
+\end{table}
+
+  \subsection{Marker based watershed}
+\label{sect:perf-marker}
+
+The performance in case of a marker based approach are shown in Table
+\ref{tbl:perf-marker}, and were obtained on the same system than the
+ones in Section \ref{sect:perf-classical}, with the command {\tt 
+./mperf3D images/ESCells.img images/ESCells-markers.tif}.
+The ITK watershed use the implementation presented in Section
+\ref{sect:implementation}. The extra steps of this implementation are
+also shown in this table.
+
+\begin{table}[htbp]
+\centering
+\begin{tabular}{cccccccc}
+\hline
+FC & MB & impose min &  itk watershed & match labels & change labels & total & Morph watershed \\
+\hline
+\hline
+\end{tabular}
+\caption{Execution times with marker based approach. Note the FC and MB refer to FullyConnected and MarkBoundary.\label{tbl:perf-marker}}
 \end{table}
@@
 \subsection{Pronucleus segmentation in bovine embryo}
-
+\label{sect:pronucleus}
 In this example, we use the watershed filter to split two nearly round
@@
 mask = itk.MaskImageFilter.IUC2IUC2IUC2.New(ws, fill)
-overlay = itk.LabelOverlayImageFilter.IUC2IUC2IRGBUC2.New(reader, mask, Opacity=0.5)
+overlay = itk.LabelOverlayImageFilter.IUC2IUC2IRGBUC2.New(reader, mask)
 \end{verbatim} \normalsize
@@
 mask2 = itk.MaskImageFilter.IUC2IUC2IUC2.New(relabel, fill)
-overlay2 = itk.LabelOverlayImageFilter.IUC2IUC2IRGBUC2.New(reader, mask2, Opacity=0.5)
+overlay2 = itk.LabelOverlayImageFilter.IUC2IUC2IRGBUC2.New(reader, mask2)
 \end{verbatim} \normalsize
@@
 
+\section{Related work}
+This work have lead us to develop some other filters to:
+\begin{itemize}
+  \item Find regional extrema \cite{regionalExtrema}
+  \item Surimpose a colored label image onto an image \cite{labelOverlay}
+  \item Impose the minima \cite{minimaImposition}
+  \item Label the connected component \cite{connectedComponent}
+  \item Invert the intensity of an image \cite{invert}
+  \item Enhance the morphological reconstruction \cite{recons}
+  \item Compute some informations about the shape of the labeled objects (not published yet)
+\end{itemize}
+
 \section{Acknowledgments}

### 2006-05-17 itkMorphologicalWatershedFromMarkersImageFilter.h
MarkWatershed -> MarkWatershedLine
darcs-hash:20060517121730-2fc9d-6f2fb0c983422af5dbc72c9ca5adfc9de1c224f1.gz

@@ -103,9 +103,9 @@ public:
    * is true. Set it to false do not only avoid writing watershed pixels,
    * it also decrease algorithm complexity.
    */
-  itkSetMacro(MarkWatershed, bool);
-  itkGetConstReferenceMacro(MarkWatershed, bool);
-  itkBooleanMacro(MarkWatershed);
+  itkSetMacro(MarkWatershedLine, bool);
+  itkGetConstReferenceMacro(MarkWatershedLine, bool);
+  itkBooleanMacro(MarkWatershedLine);
 
 protected:
   MorphologicalWatershedFromMarkersImageFilter();
@@ -140,7 +140,7 @@ private:
 
   bool m_FullyConnected;
 
-  bool m_MarkWatershed;
+  bool m_MarkWatershedLine;
 
 } ; // end of class
 

### 2006-05-17 itkMorphologicalWatershedImageFilter.h

@@ -96,9 +96,9 @@ public:
    * is true. Set it to false do not only avoid writing watershed pixels,
    * it also decrease algorithm complexity.
    */
-  itkSetMacro(MarkWatershed, bool);
-  itkGetConstReferenceMacro(MarkWatershed, bool);
-  itkBooleanMacro(MarkWatershed);
+  itkSetMacro(MarkWatershedLine, bool);
+  itkGetConstReferenceMacro(MarkWatershedLine, bool);
+  itkBooleanMacro(MarkWatershedLine);
 
   /**
    */
@@ -129,7 +129,7 @@ private:
 
   bool m_FullyConnected;
 
-  bool m_MarkWatershed;
+  bool m_MarkWatershedLine;
 
   InputImagePixelType m_Level;
 

### 2006-05-17 itkLabelOverlayImageFilter.h
update label overlay filters
darcs-hash:20060517131953-2fc9d-bbdcf2186467dd879a6fcc47c05acd631edb8a46.gz

@@ -10,34 +10,38 @@ namespace itk
 
 namespace Functor {  
  
+/** The functor class used internally by LabelOverlayImageFilter */
 template< class TInputPixel, class TLabel, class TRGBPixel >
 class LabelOverlay
 {
 public:
-
   LabelOverlay() 
     {
+    // provide some default value for external use (outside LabelOverlayImageFilter)
+    // Inside LabelOverlayImageFilter, the values are always initialized
     m_UseBackground = false;
     m_BackgroundValue = NumericTraits<TLabel>::Zero;
     }
 
-  inline TRGBPixel operator()(  const TInputPixel & p1,
-				const TLabel & p2)
+  inline TRGBPixel operator()(  const TInputPixel & p1, const TLabel & p2)
     {
     if( m_UseBackground && p2 == m_BackgroundValue )
       {
+      // value is background
+      // return a gray pixel with the same intensity than the input pixel
       typename TRGBPixel::ValueType p = static_cast< typename TRGBPixel::ValueType >( p1 );
       TRGBPixel rgbPixel;
       rgbPixel.Set( p, p, p );
       return rgbPixel;
       }
+
+     // tainte the input pixel with the colored one returned by the color functor.
      TRGBPixel rgbPixel;
      TRGBPixel opaque = m_RGBFunctor(p2);
      for( int i = 0; i<3; i++)
        {
        rgbPixel[i] = static_cast< typename TRGBPixel::ValueType >( opaque[i] * m_Opacity + p1 * ( 1.0 - m_Opacity ) );
        }
-std::cout << opaque << std::endl;
      return rgbPixel;
     }
 
@@ -68,7 +72,9 @@ std::cout << opaque << std::endl;
  *
  * Apply a colormap to a label image and put it on top of the input image. The set of colors
  * is a good selection of distinct colors. The opacity of the label image
- * can be defined by the user.
+ * can be defined by the user. The user can also choose if he want to use a background
+ * and which label value is the background. A background label produce a gray pixel
+ * with the same intensity than the input one.
  *
  * \author Gaëtan Lehmann. Biologie du Développement et de la Reproduction, INRA de Jouy-en-Josas, France.
  *
@@ -114,15 +120,20 @@ public:
      { this->SetInput2( input ); }
 
   /** Get the label image */
-  LabelImageType * GetLabelImage()
+  const LabelImageType * GetLabelImage() const
[…]

### 2006-05-17 itkLabelToRGBImageFilter.h

@@ -8,6 +8,7 @@ namespace itk
 
 namespace Functor {  
  
+/** The functor class used internally by LabelToRGBImageFilter */
 template< class TLabel, class TRGBPixel >
 class LabelToRGBFunctor
 {
@@ -59,18 +60,24 @@ public:
     addColor( 238, 130, 238 );
     addColor( 139, 0, 0 );
 
+    // provide some default value for external use (outside LabelToRGBImageFilter)
+    // Inside LabelToRGBImageFilter, the values are always initialized
     m_UseBackground = false;
     m_BackgroundValue = NumericTraits<TLabel>::Zero;
     }
 
   inline TRGBPixel operator()( const TLabel & p)
     {
+    // value is background
+    // return a gray pixel with the same intensity than the label pixel
     if( m_UseBackground && p == m_BackgroundValue )
       {
       TRGBPixel rgbPixel;
       rgbPixel.Set( p, p, p );
       return rgbPixel;
       }
+
+    // else, return a colored pixel from the color table
     return m_Colors[ p % m_Colors.size()];
     }
 
@@ -107,7 +114,9 @@ public:
  * \brief Apply a colormap to a label image
  *
  * Apply a colormap to a label image. The set of colors
- * is a good selection of distinct colors.
+ * is a good selection of distinct colors. The user can choose to use a background
+ * value. In that case, a gray pixel with the same intensity than the background
+ * label is produced.
  *
  * \author Gaëtan Lehmann. Biologie du Développement et de la Reproduction, INRA de Jouy-en-Josas, France.
  * \author Richard Beare. Department of Medicine, Monash University, Melbourne, Australia.
@@ -145,9 +154,11 @@ public:
   /** Method for creation through the object factory. */
   itkNewMacro(Self);
   
+  /** Set/Get the background value */
   itkSetMacro( BackgroundValue, LabelPixelType );
   itkGetConstReferenceMacro( BackgroundValue, LabelPixelType );
 
+  /** Set/Get if one of the labels must be considered as background */
   itkSetMacro( UseBackground, bool );
   itkGetConstReferenceMacro( UseBackground, bool );
   itkBooleanMacro(UseBackground);

### 2006-05-17 article/Article.tex
marker watershed timing
Added timing for marker based watershed. Not sure why there was such a big
difference between fully and face connected for the itk watershed. Difference
due to minima imposition isn't enough to account for it.

darcs-hash:20060517040544-fafb9-ef02a7f12e50e5a2ea52bd3a54b05b93237644fd.gz
continue the article
darcs-hash:20060517134412-2fc9d-c1dc3633a52f22ceb1f8894fb8ea63a10853ec4c.gz

@@
 transform is given in Section \ref{sect:algorithms}.
+
+Section \ref{sect:using-ws-itk} propose some applications of watershed
+with ITK.
 
@@
 or {\em recursive conditional erosion} from the seleted regional
-minima. Figure \ref{fig:recon} illustrates the results of apply a
+minima. This process is also called {\em minima imposition}. Figure
+\ref{fig:recon} illustrates the results of apply a
 reconstruction by erosion to a 1D image. The original image is shown
@@
 \centering
-\begin{tabular}{cccccccc}
+\begin{tabular}{cc|ccc|cc}
 \hline
-FC & MB & Find Min &  Label mins & Watershed from markers & Total & Morph watershed &  itk watershed \\
+FC & MB & Find min &  Label min & Ws from markers & Morph ws &  ITK ws \\
 \hline
-0  &    0  &     1.948  & 0.522  & 5.752  & 8.222   & 8.218  & 12.596 \\
-0  &    1  &     1.922  & 0.516  & 6.724  & 9.162   & 9.28   & 12.61 \\
-1  &    0  &     4.26   & 0.154  & 13.274 & 17.688  & 18.158 & 12.61 \\
-1  &    1  &     4.284  & 0.15   & 17.23  & 21.664  & 21.76  & 12.606 \\
+\hline
+0 &      0 &      1.61 &   0.0476 & 3.72 &   6.14 &   - \\
+0 &      1 &      1.59 &   0.0428 & 6.25 &   8.95 &   - \\
+1 &      0 &      3.15 &   0.0465 & 7.77 &   11.5 &   6.48 \\
+1 &      1 &      3.19  &  0.0462  &14.8 &   19.2  &  - \\
+% 0  &    0  &     1.948  & 0.522  & 5.752  & 8.222   & 8.218  & 12.596 \\
+% 0  &    1  &     1.922  & 0.516  & 6.724  & 9.162   & 9.28   & 12.61 \\
+% 1  &    0  &     4.26   & 0.154  & 13.274 & 17.688  & 18.158 & 12.61 \\
+% 1  &    1  &     4.284  & 0.15   & 17.23  & 21.664  & 21.76  & 12.606 \\
 \hline
 \end{tabular}
-\caption{Execution times. Note the FC and MB refer to FullyConnected and MarkBoundary.\label{tbl:perf}}
+\caption{Execution times. Note the FC and MWL refer to FullyConnected and MarkWatershedLine.\label{tbl:perf}}
 \end{table}
@@
 \label{sect:perf-marker}
-
 The performance in case of a marker based approach are shown in Table
@@
 \centering
-\begin{tabular}{cccccccc}
+\begin{tabular}{cc|cccc|cc}
 \hline
-FC & MB & impose min &  itk watershed & match labels & change labels & total & Morph watershed \\
+FC & MWL & Impose min &  ITK ws & Match labels & Change labels & Morph ws & ITK ws \\
 \hline
 \hline
+0 &      0 &      -    &   -   &    -      & -     &  3.67 &   - \\
+0 &      1 &      -    &   -   &    -      & -     &  6.53 &   - \\
+1 &      0 &      5.46 &   7.8 &    0.0192 & 0.227 &  14.6 &   13.5 \\
+1 &      1 &      -    &   -   &    -      & -     &  34.8 &   - \\
+% 0   &  0  &  2.47776 & 8.12256 & 4.29153e-07 & 0.0990558 &  10.6994 & 3.70646\\
+% 0   &  1  &  2.24303 & 8.20924 & 9.53674e-07 & 0.0977524 &  10.55   & 6.22279\\
+% 1   &  0  &  5.92089 & 12.2889 & 4.29153e-07 & 0.0971172 &  18.3069 & 10.7172\\
+% 1   &  1  &  6.08657 & 12.222  & 5.72205e-07 & 0.0927572 &  18.4014 & 18.4605\\
+\hline
 \end{tabular}
-\caption{Execution times with marker based approach. Note the FC and MB refer to FullyConnected and MarkBoundary.\label{tbl:perf-marker}}
+\caption{Execution times with marker based approach. Note the FC and MWL refer to FullyConnected and MarkWatershedLine.\label{tbl:perf-marker}}
 \end{table}
@@
 \section{Using watersheds in ITK}
+\label{sect:using-ws-itk}
+
+It should be noticed that all the images shown in this article have been
+produced with ITK and can be easily reproduced and tested by running 
+{\em ctest} in the build tree.
 
@@
 The program is in python for conciseness\footnote{And also to promote
-python wrappers.}.
+python wrappers.}. A C++ version, \verb$splitPronucleus2D.cxx$, is also
+available.
 
@@
 ws = itk.MorphologicalWatershedImageFilter.IUC2IUC2.New(idistance,
-                                                        MarkWatershed=False,
+                                                        MarkWatershedLine=False,
                                                         Level=10)
 mask = itk.MaskImageFilter.IUC2IUC2IUC2.New(ws, fill)
-overlay = itk.LabelOverlayImageFilter.IUC2IUC2IRGBUC2.New(reader, mask)
+overlay = itk.LabelOverlayImageFilter.IUC2IUC2IRGBUC2.New(reader,
+                                                          mask,
+                                                          UseBackground=True)
 \end{verbatim} \normalsize
@@
 mask2 = itk.MaskImageFilter.IUC2IUC2IUC2.New(relabel, fill)
-overlay2 = itk.LabelOverlayImageFilter.IUC2IUC2IRGBUC2.New(reader, mask2)
+overlay2 = itk.LabelOverlayImageFilter.IUC2IUC2IRGBUC2.New(reader,
+                                                           mask2
+                                                           UseBackground=True)
 \end{verbatim} \normalsize
 
-ITK watershed clearly produce a wrong segmentation in that case.
-The result of the morphological watershed is quite good, even if not perfect:
+ITK watershed clearly produce a wrong segmentation in that case. However, using
+a floating point image to store the distance map would have produce a better
+result.
+The segmentation of the morphological watershed is quite good, even if not perfect:
 there are some curves in the separation line of the two labels. This is the

### 2006-05-22 article/Article.tex
add WrapITK note
darcs-hash:20060522111717-2fc9d-faff6aca2e507b0c4cfc40533e58467f786a555c.gz

@@
 and \verb$MorphologicalWatershedFromMarkersImageFilter$ are the usal steps
-of the waterhed usage, the \verb$HMinimaImageFilter$ is a convenient way
+of the watershed usage, the \verb$HMinimaImageFilter$ is a convenient way
 to avoid the classical watershed's over segmentation by allowing the user
@@
 selected with \verb$SetLevel()$.
+
+  \subsection{WrapITK integration}
+
+The new filters have been fully tested with python, with the WrapITK project.
+An example of python usage is shown in Section \ref{sect:pronucleus}.
+The sources of this article contain the configuration files needed to build
+build the new filters with WrapITK, as an external project - see \cite{WrapITK}
+for more information about how to build and use an external project.
 

### 2006-05-22 article/Article.tex
enhance floating type notes
darcs-hash:20060522113650-2fc9d-124aaded080255308f83d1473323b91d3f602a6a.gz

@@
 a floating point image to store the distance map would have produce a better
-result.
+result very similar to the one of the morphological watershed. It seems to be
+another example of plateaus problem, where each level curve is seen as a plateau
+when the pixel type of the distance image is an integral.
+
 The segmentation of the morphological watershed is quite good, even if not perfect:

### 2006-05-22 article/Article.tex
update timing results
darcs-hash:20060522205312-2fc9d-9263eb42cd5c6d0c303bd1eb2585628aa7ab2219.gz

@@
 Table~\ref{tbl:perf}. The timings were obtained using the {\tt
-./perf3D images/ESCells.img}. The times for steps involved in a
+./perf3D images/ESCells.img} on an Athlon 64 
+Processor 2800+ (1802 MHz) with 512Kb cache, 512 Mb of RAM running
+Mandriva linux 64 bits with gcc 4.1.1. The input image have a size of
+$371 \times 371 \times 34$.
+
+\begin{table}[htbp]
+\centering
+\begin{tabular}{cc|ccc|cc}
+\hline
+FC & MWL & Find min &  Label min & Ws from markers & Morph ws &  ITK ws \\
+\hline
+\hline
+0 &      0 &      1.39 &   0.0566 & 2.74 &   4.61 &   7.47 \\
+0 &      1 &      1.38 &   0.056  & 4.34 &   6.53 &   - \\
+1 &      0 &      2.85 &   0.0568 & 5.8  &   9.54 &   - \\
+1 &      1 &      2.85 &   0.0568 & 10.5 &   14.4 &   - \\
+% 0  &    0  &     1.948  & 0.522  & 5.752  & 8.222   & 8.218  & 12.596 \\
+% 0  &    1  &     1.922  & 0.516  & 6.724  & 9.162   & 9.28   & 12.61 \\
+% 1  &    0  &     4.26   & 0.154  & 13.274 & 17.688  & 18.158 & 12.61 \\
+% 1  &    1  &     4.284  & 0.15   & 17.23  & 21.664  & 21.76  & 12.606 \\
+\hline
+\end{tabular}
+\caption{Execution times. Note the FC and MWL refer to FullyConnected and MarkWatershedLine.\label{tbl:perf}}
+\end{table}
+
+The times for steps involved in a
 morphological watershed are shown in the first few columns, with the
@@
 
-\begin{table}[htbp]
-\centering
-\begin{tabular}{cc|ccc|cc}
-\hline
-FC & MB & Find min &  Label min & Ws from markers & Morph ws &  ITK ws \\
-\hline
-\hline
-0 &      0 &      1.61 &   0.0476 & 3.72 &   6.14 &   - \\
-0 &      1 &      1.59 &   0.0428 & 6.25 &   8.95 &   - \\
-1 &      0 &      3.15 &   0.0465 & 7.77 &   11.5 &   6.48 \\
-1 &      1 &      3.19  &  0.0462  &14.8 &   19.2  &  - \\
-% 0  &    0  &     1.948  & 0.522  & 5.752  & 8.222   & 8.218  & 12.596 \\
-% 0  &    1  &     1.922  & 0.516  & 6.724  & 9.162   & 9.28   & 12.61 \\
-% 1  &    0  &     4.26   & 0.154  & 13.274 & 17.688  & 18.158 & 12.61 \\
-% 1  &    1  &     4.284  & 0.15   & 17.23  & 21.664  & 21.76  & 12.606 \\
-\hline
-\end{tabular}
-\caption{Execution times. Note the FC and MWL refer to FullyConnected and MarkWatershedLine.\label{tbl:perf}}
-\end{table}
-
 \subsection{Marker based watershed}
@@
 \hline
-0 &      0 &      -    &   -   &    -      & -     &  3.67 &   - \\
-0 &      1 &      -    &   -   &    -      & -     &  6.53 &   - \\
-1 &      0 &      5.46 &   7.8 &    0.0192 & 0.227 &  14.6 &   13.5 \\
-1 &      1 &      -    &   -   &    -      & -     &  34.8 &   - \\
+0 &      0 &      2.32 &   6.81 &   0.0295 & 0.273 &  3.12 &   9.43 \\
+0 &      1 &      -    &   -    &   -      & -     &  5    &   - \\
+1 &      0 &      -    &   -    &   -      & -     &  6.57 &   - \\
+1 &      1 &      -    &   -    &   -      & -     &  11.6 &   - \\
 % 0   &  0  &  2.47776 & 8.12256 & 4.29153e-07 & 0.0990558 &  10.6994 & 3.70646\\

### 2006-05-23 article/Article.tex
Added example using watershed lines
darcs-hash:20060523053318-fafb9-7c42f621007aa064e7f70e90149ca6940f4cc5ac.gz

@@
 \author{Richard Beare{$^1$} {\small{and}} Ga\"etan Lehmann{$^2$}}
-\authoraddress{{$^1$}Department of Medicine, Moansh University, Australia.\\ 
+\authoraddress{{$^1$}Department of Medicine, Monash University, Australia.\\ 
 {$^2$}Unit\'e de Biologie du D\'eveloppement et de la Reproduction, Institut National de la Recherche Agronomique, 78350 Jouy-en-Josas, France}

### 2006-05-23 itkMorphologicalWatershedFromMarkersImageFilter.h

@@ -107,6 +107,10 @@ public:
   itkGetConstReferenceMacro(MarkWatershedLine, bool);
   itkBooleanMacro(MarkWatershedLine);
 
+  itkSetMacro(PadImageBoundary, bool);
+  itkGetConstReferenceMacro(PadImageBoundary, bool);
+  itkBooleanMacro(PadImageBoundary);
+
 protected:
   MorphologicalWatershedFromMarkersImageFilter();
   ~MorphologicalWatershedFromMarkersImageFilter() {};
@@ -141,6 +145,7 @@ private:
   bool m_FullyConnected;
 
   bool m_MarkWatershedLine;
+  bool m_PadImageBoundary;
 
 } ; // end of class
 

### 2006-05-23 article/Article.tex
Minor corrections and additions to article.
darcs-hash:20060523061724-fafb9-60e8e3c282a308e5950666fb53147fc8df85152c.gz

@@
 
-Section \ref{sect:usingWT} provides an introduction to a
-methodology known as {\em marker based} watershed. Using markers in
-watershed transforms is a well established technique in the
-morphological image analysis community and has proved very useful for
-many applications. The aim of this part of the report is to describe
-the technique to a wider audience \footnote{It should also be noted
-that the lack of a marker based watershed motivated the recent work on
-ITK filter classes.}.
+Section \ref{sect:usingWT} provides an introduction to a methodology
+known as {\em marker based} watershed. Using markers in watershed
+transforms is a well established technique in the morphological image
+analysis community and has proved very useful for many
+applications. The aim of this part of the report is to describe the
+technique to a wider audience \footnote{The lack of a marker based
+watershed motivated this work on ITK filter classes.}.
 
@@
 
-An Obvious alternative that is very similar in practice is to 
+An obvious alternative that is very similar in practice is to 
 smooth the image so that there are fewer regional minima.
@@
 minima. This process is also called {\em minima imposition}. Figure
-\ref{fig:recon} illustrates the results of apply a
+\ref{fig:recon} illustrates the result of applying a
 reconstruction by erosion to a 1D image. The original image is shown
@@
 regional minima. Reconstruction approaches are often used to find
-regional minima of integer type images, but subtracting 1, carrying
-out a recursive dilation and subtracting the two. The same concept
-does not extend to floating point data. However, if regional minima
-are available the algorithms we are discussing can be applied to
-floating point data.}.
+regional minima of integer type images, by subtracting 1, carrying out
+a recursive dilation and subtracting the two. The same concept does
+not extend to floating point data. However, if regional minima are
+available the algorithms we are discussing can be applied to floating
+point data. It is possible to use other techniques, such as flooding
+approaches, to find regional minima of floating point images.}.
 
@@
 
-Algorithm 3 requires slightly more complexity to store label
-information in the queue.
-
 {\em a summary of our findings comparing the last 2 algorithms}
@@
 
-\caption{Three difficult test images.
-The first image, plateau, show that ITK's watershed produce a bad result on this image: only one minima take all the image.
-This problem is also shown in Figure \ref{fig:cthead1-markers}.
-The morphological watershed give a better result, but not yet perfect: the separation line should not be curved as it is.
-This problem is also visible in Figure \ref{fig:embryo-mws}.
-The next image, thick line, again show that ITK watershed can produce strange results with a green label which is where we should have the light blue one.
-Morphological watershed gives good results.
-The last image, button hole, is a very difficult case for watershed algorithms which mark watershed line: the watershed pixel at the bottom of the image stop the label propagation.
-There is several possible behavior: produce a large watershed line which cover all the zone (Vincent-Soille algorithm, not shown here), let the label on the top of the image propagate in the zone (Meyer algorithm), split the zone in 2 zones with the 2 labels of the bottom of the image (topological watershed, not shown here).\label{fig:difficult-cases}}
+\caption{Three difficult test images. The first image, plateau, illustrates a 
+bad result produced by ITK's watershed in which most of the area is
+assigned to one minima. This problem is also shown in Figure
+\ref{fig:cthead1-markers}. The morphological watershed give a better
+result, but not yet perfect: the separation line should not be curved
+as it is. This problem is also visible in Figure
+\ref{fig:embryo-mws}. The reason for the problem is as follows - the 
+example is attempting to split a plateau into two regions, and should
+therefore produce a Voronoi tesselation, with the two markers being
+separated by a straight line. This does not happen because various
+approximations have been made, as mentioned in Section \ref{}, that
+influence the estimation of the distance function. Even if a more
+accurate underlying distance function was used the computations are
+still being carried out on a grid and so are limited by the digital
+representation of lines.
+
+The next image, thick line, again show that ITK watershed can produce
+strange results with a green label which is where we should have the
+light blue one. Morphological watershed gives good results. The last
+image, button hole, is a very difficult case for watershed algorithms
+which mark watershed lines: the watershed pixel at the bottom of the
+image stop the label propagation. There are several possible
+behaviors: produce a large watershed line which covers all the zone
+(Vincent-Soille algorithm, not shown here), let the label on the top
+of the image propagate into the zone (Meyer algorithm), split the zone
+in 2 zones with the 2 labels of the bottom of the image (topological
+watershed, not shown here).\label{fig:difficult-cases}}
 \end{center}
@@
 
-\verb$MorphologicalWatershedFromMarkersImageFilter$ implement the
+\verb$MorphologicalWatershedFromMarkersImageFilter$ implements
 algorithms \ref{sect:BeucherAlgorithm} and \ref{sect:MeyerAlgorithm}.
-They are accessible respectively when the attribute \verb$MarkWatershed$
-is \verb$false$ or is \verb$true$. The filter is templated over the
-input image type, which can be any scalar type, and the label image type
-which should be an integer. The filter take 2 inputs:
+Setting attribute \verb$MarkWatershedLine$ to \verb$true$ selects
+algorithm \ref{sect:MeyerAlgorithm} (Meyer) while setting to
+\verb$false$ selects algorithm \ref{sect:BeucherAlgorithm} (Beucher).
+The filter is templated over the input image type, which can be any
+scalar type, and the label image type which should be an integer. The
+filter takes 2 inputs:
 \begin{itemize}
@@
 
-The algorithm \ref{sect:MeyerAlgorithm} require to keep the status of all
-the pixels: {\em already added to the queue} or {\em not yet added to the queue}.
-This task is often done by using negative values in the output image
-for the pixels not yet added to the queue. In our implementation, we have
-chosen to use a binary image to store this, while labels always take values
-above 0 and watershed lines are marked with value 0. The negative values
-option have several drawbacks:
+Algorithm \ref{sect:MeyerAlgorithm} requires the status of all pixels
+to be tracked -- pixels are either {\em already added to the queue} or
+{\em not yet added to the queue}.  This task is often done by using
+negative values in the output image for the pixels not yet added to
+the queue. In our implementation, we have chosen to use a binary image
+to store this, while labels always take values above 0 and watershed
+lines are marked with value 0. The negative values option have several
+drawbacks:
 \begin{itemize}
@@
 
-We prefered using a separated binary image (of type \verb$itk::Image<bool, dim>$)
-to store this information. This solution avoid the need to copy the image at
-the end, and so should be more efficient in computation time, and require only
-one bit per pixel and so should be more memory efficient as well.
+{\em Don't think this should be here. Is there any compiler that will
+represent bool images as bit vectors? It might save memory but will
+increase access time}.  We prefered using a separated binary image (of
+type \verb$itk::Image<bool, dim>$) to store this information. This
+solution avoid the need to copy the image at the end, and so should be
+more efficient in computation time, and require only one bit per pixel
+and so should be more memory efficient as well.
 
@@
 
-\verb$MorphologicalWatershedImageFilter$ perform a watershed transformation
-without marker. It encapsulate several filters:
+\verb$MorphologicalWatershedImageFilter$ performs a watershed transformation
+without markers. It encapsulate several filters:
 \begin{itemize}

### 2006-05-23 article/Article.tex
Added label
darcs-hash:20060523062722-fafb9-159f48d60a0d3b7c980c6f3611197cc00da1772f.gz

@@
 \subsection{Hill climbing}
+\label{sect:hillclimbing}
 Hill climbing is an approximation that eliminates the need to maintain
@@
 separated by a straight line. This does not happen because various
-approximations have been made, as mentioned in Section \ref{}, that
-influence the estimation of the distance function. Even if a more
-accurate underlying distance function was used the computations are
-still being carried out on a grid and so are limited by the digital
-representation of lines.
-
+approximations have been made, as mentioned in Section
+\ref{sect:hillclimbing}, that influence the estimation of the distance
+function. Even if a more accurate underlying distance function was
+used the computations are still being carried out on a grid and so are
+limited by the digital representation of lines.
 The next image, thick line, again show that ITK watershed can produce

### 2006-05-29 article/Article.tex
Article updates and gel segmentation code
darcs-hash:20060529003507-fafb9-322bfc10e63dd03b7bfa6b3de09b8a9d4048ce10.gz

@@
 $(g)$ and $(h)$ show a marker defined by a user, and the result over the input image.
-All the watersheds were performed with $MarkWatershed = true$ and $FullyConnected = false$.\label{fig:artificial-itk}}
+All the watersheds were performed with $MarkWatershedLine = true$ and $FullyConnected = false$.\label{fig:artificial-itk}}
 \end{center}
@@
 
-\subsection{???}
+\subsection{2D electrophoresis gel segmentation}
+Protein separation in two dimensions on electrophoresis gels is one of
+the many situations where image segmentation techniques are commonly
+used to make measurements of spots which are often touching. This
+section describes a multi-stage watershed approach and illustrates the
+use of the watershed transform on gradient images and the use of
+watershed lines, which are a new capability of the filters introduced
+in this paper, as markers. This technnique almost certainly isn't the
+best way of segmenting gels, and probably won't work on many gels
+without tuning. It is simply offered as an example of how watershed
+transforms may be used. There are many variants of the basic
+algorithm, some of which are discussed in the text.
+
+The algorithm source code is in {\tt segGel.cxx} and the images used
+in this example are from \url{http://www-lecb.ncifcrf.gov/flicker/}.
+
+The original image is shown in Figure \ref{fig:origGel}. The spots we
+are interested in are dark, and frequently touching. In gels often
+have all sorts of artifacts, ranging from streaks along separation
+dimensions to an over abundance of a protein that leads to a large
+dark spot that may saturate the image brightness (i.e be darker than
+the dark floor of the imaging system). We will not address these sorts
+of issues here.
+
+The basic steps are
+\begin{itemize}
+\item {\bf Find spot markers.} Here we'll use a combination of smoothing 
+and minima detection. Depending on the noise levels of the image there
+are likely to be minima in bright regions of the image. It may be
+possible to discard a number of minima based on brightness, depending
+on how certain we are about the maximum brightness of a valid spot.
+
+The results of an unmodified minima detection after median smoothing
+is shown in Figure \ref{fig:minimaGel}.
+\item {\bf Find background marker.} We'll use a watershed on the raw image 
+to generate a background marker. The background marker will be the
+watershed line. Setting attribute \verb$MarkWatershedLine$ to
+\verb$true$ will result in the watershed line being marked with zero 
+values. In this situation we are using the watershed to tesselate the
+image into regions. The region boundaries will be in the brightest
+parts of the image between markers. Figure \ref{fig:backgroundGel} 
+illustrates the results of this step.
+\item {\bf Combine markers.} The spot and background markers are combined 
+by adding them together, after using the binary threshold filter to
+generate the background marker from the first watershed -- see Figure \ref{fig:combinedGel}.
+\item {\bf Compute gradient.} A gradient of the raw image is then computed. 
+We use the {\tt GradientRecursiveGaussianImageFilter}, but many
+options are available. The gradient image is shown in Figure
+\ref{fig:gradientGel}.
+\item {\bf Compute watershed transform} using the gradient image as the 
+control surface. Figure \ref{fig:resultGel}.
+\item {\bf Optional additional watershed steps.} The results of this sort 
+of segmentation, based on an image gradient, often appear to
+underestimate the spot size. This is because the watershed transform
+will produce edges along image ridges, which correspond to the
+locations of maximum image gradient. Unfortunately this tends not to
+correspond to the location a human observer would select.
+
+This easy easy to understand if a gaussian spot is considered -- the
+position of maximal gradient is halfway down the edge of the spot,
+while a human would tend to mark the edge as being near the point
+where the spot brightness becomes indistinguishable from the
+background brightness. This location is often closer to the position
+of the maximum of the second derivative.
+
+Therefore we can ``improve'' the segmentation by using the previous
+results to develop a new set of markers and apply the watershed
+transform using the second derivative as a control surface. It is
+sensible to use a different set of spot markers to the original
+because there are two maxima of second derivative, and the original
+markers will tend to be inside both, while the results of the first
+stage will be contain the closer maxima. Figures
+\ref{fig:combinedGel2}, \ref{fig:gradientGel2} and \ref{fig:resultGel2}
+illustrate these steps.
+
+\item {\bf Postprocessing.} This implementation doesn't do any post 
+processing, but in most cases it would be a useful step. Error
+situations, such as spots being a crazy size, shape or brightness
+could be detected brought to the attention of the user or attempts
+could be made to automatically correct the result.
+\end{itemize}
+\begin{figure}[htbp]
+\begin{center}
+\subfigure[Input]{\includegraphics[scale=0.4]{leukemia-ALL-Human-b01633-Flicker-NCI}\label{fig:origGel}}
+\subfigure[Regional minima]{\includegraphics[scale=0.4]{gel_minima}\label{fig:minimaGel}}
+\subfigure[Background marker]{\includegraphics[scale=0.4]{gel_tesselation}\label{fig:backgroundGel}}
+
+\subfigure[All markers]{\includegraphics[scale=0.4]{gel_allmarkers}\label{fig:combinedGel}}
+\subfigure[Gradient]{\includegraphics[scale=0.4]{gel_gradient}\label{fig:gradientGel}}
+\subfigure[Result (background marker removed)]{\includegraphics[scale=0.4]{gel_result}\label{fig:resultGel}}
+\subfigure[All markers - stage 2]{\includegraphics[scale=0.4]{gel_allmarkers2}\label{fig:combinedGel2}}
+\subfigure[Second derivative (gamma corrected)]{\includegraphics[scale=0.4]{gel_gradient2_gamma2}\label{fig:gradientGel2}}
+\subfigure[Result stage 2 (background marker removed)]{\includegraphics[scale=0.4]{gel_result2}\label{fig:resultGel2}}
+
+\caption{Steps used in segmenting a gel using multiple watershed stages.}
+\end{center}
+\end{figure}
+
+A cursory examination of the results will show that segmentation is
+far from perfect. Some touching spots only have one marker so only one
+spot is found and there are markers placed in what should probably be
+condsidered as background. These problems may be addressed in a number
+of ways, such as performing more careful selection of regional minima
+and sharpening the original image using deconvolution, as well as
+various post processing options. We are not going to explore these
+issues here because the aim is to illustrate a few ways in which the
+watershed transform can be used.
 
@@
 There is a level of erroneous folklore surrounding the watershed
-transform and it is hoped that this report helps to clear the air.
+transform and it is hoped that this report helps to clear the air, as
+well as providing a useful tutorial/introduction to a useful tool.
 
@@
   \item Find regional extrema \cite{regionalExtrema}
-  \item Surimpose a colored label image onto an image \cite{labelOverlay}
+  \item Superimpose a colored label image onto an image \cite{labelOverlay}
   \item Impose the minima \cite{minimaImposition}

### 2006-06-08 article/Article.tex
try to clean and complete the article
darcs-hash:20060608151103-2fc9d-3e0667377f66e4c80494342e372b5ee89fb24101.gz

@@
 \cite{Beucher93a,Meyer1994a}. Meyer defined the watershed transform in terms of {\em
-topographic distances} through {\em lower complete} images. These
+topographic distances} through {\em lower complete} images\footnote{A lower complete
+image is an image where there is no plateaus, excepted plateaus
+which are regional minima. A common way to compute the lower completion of an
+image, is to add the distance from the nearest lower pixel to the pixels of all
+the plateaus. Lower completion have the property to produce the same image than
+the input image if the input image is already lower complete. By lower completing
+an image, we introduce a new order relation in the plateaus pixels which can be
+used to solve the plateaus problem with watershed.}. These
 definitions rigorously defined the behavior in the presence of
@@
 \subfigure[Input image]{\includegraphics[scale=0.7]{cthead1}}
-\subfigure[User defined labels colored with LabelOverlayImageFilter]{\includegraphics[scale=0.7]{cthead1-markers-rgb}}
+\subfigure[User defined labels]{\includegraphics[scale=0.7]{cthead1-markers-rgb}}
 % \subfigure[]{\includegraphics[scale=0.7]{cthead1mf-rgb}}
@@
 \subfigure[ITK watershed]{\includegraphics[scale=0.7]{cthead1itk-rgb}}
-\caption{Example of plateau problem. The blue marker is clearly extended where it should not with the ITK watershed.\label{fig:cthead1-markers}}
+\caption{Example of plateau problem.
+The blue marker is clearly extended where it should not with the ITK watershed.
+Labels are colorized and superimposed on the input image.\label{fig:cthead1-markers}}
 \end{center}
@@
 
-\subsection{The topological watershed}
+% \subsection{The topological watershed}
 An alternative approach, known as the topological watershed, has
@@
 
-I don't understand the implications of the topological watershed yet.
-
-\section{Rainfall model}
+% I don't understand the implications of the topological watershed yet.
+
+% \section{Rainfall model}
 \label{sect:rainfall}
-{\em where should we put this?}  The original concept behind the
+The original concept behind the
 watershed transform was rain falling on a terrain and flowing down
@@
 
-Using ordered queues.
-
-This algorithm does not mark the watershed lines.
-
-The queue only contains the borders of the growing regions.
-
 This algorithm take a grayscale image $I$ and a marker image $M$
@@
 
+It should be noticed that:
+\begin{itemize}
+  \item this algorithm use ordered queues,
+  \item this algorithm doesn't mark the watershed lines,
+  \item the ordered queue only contains the borders of the growing regions.
+\end{itemize}
+
 % \begin{itemize}
@@
 \label{sect:MeyerAlgorithm}
-
-Also using ordered queues.
 
@@
 
-This algorithm mark the watershed lines.
-
-The queue only contains the borders of the growing regions.
-
-It should be noted that this algorithm never propagates watershed
+It should be noticed that:
+\begin{itemize}
+  \item this algorithm mark the watershed lines,
+  \item the queue only contains the borders of the growing regions,
+  \item this algorithm never propagates watershed
 lines, and that the watershed lines it produces have a width of 1 pixel.
+Also, the watershed lines form a connected obejct if we consider the
+opposite connectivity than the one used in the algorithm: with a 2D
+image the watershed lines are 8-connected if the algorithm is run
+as 4-connected, and 4-connected if it is run as 8-connected.
+\end{itemize}
 
@@
 
-The difference between algorithms 2 and 3 is the order in which
-labelling and queueing happens.
-
-{\em a summary of our findings comparing the last 2 algorithms}
-
-The new work implements algorithms 2 and 3 -- algorithm 2 is used when
-the watershed lines are not marked while algorithm 3 is used when they
-are.
+% The difference between algorithms 2 and 3 is the order in which
+% labelling and queueing happens.
+% 
+% {\em a summary of our findings comparing the last 2 algorithms}
+% 
+% The new work implements algorithms 2 and 3 -- algorithm 2 is used when
+% the watershed lines are not marked while algorithm 3 is used when they
+% are.
 
@@
 in 2 zones with the 2 labels of the bottom of the image (topological
-watershed, not shown here).\label{fig:difficult-cases}}
+watershed, not shown here). See \cite{najman03watershed} for more
+details.\label{fig:difficult-cases}}
 \end{center}
@@
 
-\section{Finding regional minima}
-The marker based watershed filter can be used as a convention
-watershed algorithm by finding the regional minima and using them as
-markers. Regional minima (and maxima) are often useful, and there are
-a number of different ways of finding them:
-\begin{itemize}
-\item Using reconstruction approaches mentioned earlier - 
-the problem with these is that a value of 1 is subtracted from the
-original to produce the image from which the reconstruction is carried
-out, which makes the approach less useful for floating point
-types. The advantage of this approach is that it can be easily
-extended to find {\em image dynamics}, which are often more stable and
-less subject to noise than regional minima.
-\item The way ITK does it, which is related to the connected 
-component labeller. Finds all flat regions and then figures out which
-are plateaus, minima and maxima by looking at the boundary values.
-\item Flooding transform (not sure where this came from) which 
-sets anything not a regional minima to a specified max value. The
-algorithm assumes that the border is set to the maximum value. Pixels
-are visited in raster order. If a pixel $A$ has a neighbour $B$ and
-$P(A) > p(B)$ (i.e grey level of A is greater than B) then all pixels
-connected to $A$ with value $P(A)$ are set to the maximum value. This
-is a recursive algorithm that finds pixels that cannot be a regional
-minimum (because they have a lower neighbour) and then eliminates any
-connected region that also cannot be a regional minima.
-
-The result would then need to be thresholded and labelled to produce
-markers. This approach is not limited by the pixel type, but does
-require a more random visit order when flooding, which could restrict
-performance on larger images.
-\end{itemize}
+% \section{Finding regional minima}
+% The marker based watershed filter can be used as a convention
+% watershed algorithm by finding the regional minima and using them as
+% markers. Regional minima (and maxima) are often useful, and there are
+% a number of different ways of finding them:
+% \begin{itemize}
+% \item Using reconstruction approaches mentioned earlier - 
+% the problem with these is that a value of 1 is subtracted from the
+% original to produce the image from which the reconstruction is carried
+% out, which makes the approach less useful for floating point
+% types. The advantage of this approach is that it can be easily
+% extended to find {\em image dynamics}, which are often more stable and
+% less subject to noise than regional minima.
+% \item The way ITK does it, which is related to the connected 
+% component labeller. Finds all flat regions and then figures out which
+% are plateaus, minima and maxima by looking at the boundary values.
+% \item Flooding transform (not sure where this came from) which 
+% sets anything not a regional minima to a specified max value. The
+% algorithm assumes that the border is set to the maximum value. Pixels
+% are visited in raster order. If a pixel $A$ has a neighbour $B$ and
+% $P(A) > p(B)$ (i.e grey level of A is greater than B) then all pixels
+% connected to $A$ with value $P(A)$ are set to the maximum value. This
+% is a recursive algorithm that finds pixels that cannot be a regional
+% minimum (because they have a lower neighbour) and then eliminates any
+% connected region that also cannot be a regional minima.
+% 
+% The result would then need to be thresholded and labelled to produce
+% markers. This approach is not limited by the pixel type, but does
+% require a more random visit order when flooding, which could restrict
+% performance on larger images.
+% \end{itemize}
 
@@
 find the label correspondance in ITK, so this step must be performed by
-hand with a code like the one before. We can use a
-\verb$ChangeLabelImageFilter$ to change the labels.
+hand with a code like the one before. The \verb$ChangeLabelImageFilter$
+is well designed for this task:
 \small \begin{verbatim}
@@
 \begin{itemize}
-  \item it doesn't work if there some markers not separated by background
+  \item it doesn't work if there some markers are not separated by background
 pixels - they are considered as a single marker
@@
   \item it's less efficient, as shown in Section \ref{sect:perf}
-  \item it gives poor results with plateaus
+  \item it still gives poor results with plateaus. A lower completion
+sould be performed before to also solve this problem\footnote{Lower
+completion is not implemented yet in ITK}.
 \end{itemize}
@@
 
-{\em Don't think this should be here. Is there any compiler that will
-represent bool images as bit vectors? It might save memory but will
-increase access time}.  We prefered using a separated binary image (of
-type \verb$itk::Image<bool, dim>$) to store this information. This
-solution avoid the need to copy the image at the end, and so should be
-more efficient in computation time, and require only one bit per pixel
-and so should be more memory efficient as well.
+% {\em Don't think this should be here. Is there any compiler that will
+% represent bool images as bit vectors? It might save memory but will
+% increase access time}.  We prefered using a separated binary image (of
+% type \verb$itk::Image<bool, dim>$) to store this information. This
+% solution avoid the need to copy the image at the end, and so should be
+% more efficient in computation time, and require only one bit per pixel
+% and so should be more memory efficient as well.
 
@@
 \verb$MorphologicalWatershedImageFilter$ performs a watershed transformation
-without markers. It encapsulate several filters:
+{\em without} markers. It encapsulate several filters:
 \begin{itemize}
@@
 of small objects. To remove them, we use an {\em opening by reconstruction}
-which have the advantage to preserve the egdes of the object.
+which has the advantage to preserve the egdes of the objects.
 
@@
 There is now only one binary object in the image, but it contains lots of holes
-which will produce an unsable distance map for the watershed. We fill them
+which will produce an unusable distance map for the watershed. We fill them
 with a {\em hole filling} filter. This algorithm, based on a morphological
@@
 binary object must be inverted to compute the distance {\em inside} the
-object, and the distance map must be inverted to used by the watershed.
+object, and the distance map must be inverted to be used by the watershed.
 
@@
 
-The watershed is then performed to split the two pronucleus. A level
-is used to avoid the too small minima in the border of the object and thus
+The watershed is then performed to split the two pronucleus. A \verb$Level$
+is set to avoid the too small minima in the border of the object and thus
 avoid over segmentation. The label produced by the watershed are then
@@
 The watershed extends the labels on the entire image. We use a
-\verb$MaskImageFilter$ to limit the labels to the object. Another commonly
-used method is to use the background as a marker and to perform a watershed
-from markers.
+\verb$MaskImageFilter$ to keep the labels inside the object. Another commonly
+used method to keep the labeled zone in the object is to use the regional
+minima {\em and} the background as markers, to perform a marker based watershed.
 
@@
 The same watershed is done with the ITK watershed filter\footnote{A
-$RelabelComponentImageFilter$ is used here because unsigned long is
-not available in python for the $MaskImageFilter$}.
+$RelabelComponentImageFilter$ is used here because $unsigned long$ is
+not one of the available types for the $MaskImageFilter$ in python}.
 
@@
 \section{Related work}
-This work have lead us to develop some other filters to:
+This work has lead us to develop some other filters to:
 \begin{itemize}

### 2006-06-09 article/Article.tex
Article and bibliography cleanup
Ran spell check over article, corrected grammar. Removed my big bibliography and replaced with a customized one.

darcs-hash:20060609003028-fafb9-24b197713b6513af0f4780eebc5d83f4af4004a2.gz

@@
 known techniques for applying the watershed transform to practical
-problems. The discussion will be focussed on new and existing ITK
+problems. The discussion will be focused on new and existing ITK
 classes.
@@
 
-Section \ref{sect:using-ws-itk} propose some applications of watershed
+Section \ref{sect:using-ws-itk} proposes some applications of watershed
 with ITK.
@@
 falling on a terrain. The idea is that a raindrop falling on a surface
-will trickle down the path of steepest descent to a minima. The set of
-points on the surface that lead to the same minima is known as a
+will trickle down the path of steepest descent to a minimum. The set of
+points on the surface that lead to the same minimum is known as a
 catchment basin and borders between catchment basins are watershed
@@
 meet neighboring lakes. Virtual dams are constructed to keep the
-neighboring lakes separate as the waterlevel rises. When the image
-surface is completely fooded the virtual dams correspond to the
+neighboring lakes separate as the water level rises. When the image
+surface is completely flooded the virtual dams correspond to the
 watershed lines\footnote{In 3D neighboring lakes meet at surfaces.}.
@@
 \cite{Beucher93a,Meyer1994a}. Meyer defined the watershed transform in terms of {\em
-topographic distances} through {\em lower complete} images\footnote{A lower complete
-image is an image where there is no plateaus, excepted plateaus
-which are regional minima. A common way to compute the lower completion of an
-image, is to add the distance from the nearest lower pixel to the pixels of all
-the plateaus. Lower completion have the property to produce the same image than
-the input image if the input image is already lower complete. By lower completing
-an image, we introduce a new order relation in the plateaus pixels which can be
-used to solve the plateaus problem with watershed.}. These
-definitions rigorously defined the behavior in the presence of
-plateaus and lead to some new algorithms based on graph theory. Two
-examples of these algorithms are {\em hill-climbing} and {\em image
-integration} and will be discussed in more detail in this report.
-Figures \ref{fig:cthead1-markers}, \ref{fig:circle} and \ref{fig:difficult-cases} illustrate
-the possible problems with plateaus.
+topographic distances} through {\em lower complete} images\footnote{A
+lower complete image is an image where the only plateaus are regional
+minima. A common way to compute the lower completion of an image, is
+to add the distance from the nearest lower pixel to the pixels of all
+the plateaus. Lower completion have the property to produce the same
+image than the input image if the input image is already lower
+complete. By lower completing an image, we introduce a new order
+relation in the plateaus pixels which can be used to solve the
+plateaus problem with watershed.}. These definitions rigorously
+defined the behavior in the presence of plateaus and lead to some new
+algorithms based on graph theory. Two examples of these algorithms are
+{\em hill-climbing} and {\em image integration} and will be discussed
+in more detail in this report.  Figures \ref{fig:cthead1-markers},
+\ref{fig:circle} and \ref{fig:difficult-cases} illustrate the possible
+problems with plateaus.
 
@@
 \caption{Another example of plateau problem.
-The inside and outside of the ring have the same brightness, but the ring is arbitarily assigned to the outer region by the ITK watershed.
+The inside and outside of the ring have the same brightness, but the ring is arbitrarily assigned to the outer region by the ITK watershed.
 It is split by the morphological watershed.\label{fig:circle}}
@@
 
-It should be evident that a number of arbitary choices have been made
+It should be evident that a number of arbitrary choices have been made
 here - for example it is impossible to break a flat region into
@@
 down the middle). What happens of there are multiple neighboring
-regions with equal height - is the labelling decision now somewhat arbitary?
-
-In general the flooding models avoid these dilemas and lead to more
+regions with equal height - is the labelling decision now somewhat arbitrary?
+
+In general the flooding models avoid these dilemmas and lead to more
 elegant algorithms.
 
-\section{Using the watershed tranform}
+\section{Using the watershed transform}
 \label{sect:usingWT}
@@
 \subfigure[Watershed from markers]{\includegraphics[scale=3.2]{level-markersMf-rgb}}
-\caption{Different results of morphological watershed on an articial image.
-The labeled image are overlayed to the input image with {\em LabelOverlayImageFilter}.
+\caption{Different results of morphological watershed on an artificial image.
+The labeled image are overlayed on the input image with {\em LabelOverlayImageFilter}.
 From $(b)$ to $(f)$, an example of reduction of the classical watershed over segmentation by minima dynamic filtering.
@@
 \subfigure[Watershed from markers]{\includegraphics[scale=3.2]{level-markers-itk-rgb}}
-\caption{Different results of ITK's watershed on an articial image.
-The labeled image are overlayed to the input image with {\em LabelOverlayImageFilter}.
+\caption{Different results of ITK's watershed on an artificial image.
+The labeled image are overlayed on the input image with {\em LabelOverlayImageFilter}.
 From $(b)$ to $(f)$, an example of reduction of the classical watershed over segmentation by minima dynamic filtering.
@@
 This can be achieved by performing a {\em reconstruction by erosion}
-or {\em recursive conditional erosion} from the seleted regional
+or {\em recursive conditional erosion} from the selected regional
 minima. This process is also called {\em minima imposition}. Figure
@@
 regional minimum B. The red plateaus illustrate the parts of the
-result image that are above the orginal. The resulting image has only
+result image that are above the original. The resulting image has only
 on regional minima that corresponds to B.
@@
 are we going to decide which minima are useful to us? The key step to
-using these ideas in practice is to realize that an arbitary mask
+using these ideas in practice is to realize that an arbitrary mask
 image can be used to create minima in an image in locations we want
@@
 \item \label{recipe:erode}Recursively erode the control surface to eliminate all minima besides those introduced in the previous step. This produces a new control surface.
-\item Apply the watershed tranform to the new surface.
+\item Apply the watershed transform to the new surface.
 \end{enumerate}
@@
 algorithms and are examples of graph flooding approaches. These
-algorithms implement the watershed tranform by topographic distance
+algorithms implement the watershed transform by topographic distance
 (or approximate it) and are derived from Dijkstra-Moore minimal path
@@
 need for explicit computation of lower complete images and recursive
-erosion. An ordered queue is a heirarchical priority queue with two
+erosion. An ordered queue is a hierarchical priority queue with two
 levels. One level is defined by the pixel gray level, with darker
@@
 the same grey level and provides a FIFO ordering. The FIFO ordering
-leads to a breadth first propogation across plateaus because the
+leads to a breadth first propagation across plateaus because the
 ordering becomes related to distance from the regional minima. This is
@@
 
-Wavefronts are propogated from each minima, using the ordered queue
+Wavefronts are propagated from each minima, using the ordered queue
 approach. As the wavefront reaches pixel the distance travelled by the
@@
 
-This algorithm take a grayscale image $I$ and a marker image $M$
+This algorithm take a gray scale image $I$ and a marker image $M$
 which contains a set of markers with different labels. It can be
-splitted in to steps:
+split into two steps:
 \begin{enumerate}
@@
 $MAX(a,b)$ returns the maximum of $a$, and $b$.
-$WSLABEL$ is the label used to define background and watershed pixels (usally $0$).
+$WSLABEL$ is the label used to define background and watershed pixels (usually $0$).
 The lower completion is done by the statement (\ref{algo:Beucher:max}).
@@
 lines, and that the watershed lines it produces have a width of 1 pixel.
-Also, the watershed lines form a connected obejct if we consider the
+Also, the watershed lines form a connected object if we consider the
 opposite connectivity than the one used in the algorithm: with a 2D
@@
 $MAX(a,b)$ returns the maximum of $a$, and $b$.
-$WSLABEL$ is the label used to define background and watershed pixels (usally $0$).
+$WSLABEL$ is the label used to define background and watershed pixels (usually $0$).
 The lower completion is done by the statement (\ref{algo:Meyer:max})}
@@
 example is attempting to split a plateau into two regions, and should
-therefore produce a Voronoi tesselation, with the two markers being
+therefore produce a Voronoi tessellation, with the two markers being
 separated by a straight line. This does not happen because various
@@
 the labels in the marker image. There is currently no filter able to
-find the label correspondance in ITK, so this step must be performed by
+find the label correspondence in ITK, so this step must be performed by
 hand with a code like the one before. The \verb$ChangeLabelImageFilter$
@@
   \item it still gives poor results with plateaus. A lower completion
-sould be performed before to also solve this problem\footnote{Lower
+souled be performed before to also solve this problem\footnote{Lower
 completion is not implemented yet in ITK}.
@@
 \begin{itemize}
-  \item a grayscale image which can be set with \verb$SetInput(image)$,
+  \item a gray scale image which can be set with \verb$SetInput(image)$,
 \verb$SetInput(0, image)$ or \verb$SetInput1(image)$.
@@
 by the user in case the user have chosen an unsigned pixel type.
-  \item it requires to use one of the value which can be possibly aready used
+  \item it requires to use one of the value which can be possibly already used
 by a label in the marker image, in case of a signed type.
@@
 pixel type} by using a type which contains a larger range of value than the
-type choosed by the user, but then, we must copy the data to an image of
+type chosen by the user, but then, we must copy the data to an image of
 the type chosen by the user at the end of the algorithm.
@@
 While \verb$RegionalMinimaImageFilter$, \verb$ConnectedComponentImageFilter$,
-and \verb$MorphologicalWatershedFromMarkersImageFilter$ are the usal steps
+and \verb$MorphologicalWatershedFromMarkersImageFilter$ are the usual steps
 of the watershed usage, the \verb$HMinimaImageFilter$ is a convenient way
@@
 \end{tabular}
-\caption{Execution times. Note the FC and MWL refer to FullyConnected and MarkWatershedLine.\label{tbl:perf}}
+\caption{Execution times. Note - FC and MWL refer to FullyConnected and MarkWatershedLine.\label{tbl:perf}}
 \end{table}
@@
 morphological watershed are shown in the first few columns, with the
-time from a combined filter and the itk watershed shown last. It
-should be noted that the itk watershed does not allow connectivity to
+time from a combined filter and the ITK watershed shown last. It
+should be noted that the ITK watershed does not allow connectivity to
 be modified by the user - it always uses face connected mode and
@@
 \end{tabular}
-\caption{Execution times with marker based approach. Note the FC and MWL refer to FullyConnected and MarkWatershedLine.\label{tbl:perf-marker}}
+\caption{Execution times with marker based approach. Note - FC and MWL refer to FullyConnected and MarkWatershedLine.\label{tbl:perf-marker}}
 \end{table}
@@
 \begin{itemize}
-\item Preventing boundary checks - not really the itk way, but if a user is
+\item Preventing boundary checks - not really the ITK way, but if a user is
 willing to throw away the image border by treating it as a watershed
@@
 available so prevent streaming from working well. The existing ITK
-watershed does provide a multithreaded implementation and similar
+watershed does provide a multi threaded implementation and similar
 techniques may be possible with the new implementation. Other
@@
 to get a binary image\footnote{The methods to remove the noise and binarize the
-image are volontary kept simple for this example}.
+image are kept simple for this example}.
 
@@
 of small objects. To remove them, we use an {\em opening by reconstruction}
-which has the advantage to preserve the egdes of the objects.
+which has the advantage to preserve the edges of the objects.
 
@@
 watershed lines, which are a new capability of the filters introduced
-in this paper, as markers. This technnique almost certainly isn't the
+in this paper, as markers. This technique almost certainly isn't the
 best way of segmenting gels, and probably won't work on many gels
@@
 \verb$true$ will result in the watershed line being marked with zero 
-values. In this situation we are using the watershed to tesselate the
+values. In this situation we are using the watershed to tessellate the
 image into regions. The region boundaries will be in the brightest
@@
 spot is found and there are markers placed in what should probably be
-condsidered as background. These problems may be addressed in a number
+considered as background. These problems may be addressed in a number
 of ways, such as performing more careful selection of regional minima
@@
 \bibliographystyle{plain}
-\bibliography{InsightJournal,Refnames.bib}
+\bibliography{InsightJournal,local.bib}
 \nocite{ITKSoftwareGuide}

### 2006-06-09 article/Article.tex
fix affiliation, according to new directives
darcs-hash:20060609081932-2fc9d-d21fa9fefcfee005885854eef9ed262a39cd7bb7.gz

@@
 \authoraddress{{$^1$}Department of Medicine, Monash University, Australia.\\ 
-{$^2$}Unit\'e de Biologie du D\'eveloppement et de la Reproduction, Institut National de la Recherche Agronomique, 78350 Jouy-en-Josas, France}
+{$^2$}INRA, UMR 1198; ENVA; CNRS, FRE 2857, Biologie du D\'eveloppement et 
+Reproduction, Jouy en Josas, F-78350, France}
 
